### 2024-08-27 docs/paper/article.tex
start report & debug mamba output

\documentclass[10pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{geometry}
\usepackage{cite}

% Page layout
\geometry{margin=1.5cm}

% Title page
\title{\Large\textbf{PixelBytes: Catching Unified Embedding for Multimodal Generation}}
\author{\large Fabien Furfaro}
\date{\large 2024}

\begin{document}

\maketitle

\begin{abstract}
This report presents PixelBytes, an innovative multimodal embedding model designed for the simultaneous generation of text and images pixel by pixel. We introduce the PxByEmbed algorithm as a key component of our architecture, enabling efficient representation of mixed sequences of text and images. Our approach leverages web scraping, image processing, and advanced machine learning techniques to create a unique dataset and training pipeline. We compare the performance of RNN, Transformer, and Mamba models in this task, providing insights into the effectiveness of different architectures for multimodal generation.
\end{abstract}

\begin{figure}[h]
\centering
\includegraphics[width=0.8\linewidth]{example_generation.png}
\caption{Example of a generated pixelated Pokémon and its description}
\label{fig:example_generation}
\end{figure}

\section{Introduction}
Multimodal generation, which combines text and images, represents a significant challenge in artificial intelligence. Recent advancements in pixel-by-pixel image generation have shown promising results in creating coherent visual content \cite{parmar2018image}. Plus connu egalmeent les modele de generation de texte comme GPT. Building upon these findings, PixelBytes aims to address the challenge of unified text and image generation by proposing a model capable of producing mixed sequences of text and images in a coherent and unified manner.

Our approach draws inspiration from the state-of-the-art in image transformers, which have demonstrated the ability to generate high-quality images with fine-grained control \cite{parmar2018image}. We also build upon generative sequence models such as GPT, PixelRNN, and PixelCNN. Additionally, we leverage recent developments in bidirectional state space models for time series prediction, which have shown improved performance in capturing long-range dependencies \cite{bimamba}. We also incorporate insights from tokenless selective state space models, which have proven effective in processing sequential data efficiently \cite{mambabyte}.

Notre approche a permis de montrer (...)

\section{Model Architecture}

\subsection{Overview}
PixelBytes integrates several innovative components to achieve its multimodal generation capabilities:
\begin{itemize}
    \item A tokenizer sequence constructor with pixelizable image, and bytes texte
    \item A unified multimodal embedding (PxByEmbed)
\end{itemize}

\subsection{Dataset Construction}

Pour tester notre approche, nous avons besoin d'une base de donnée d'image-captionning. Néanmoins, les bases de données actuelles ne sont pas adapté pour combiner la génération de texte et d'image. Les textes sont souvent trop petit, et une image pixelisé des base de données d'image captionning comme (COCO, Flick) sont difficilement interpretable (les images d'origine sont trop volumineuse). C'est pourquoi, nous avons choisi de construire notre base de donnée à partir des données wiki de Pokemon. Pokemon a l'avantage d'exister depuis plus de 20 ans, lors de l'apparition des jeux video portable, il y a encore des dessins pixelisé ou pixelisable de pokemon. de plus, leur description est une bonne base pour decrire le pokemon. Aujourd'hui, il existe plus de 1000 pokemon, ce qui permet d'avoir une base de donnée minimal pour construire un jeu de données specialisé dans la generation de description et de pixel pokemon.

\subsubsection{Web Scraping from Pokepedia}
We constructed our dataset by scraping miniatures and descriptions from Pokepedia, ensuring a rich and diverse collection of Pokémon data.

Utilisation de beautifull soup.

\subsubsection{Image Pixelation and Text-Image Balance}
The scraped images were pixelated to create a retro aesthetic. We carefully balanced the dataset to maintain a 2/3 text to 1/3 image ratio, ensuring comprehensive training on both modalities.

Utilisation d'OpenCV et skimage.

\subsubsection{Image Quantization and Token Creation}
We quantized the pixelated images using a palette inspired by the NES color scheme (55 colors). This process allowed us to create tokens representing different color and position combinations, effectively translating visual information into a format suitable for sequence modeling.

Utilisation d'OpenCV et skimage. (NES palette via le site web des palettes)

Construction des token d'image

Enregistrement dans le Hub dataset de HuggingFace

\section{Multimodal Embedding Algorithm}

\subsection{PxByEmbed: Multimodal Embedding Algorithm}
At the core of our approach is the PxByEmbed algorithm, which represents mixed sequences of text and images in a unified manner. This algorithm extends classical embedding techniques by incorporating spatial adaptivity, allowing for more effective representation of both textual and visual information.

\begin{algorithm}[h]
\caption{PxByEmbed: Multimodal Embedding Algorithm (k=3)
\newline
\textbf{Input:} $V$: vocabulary size, $D$: embedding dimension
\newline
\textbf{Output:} Embedded representation $\mathbf{E} \in \mathbb{R}^{B \times L \times D}$
\newline
\textbf{Note:} $\mathbf{X}_{emb} \in \mathbb{R}^{B \cdot L \times E_{int} \times k \times k}$, 
$\mathbf{X}_{flat} \in \mathbb{R}^{B \cdot L \times E_{int}k^2}$, 
$\mathbf{X}_{proj} \in \mathbb{R}^{B \cdot L \times D}$
}
\begin{algorithmic}[0]
\State \textbf{Initialize:}
\State $k \gets 3$
\State $E_{int} \gets \max(9, \lfloor D / k^2 \rfloor)$
\State $\mathbf{\alpha} \in \mathbb{R}^{1 \times 1 \times k \times k}$ 
\State $\mathbf{W}_{emb} \in \mathbb{R}^{V \times E_{int}}$ 
\State $\mathbf{W}_{proj} \in \mathbb{R}^{E_{int}k^2 \times D}$ 
\State $\mathbf{W}_{patch} \in \mathbb{R}^{E_{int} \times E_{int} \times k \times k}$

\Function{PxByEmbed}{$\mathbf{X} \in \mathbb{Z}^{B \times L \times k \times k}$}
    \State $\mathbf{X}_{emb} \gets \text{Permute}(\text{Embed}(\mathbf{X}, \mathbf{W}_{emb}), [0, 3, 1, 2])$

    \State $\mathbf{X}_{patch} \gets \text{Conv2D}(\mathbf{X}_{emb}, \mathbf{W}_{patch}, \text{padding}=1)$
    \State $\mathbf{X}_{combined} \gets \sigma(\mathbf{\alpha}) \odot \mathbf{X}_{emb} + (1 - \sigma(\mathbf{\alpha})) \odot \mathbf{X}_{patch}$

    \State $\mathbf{X}_{flat} \gets \text{Flatten}(\mathbf{X}_{combined})$ 
    \State $\mathbf{X}_{proj} \gets \mathbf{X}_{flat}\mathbf{W}_{proj}$ 
    \State $\mathbf{E} \gets \text{LayerNorm}(\mathbf{X}_{proj})$
    \State $\mathbf{E} \gets \text{Reshape}(\mathbf{E}, [B, L, D])$
    \State \Return $\mathbf{E}$
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Managing Transitions}
PixelBytes employs newline characters (ASCII 0A) to manage transitions between text and image, ensuring coherence in the generation of mixed sequences.

\section{Model Variants and Training}

\subsection{Model Architectures}
We implemented and compared three different model architectures:
\begin{itemize}
    \item Recurrent Neural Network (RNN)
    \item Transformer
    \item Mamba (State Space Model)
\end{itemize}

\subsection{Training Process}
Training was conducted on dual T4 GPUs from Kaggle, enabling experimentation with these advanced architectures and our large multimodal dataset.

\subsection{Evaluation Metrics}
We employed various metrics to evaluate the performance of our models, including perplexity, BLEU score for text generation, and structural similarity index (SSIM) for image generation.

\section{Results and Discussion}
[This section will be filled with findings and observations from our experiments, including:
- Performance comparison of RNN, Transformer, and Mamba models
- Ablation studies to understand the importance of each component
- Qualitative analysis of generated embeddings and outputs
- Methods and results of measuring similarity in generated sequences]

\section{Conclusion and Future Directions}
This report on PixelBytes opens new avenues in unified multimodal generation. Our findings suggest promising potential for coherent text-image generation. Future work will focus on refining the PxByEmbed algorithm, exploring larger datasets, and investigating applications in creative content generation.

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2024-08-27 pixelbytes/model.py

@@ -95,7 +95,7 @@ class bMamba(PreTrainedModel):
         # Remaining Mamba layers
         if self.layers:
             for layer in self.layers: x = layer(x)
-        return self.lm_head(x)
+        return self.lm_head(x[:, -1, :])
 
 ### Comparizon model
 # simple lstm (like simplified PixelRNN)

### 2024-08-27 pixelbytes/train.py

@@ -30,6 +30,7 @@ class TrainConfig:
     save_dir : str = "models"
     dataset_name : str = "dataset"
     eval_every : int = 5
+    compile_model : bool = False
 
     def __post_init__(self):
         self.info = "_".join([self.model.name, 
@@ -44,6 +45,7 @@ class Trainer:
         self.config = config
         self.device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
         self.model = self.config.model.to(self.device)
+        self.model = torch.compile(self.model) if config.compile_model else self.model
         self.save_dir = os.path.join(config.save_dir, f"{config.info}_{config.dataset_name}")
         self.hf_dir = os.path.join(config.save_dir, f"{config.info}")
         os.makedirs(self.save_dir, exist_ok=True)

### 2024-08-27 tests.py

@@ -30,7 +30,7 @@ if __name__ == '__main__' :
     # train simple model (one epoch)
     model = SimpleRNNModel(model_config) #SimpleTransformerModel(config)
     
-
+    """
     token = None #input("Input Hugging Face Token: ")
     train_config = TrainConfig(model=model, model_config=model_config, dataset_name="PixelBytes-Pokemon", hf_token=token,
                                train_dataset=train_dataset,test_dataset=test_dataset, num_epochs=2, repo_name="PixelBytes-Pokemon")
@@ -58,7 +58,7 @@ if __name__ == '__main__' :
         reconstructor.process(token_id)
     result = reconstructor.get_result()
     display_result(result)
-    """
+
     """
     train_loader = DataLoader(train_dataset, batch_size=32, shuffle=True)
     

### 2024-08-27 pixelbytes/eval.py
some update

@@ -8,7 +8,7 @@ import torch
 import numpy as np
 import pandas as pd
 from dataclasses import dataclass
-from typing import List
+from typing import List, Dict
 from scipy.spatial.distance import hamming, cosine
 from nltk.translate.bleu_score import sentence_bleu
 from collections import Counter
@@ -18,15 +18,17 @@ from tqdm import tqdm
 class EvaluateMetricConfig:
     metrics: List[str] = ("edit", "hamming", "cosine", "bleu", "diversity", "perplexity")
     batch_size: int = 32
-    num_generated_sequences: int = 100
+    num_input_sequences: int = 100  # Nombre de séquences d'entrée à utiliser
     sequence_length: int = 50
-    vocab: str = "ATGC"  # Pour les séquences ADN
+    vocab: str = None  # Sera déterminé automatiquement
+    output_file: str = "evaluation_results.csv"
 
 class EvaluateMetric:
-    def __init__(self, config: EvaluateMetricConfig, model: torch.nn.Module, dataloader: torch.utils.data.DataLoader):
+    def __init__(self, config: EvaluateMetricConfig, models: Dict[str, torch.nn.Module], dataloader: torch.utils.data.DataLoader):
         self.config = config
-        self.model = model
+        self.models = models
         self.dataloader = dataloader
+        self.vocab = self._determine_vocab()
         self.metric_funcs = {
             "edit": self._edit_distance,
             "hamming": self._hamming_distance,
@@ -36,43 +38,88 @@ class EvaluateMetric:
             "perplexity": self._perplexity_score
         }
 
-    def evaluate(self) -> pd.DataFrame:
-        input_sequences = self._tensor_to_strings(next(iter(self.dataloader)))
-        generated_sequences = self._tensor_to_strings(self.generate_sequences())
-        
-        results = []
-        for gen_seq in tqdm(generated_sequences, desc="Evaluating sequences"):
-            row = {"generated": gen_seq}
-            for metric in self.config.metrics:
-                if metric == "diversity":
-                    row[f"{metric}"] = self.metric_funcs[metric](generated_sequences)
-                elif metric == "perplexity":
-                    row[f"{metric}"] = self.metric_funcs[metric](gen_seq)
-                else:
-                    scores = [self.metric_funcs[metric](input_seq, gen_seq) for input_seq in input_sequences]
-                    row[f"{metric}_min"] = np.min(scores)
-                    row[f"{metric}_avg"] = np.mean(scores)
-            results.append(row)
[…]

### 2024-08-27 pixelbytes/results.py

@@ -1,0 +1,123 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabien
+"""
+
+import seaborn as sns
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+
+# Configuration du style (inchangée)
+plt.style.use('seaborn-whitegrid')
+plt.rcParams.update({
+    'font.family': 'serif',
+    'font.serif': ['Times New Roman'] + plt.rcParams['font.serif'],
+    'font.size': 10,
+    'axes.labelsize': 12,
+    'axes.titlesize': 14,
+    'xtick.labelsize': 10,
+    'ytick.labelsize': 10,
+    'legend.fontsize': 10,
+    'figure.titlesize': 16
+})
+
+def load_csv_files(directory_or_file):
+    data_dict = {}
+    if os.path.isdir(directory_or_file):
+        for file in os.listdir(directory_or_file):
+            if file.endswith('.csv'):
+                model_name = os.path.splitext(file)[0]
+                file_path = os.path.join(directory_or_file, file)
+                data_dict[model_name] = pd.read_csv(file_path)
+    elif os.path.isfile(directory_or_file) and directory_or_file.endswith('.csv'):
+        df = pd.read_csv(directory_or_file)
+        if 'model' in df.columns:
+            for model in df['model'].unique():
+                data_dict[model] = df[df['model'] == model].reset_index(drop=True)
+        else:
+            data_dict['unnamed_model'] = df
+    return data_dict
+
+def save_plot(fig, base_name):
+    for ext in ['svg', 'png']:
+        fig.savefig(f"{base_name}.{ext}", format=ext, dpi=300, bbox_inches='tight')
+    plt.close(fig)
+
+def plot_training_results(train_dfs, save_dir):
+    os.makedirs(save_dir, exist_ok=True)
+    metrics = {
+        'loss': ['train_eval_loss', 'test_loss'],
+        'accuracy': ['train_accuracy', 'test_accuracy'],
+        'f1': ['train_f1', 'test_f1']
+    }
+    
+    for metric_name, columns in metrics.items():
+        fig, ax = plt.subplots(figsize=(10, 6))
+        for model_name, df in train_dfs.items():
+            for col in columns:
+                if col in df.columns:
+                    sns.lineplot(data=df, x='epoch', y=col, ax=ax, label=f'{model_name}_{col}', linewidth=2, marker='o', markersize=4)
+        ax.set_title(f'{metric_name.capitalize()} over Epochs')
[…]

### 2024-08-27 tests.py

@@ -37,7 +37,8 @@ if __name__ == '__main__' :
     trainer = Trainer(train_config)
     trainer.train_and_evaluate()
     """
-    model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    #model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
     # generate and display
     generator = SequenceGenerator(model, tokenizer)
     reconstructor = SequenceReconstructor(tokenizer)

### 2024-08-27 pixelbytes/__init__.py
some update

@@ -5,3 +5,5 @@ from .pokemon import *
 from .display import *
 from .tokenizer import *
 from .train import *
+from .eval import *
+from .results import *

### 2024-08-28 pixelbytes/display.py
update generator and all train

@@ -9,63 +9,271 @@ import matplotlib.pyplot as plt
 import warnings
 warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
 
-class SequenceReconstructor:
-    def __init__(self, tokenizer, min_row_length=3):
-        self.tokenizer = tokenizer
-        self.min_row_length = min_row_length
-        self.elements = []
-        self.current = {'type': None, 'data': []}
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
 
-    def process(self, token_id):
-        token = self.tokenizer.ids_to_tokens[token_id]
-        
-        if isinstance(token, tuple) and len(token) == 3:
-            self._process_element('image', token)
-        elif isinstance(token, bytes):
-            self._process_element('text', token)
+def reconstruct_sequence(tokens_arr_obj, min_row_length=3, max_gap=50, max_image_gap=100):
+    def reconstruct_imgs(img_tokens, start_index):
+        img_idx = np.arange(len(img_tokens)) + start_index
+        big_gaps = np.where(np.diff(img_idx) > max_gap)[0] + 1
+        image_splits = np.split(img_idx, big_gaps)
+        images = []
+        for split in image_splits:
+            if len(split) < min_row_length:
+                continue
+            row_breaks = np.concatenate(([0], np.where(np.diff(split) > 1)[0] + 1, [len(split)]))
+            row_lengths = np.diff(row_breaks)
+            valid_rows = row_lengths >= min_row_length
+            if not np.any(valid_rows):
+                continue
+            most_common_length = np.bincount(row_lengths[valid_rows]).argmax()
+            num_rows = np.sum(valid_rows)
+            img = np.full((num_rows, most_common_length, 3), 255, dtype=np.uint8)
+            valid_indices = split[np.concatenate([np.arange(start, end) for start, end, valid 
+                                                  in zip(row_breaks[:-1], row_breaks[1:], valid_rows) 
+                                                  if valid])]
+            img_flat = img.reshape(-1, 3)
+            img_flat[:len(valid_indices) - start_index] = [tokens_arr_obj[i] for i in valid_indices]
+            images.append({'image': img, 'start_index': split[0], 'end_index': split[-1]})
+        return images
 
-    def _process_element(self, type, data):
-        if self.current['type'] != type:
-            self._flush_current()
-            self.current['type'] = type
-        self.current['data'].append(data)
-        
[…]

### 2024-08-28 pixelbytes/generator.py

@@ -2,95 +2,85 @@
 # -*- coding: utf-8 -*-
 """
 @author: fabien
-In dev.. (process_token not really working)
-https://www.perplexity.ai/search/tu-peux-me-dire-combien-obtien-z6lt6R6IRrmwnILyCqwZxw
 """
-import torch
-import torch.nn.functional as F
+import numpy as np
 
 class SequenceGenerator:
-    def __init__(self, model, tokenizer):
-        self.model = model
+    def __init__(self, tokenizer):
         self.tokenizer = tokenizer
         self.newline_token = self.tokenizer.vocab[b'\n']
         self.tab_token = self.tokenizer.vocab[b'\t']
+        self.sequence = None
+        self.clock = None
 
-    def generate_complete(self, start_sequence, max_length=100, temperature=1.0):
-        current_sequence, sequence_clock = self.ids_to_matrices(start_sequence)
-        generated_sequence = current_sequence.clone()
-        
-        for _ in range(max_length):
-            next_token_id = self._generate_next_token(current_sequence, temperature)
-            next_matrix, sequence_clock = self._process_token(next_token_id, current_sequence, sequence_clock)
-            current_sequence = torch.cat([current_sequence[1:], next_matrix.unsqueeze(0)])
-            generated_sequence = torch.cat([generated_sequence, next_matrix.unsqueeze(0)])
+    def reset(self, start_sequence, line_period=64):
+        # shape : start_sequence : (L>50, M=3, N=3) : long --> input of model
+        L, M, N = start_sequence.shape
+        if L < line_period : print(f"Input sequence dimensions should be > 2*25 for image minimum, but got {L}, choose 64 (images line)")
+        # clone sequence and find clock
+        self.sequence = start_sequence.copy()
+        self.clock = self._find_clock(self.sequence)
 
-        return generated_sequence.cpu().numpy()
-
-    def generate_stream(self, start_sequence, max_length=100, temperature=1.0):
-        current_sequence, sequence_clock = self.ids_to_matrices(start_sequence)
-        for matrix in current_sequence:
-            yield matrix.cpu().numpy()
-
-        for _ in range(max_length):
-            next_token_id = self._generate_next_token(current_sequence, temperature)
-            next_matrix, sequence_clock = self._process_token(next_token_id, current_sequence, sequence_clock)
[…]

### 2024-08-28 pixelbytes/model.py

@@ -51,16 +51,18 @@ class PxByEmbed(nn.Module):
 
 ### Main model
 @dataclass
+model_config = ModelConfig(dim=81, d_state=64, depth=2, vocab_size=vocab_size)
+
 class ModelConfig(PretrainedConfig):
-    dim : int # The input dimension of the input tensor. (embedding dim output)
+    dim : int #81 # The input dimension of the input tensor. (embedding dim output)
     model_type: str = "sequence-generator"
     pembed : bool = True # convolutionnal embedding
     pxby_embed : bool = True # PixelBytes or only center sequence
     bidirectional : bool = True # For RNN or SSM model
-    d_state: int = 16 # The dimension of the state space model
+    d_state: int = 64 # The dimension of the state space model (or hidden state for RNN)
     d_conv : int = 4 # The convolutionnal windows
     expand: int = 2 # E in paper/comments
-    depth : int = 3 # The number of residual layers
+    depth : int = 2 # The number of residual layers
     vocab_size : int = 113 # ASCII bytes + NES Pixel
     
     def __init__(self, dim: int = None, **kwargs):

### 2024-08-28 pixelbytes/results.py

@@ -9,8 +9,9 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import os
 
-# Configuration du style (inchangée)
-plt.style.use('seaborn-whitegrid')
+# Configuration du style
+plt.style.use('default')  # Utilisation du style par défaut de Matplotlib
+sns.set_theme()  # Application des paramètres de base de Seaborn
 plt.rcParams.update({
     'font.family': 'serif',
     'font.serif': ['Times New Roman'] + plt.rcParams['font.serif'],
@@ -20,8 +21,16 @@ plt.rcParams.update({
     'xtick.labelsize': 10,
     'ytick.labelsize': 10,
     'legend.fontsize': 10,
-    'figure.titlesize': 16
+    'figure.titlesize': 16,
+    'axes.grid': True,
+    'grid.alpha': 0.3,
+    'axes.axisbelow': True,
+    'axes.facecolor': '#EAEAF2',
+    'axes.edgecolor': 'none',
+    'axes.linewidth': 0,
+    'figure.facecolor': 'white'
 })
+
 
 def load_csv_files(directory_or_file):
     data_dict = {}

### 2024-08-28 tests.py

@@ -8,12 +8,71 @@ TESTS FILE
 from pixelbytes import *
 from datasets import load_dataset
 
+
+
 ### basic test
 if __name__ == '__main__' :
     # train part
-    
     hf_dataset = load_dataset("ffurfaro/PixelBytes-Pokemon")
+    tokenizer = PixelBytesTokenizer()
+    generator = SequenceGenerator(tokenizer)
+    # initialize sequence (ARRAY)
+    first_sequence = np.array(hf_dataset["train"]['pixelbyte'][0])
+    for start in np.arange(0,1000,50) :
+        start_lenght, sup_lenght = 64, 50
+        # init
+        reference_sequence = first_sequence[start: start + start_lenght + sup_lenght]
+        start_sequence = reference_sequence[:start_lenght]
+        end_sequence = reference_sequence[start_lenght : start_lenght + sup_lenght]
+        # reset
+        generator.reset(start_sequence)
+        generated_sequence = generator.update_sequence(end_sequence[:,1,1])
+        print(np.mean(reference_sequence - generated_sequence)) ## all zero with 50 lenght ! prefer 64 to be large
+    '''
+    tokenizer = PixelBytesTokenizer()
+    model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    #model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    # generate and display
+    generator = SequenceGenerator(tokenizer)
+    reconstructor = SequenceReconstructor(tokenizer)
+    # generate complete sequence
+
+    complete_sequence = generator.generate_complete(start_sequence, max_length=0, temperature=0.7) 
+    # reconstruction 2D complexe ! car si on commence au milieu de l'image, il est compliqué de savoir les bords !
+    print(complete_sequence)
+    # Génération en streaming
+    print("\nGénération en streaming (incluant la séquence initiale):")
+    #for matrix in generator.generate_stream(start_sequence, max_length=25, temperature=0.7):
+    #    print(matrix)
+    # display generated sequence
+    """
+    # generate complete sequence
+    first_sequence = np.array(hf_dataset["train"]['pixelbyte'][0])[:,1,1]
[…]

### 2024-08-29 pixelbytes/dataset.py
change package requirement

@@ -165,6 +165,17 @@ def kmean_quantization(img, num_colors) :
     quantized = centers[labels.flatten()]
     return cv2.cvtColor(quantized.reshape(img.shape), cv2.COLOR_Lab2BGR)
 
+def image_edging_enhancer(img) :
+    gray_image = cv2.cvtColor(img, cv2.COLOR_RGB2GRAY)
+    edges = cv2.Canny(gray_image, 100, 200)
+    # kernel & dilatation
+    kernel_size = int(min(img.shape[0], img.shape[1]) * 0.005)  # 0,5% of image size
+    kernel = cv2.getStructuringElement(cv2.MORPH_RECT, (kernel_size, kernel_size))
+    dilated_edges = cv2.dilate(edges, kernel)
+    # border dark coloring
+    img[dilated_edges != 0] = [0, 0, 0]
+    return img
+
 def image_paletization(img, palette) :
     palette_rgb = palette.copy()
     palette = color.rgb2lab(palette_rgb.reshape(1, -1, 3) / 255.0).reshape(-1, 3)
@@ -176,7 +187,7 @@ def image_paletization(img, palette) :
         return palette_rgb[np.argmin(distances)]
     return np.apply_along_axis(closest_color, 2 , img, palette)
     
-def image_pixelization(img, palette, max_size=25):
+def image_pixelization(img, palette, max_size=25, edging=False):
     # Parameter
     h, w = img.shape[:2]
     num_colors = len(palette)
@@ -186,6 +197,8 @@ def image_pixelization(img, palette, max_size=25):
     # Remove alpha (if)
     if img.shape[2] == 4:
         img = alpha_to_blank(img)
+    # edging (experimental for big img)
+    img = image_edging_enhancer(img) if edging else img
     # First resizing (a=2)
     img = resize_image(img, h, w, m=max_size, a=2)
     ## Quantize image

### 2024-08-29 pixelbytes/model.py

@@ -13,7 +13,8 @@ from dataclasses import dataclass
 try :
     from mamba_ssm.modules.mamba_simple import Mamba
 except :
-    print('No mamba_ssm modules installed...')
+    print('No mamba_ssm CUDA modules installed...')
+    from mambapy.mamba import Mamba
 
 ### Multimodal embedding        
 class PxByEmbed(nn.Module):
@@ -51,8 +52,6 @@ class PxByEmbed(nn.Module):
 
 ### Main model
 @dataclass
-model_config = ModelConfig(dim=81, d_state=64, depth=2, vocab_size=vocab_size)
-
 class ModelConfig(PretrainedConfig):
     dim : int #81 # The input dimension of the input tensor. (embedding dim output)
     model_type: str = "sequence-generator"

### 2024-08-29 pixelbytes/dataset.py
change name of displaysss

@@ -173,8 +173,9 @@ def image_edging_enhancer(img) :
     kernel = cv2.getStructuringElement(cv2.MORPH_RECT, (kernel_size, kernel_size))
     dilated_edges = cv2.dilate(edges, kernel)
     # border dark coloring
-    img[dilated_edges != 0] = [0, 0, 0]
-    return img
+    img_ = img.copy()
+    img_[dilated_edges != 0] = [0, 0, 0]
+    return img_
 
 def image_paletization(img, palette) :
     palette_rgb = palette.copy()

### 2024-08-29 pixelbytes/__init__.py
update generator and all train

@@ -2,7 +2,7 @@ from .dataset import *
 from .generator import *
 from .model import *
 from .pokemon import *
-from .display import *
+from .displays import *
 from .tokenizer import *
 from .train import *
 from .eval import *

### 2024-08-29 pixelbytes/eval.py
update eval code

@@ -1,165 +1,88 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
 """
-@author: fabien
+@author: fabienfrfr
 """
 
 import torch
 import numpy as np
 import pandas as pd
 from dataclasses import dataclass
-from typing import List, Dict
 from scipy.spatial.distance import hamming, cosine
 from nltk.translate.bleu_score import sentence_bleu
-from collections import Counter
 from tqdm import tqdm
+import os
 
 @dataclass
 class EvaluateMetricConfig:
-    metrics: List[str] = ("edit", "hamming", "cosine", "bleu", "diversity", "perplexity")
-    batch_size: int = 32
-    num_input_sequences: int = 100  # Nombre de séquences d'entrée à utiliser
-    sequence_length: int = 50
-    vocab: str = None  # Sera déterminé automatiquement
-    output_file: str = "evaluation_results.csv"
+    input_seq_length: int = 64
+    generation_length: int = 16
+    output_dir: str = "eval_results"
+    device: str = "cuda" if torch.cuda.is_available() else "cpu"
 
 class EvaluateMetric:
-    def __init__(self, config: EvaluateMetricConfig, models: Dict[str, torch.nn.Module], dataloader: torch.utils.data.DataLoader):
+    def __init__(self, config: EvaluateMetricConfig, data: np.ndarray, tokenizer):
         self.config = config
-        self.models = models
-        self.dataloader = dataloader
-        self.vocab = self._determine_vocab()
-        self.metric_funcs = {
-            "edit": self._edit_distance,
-            "hamming": self._hamming_distance,
-            "cosine": self._cosine_distance,
-            "bleu": self._bleu_score,
-            "diversity": self._diversity_score,
-            "perplexity": self._perplexity_score
-        }
+        self.data = data
+        self.model = None
+        self.device = torch.device(config.device)
+        self.sequence_generator = SequenceGenerator(tokenizer)
+        self.df = None
+        os.makedirs(self.config.output_dir, exist_ok=True)
 
-    def _determine_vocab(self):
-        # Déterminer le vocabulaire à partir d'un échantillon de données
-        sample = next(iter(self.dataloader))
-        if isinstance(sample, torch.Tensor):
-            sample = self._tensor_to_strings(sample)
-        return ''.join(sorted(set(''.join(sample))))
+    def reset(self, model: torch.nn.Module):
+        self.model = model.to(self.device).eval()
 
-    def evaluate(self) -> pd.DataFrame:
[…]

### 2024-08-29 pixelbytes/results.py

@@ -49,6 +49,14 @@ def load_csv_files(directory_or_file):
             data_dict['unnamed_model'] = df
     return data_dict
 
+def combine_eval_results(self):
+    all_results = []
+    for file in os.listdir(self.config.output_dir):
+        if file.endswith(".csv"):
+            df = pd.read_csv(os.path.join(self.config.output_dir, file))
+            all_results.append(df)
+    return pd.concat(all_results, ignore_index=True)
+
 def save_plot(fig, base_name):
     for ext in ['svg', 'png']:
         fig.savefig(f"{base_name}.{ext}", format=ext, dpi=300, bbox_inches='tight')

### 2024-08-29 pixelbytes/eval.py
update eval code

@@ -10,6 +10,8 @@ import pandas as pd
 from dataclasses import dataclass
 from scipy.spatial.distance import hamming, cosine
 from nltk.translate.bleu_score import sentence_bleu
+from Levenshtein import distance as levenshtein_distance
+from dtaidistance import dtw
 from tqdm import tqdm
 import os
 
@@ -21,14 +23,14 @@ class EvaluateMetricConfig:
     device: str = "cuda" if torch.cuda.is_available() else "cpu"
 
 class EvaluateMetric:
-    def __init__(self, config: EvaluateMetricConfig, data: np.ndarray, tokenizer):
+    def __init__(self, config: EvaluateMetricConfig, data_pxby, tokenizer):
         self.config = config
-        self.data = data
+        self.data = [np.array(d) for d in data_pxby] 
         self.model = None
         self.device = torch.device(config.device)
         self.sequence_generator = SequenceGenerator(tokenizer)
-        self.df = None
         os.makedirs(self.config.output_dir, exist_ok=True)
+        self.df = None
 
     def reset(self, model: torch.nn.Module):
         self.model = model.to(self.device).eval()
@@ -63,13 +65,21 @@ class EvaluateMetric:
 
             for seq_idx, (input_seq, target_seq) in enumerate(zip(inputs, targets)):
                 generated_seq = self._generate_sequence(input_seq)
+                
+                target_flat = target_seq.flatten()
+                generated_flat = generated_seq.flatten()
+                target_str = ''.join(map(str, target_flat))
+                generated_str = ''.join(map(str, generated_flat))
+                
                 results.append({
                     "row": row_idx,
                     "n_seq": seq_idx,
                     "model": model_name,
-                    "hamming": hamming(generated_seq.flatten(), target_seq.flatten()),
-                    "cosine": cosine(generated_seq.flatten(), target_seq.flatten()),
-                    "bleu": sentence_bleu([list(map(str, target_seq.flatten()))], list(map(str, generated_seq.flatten())))
+                    "hamming": hamming(generated_flat, target_flat),
+                    "cosine": cosine(generated_flat, target_flat),
+                    "bleu": sentence_bleu([list(map(str, target_flat))], list(map(str, generated_flat))),
+                    "levenshtein": levenshtein_distance(target_str, generated_str),
+                    "dtw": dtw.distance(target_flat, generated_flat)
                 })
 
         self.df = pd.DataFrame(results)

### 2024-08-30 pixelbytes/eval.py
debug eval code

@@ -4,14 +4,14 @@
 @author: fabienfrfr
 """
 
+from .generator import SequenceGenerator
+
 import torch
 import numpy as np
 import pandas as pd
 from dataclasses import dataclass
 from scipy.spatial.distance import hamming, cosine
 from nltk.translate.bleu_score import sentence_bleu
-from Levenshtein import distance as levenshtein_distance
-from dtaidistance import dtw
 from tqdm import tqdm
 import os
 
@@ -37,15 +37,15 @@ class EvaluateMetric:
 
     def _generate_sequence(self, input_seq: np.ndarray) -> np.ndarray:
         self.sequence_generator.reset(input_seq)
-        input_tensor = torch.from_numpy(input_seq).float().to(self.device)
-        
+        input_tensor = torch.from_numpy(input_seq).long().to(self.device)
+
         with torch.no_grad():
             for _ in range(self.config.generation_length):
                 output = self.model(input_tensor.unsqueeze(0))
-                next_token = output.squeeze().cpu().numpy()
-                self.sequence_generator.update_sequence(next_token)
-                input_tensor = torch.from_numpy(self.sequence_generator.sequence[-self.config.input_seq_length:]).float().to(self.device)
-
+                _, predicted = output.max(1)
+                next_token_id = predicted.squeeze().cpu().numpy().item() # ensure is not numpy object
+                self.sequence_generator.update_sequence(next_token_id)
+                input_tensor = torch.from_numpy(self.sequence_generator.sequence).long().to(self.device)
         return self.sequence_generator.sequence[-self.config.generation_length:]
 
     def evaluate(self, model_name: str) -> pd.DataFrame:
@@ -68,31 +68,15 @@ class EvaluateMetric:
                 
                 target_flat = target_seq.flatten()
                 generated_flat = generated_seq.flatten()
-                target_str = ''.join(map(str, target_flat))
-                generated_str = ''.join(map(str, generated_flat))
                 
                 results.append({
                     "row": row_idx,
                     "n_seq": seq_idx,
-                    "model": model_name,
                     "hamming": hamming(generated_flat, target_flat),
-                    "cosine": cosine(generated_flat, target_flat),
+                    "cosine": 1 - cosine(generated_flat, target_flat),
                     "bleu": sentence_bleu([list(map(str, target_flat))], list(map(str, generated_flat))),
[…]

### 2024-08-30 pixelbytes/generator.py

@@ -23,7 +23,7 @@ class SequenceGenerator:
 
     def update_sequence(self, tokens_ids):
         # shape : (1) --> output of model
-        if not hasattr(tokens_ids, '__iter__'):
+        if not(hasattr(tokens_ids, '__iter__')):
             tokens_ids = [tokens_ids]
         for token_id in tokens_ids :
             next_matrix, self.clock = self._process_token(token_id, self.sequence, self.clock)
@@ -53,7 +53,6 @@ class SequenceGenerator:
         # index calculation for sequence ref
         a, b = sequence_clock[-2:]  # step
         i, gap = a + len(seq) - b - 1, b - a  # (0-based)
-    
         def safe_get(seq, index,default=0):
             return seq[index][1,1].item() if 0 <= index < len(seq) else default
     
@@ -75,7 +74,7 @@ class SequenceGenerator:
             elif isinstance(prev_token, bytes):
                 matrix = np.array([[0, 0, 0], [0, token_id, 0], [0,0,0]]) # warning here : change following bytes token, not \n, can cause problem during next generation
                 if prev_token != b'\n' : sequence_clock.append(len(seq))
-            elif gap <= 3:
+            elif len(seq) - b > gap : # line2 > line 1
                 matrix = np.array([[0, 0, 0], [prev_token_id, token_id, 0], [0,0,0]])
             else:
                 matrix = np.array([[safe_get(seq, i), safe_get(seq, i+1), safe_get(seq, i+2)],

### 2024-08-30 tests.py

@@ -9,11 +9,29 @@ from pixelbytes import *
 from datasets import load_dataset
 
 
-
 ### basic test
 if __name__ == '__main__' :
-    # train part
+    # common part
     hf_dataset = load_dataset("ffurfaro/PixelBytes-Pokemon")
+
+
+    # evaluation part
+    #config = EvaluateMetricConfig(output_dir="models/eval_results")
+    data_columns = hf_dataset["train"]['pixelbyte']
+    tokenizer = PixelBytesTokenizer()
+    evaluator = EvaluateMetric(config, data_columns, tokenizer)
+    
+    ## manual loop part
+    model_name = "rnn_bi_pxby_81_dim_64_state_2_layer_last"
+    model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder=model_name)  # Your actual model here
+    evaluator.reset(model)
+    results = evaluator.evaluate(model_name)
+    print(results.describe())
+
+
+
+    """    
+    # generation test
     tokenizer = PixelBytesTokenizer()
     generator = SequenceGenerator(tokenizer)
     # initialize sequence (ARRAY)
@@ -28,6 +46,8 @@ if __name__ == '__main__' :
         generator.reset(start_sequence)
         generated_sequence = generator.update_sequence(end_sequence[:,1,1])
         print(np.mean(reference_sequence - generated_sequence)) ## all zero with 50 lenght ! prefer 64 to be large
+    """
+
     '''
     tokenizer = PixelBytesTokenizer()
     model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")

### 2024-08-30 docs/paper/article.tex
adding eval result and strat display and report

@@
 \begin{abstract}
-This report presents PixelBytes, an innovative multimodal embedding model designed for the simultaneous generation of text and images pixel by pixel.
-We introduce the PxByEmbed algorithm as a key component of our architecture, enabling efficient representation of mixed sequences of text and images.
-Our approach leverages web scraping, image processing, and advanced machine learning techniques to create a unique dataset and training pipeline.
-We compare the performance of RNN, Transformer, and Mamba models in this task, providing insights into the effectiveness of different architectures for multimodal generation.
+This report introduces PixelBytes, a multimodal embedding model designed for generating text and pixelated images simultaneously.
+We present the PxByEmbed algorithm, which facilitates the efficient representation of mixed sequences.
+Our approach utilizes web-scraped Pokémon data, image processing techniques, and machine learning methods to create a specialized dataset and training pipeline.
+The dataset balances text and pixelated images, ensuring an equilibrium in token representation.
+We constructed it by scraping data from Pokepedia, applying pixelation using OpenCV and scikit-image, and quantizing images with a NES-inspired color palette.
+We evaluate various model architectures, including Recurrent Neural Networks (RNNs), Transformers, and Mamba models, using metrics such as Hamming distance, cosine similarity, and BLEU scores.
+Initial results suggest that bidirectional RNN models with PxBy embedding and convolutional layers perform well in generating coherent sequences.
+Our findings also indicate that compact models can be effective, challenging the assumption that larger models are always superior.
+Additionally, we observe robustness to input size variations and potential benefits from increased model depth.
+These insights contribute to the ongoing exploration of model architecture design and the trade-offs between complexity and performance in multimodal generation tasks.
 \end{abstract}
@@
 \section{Introduction}
-Multimodal generation, which combines text and images, represents a significant challenge in artificial intelligence.
-Recent advancements in pixel-by-pixel image generation have shown promising results in creating coherent visual content \cite{parmar2018image}.
-Plus connu egalmeent les modele de generation de texte comme GPT.
+
+Multimodal sequence generation, combining diverse data types such as text and numerical sequences, presents a significant challenge in artificial intelligence \cite{baltrusaitis2019multimodal}.
+While models like GPT have excelled in text generation \cite{brown2020language}, there's a growing need for unified approaches that can seamlessly handle varied data types.
 Building upon these findings, PixelBytes aims to address the challenge of unified text and image generation by proposing a model capable of producing mixed sequences of text and images in a coherent and unified manner.
+We draw inspiration from state-of-the-art sequence models, including Image Transformers \cite{parmar2018image}, and recent developments in bidirectional architectures \cite{schuster1997bidirectional, bimamba} and embedding techniques \cite{mikolov2013efficient, mambabyte}.
+We explore various architectures including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on:
 
-Our approach draws inspiration from the state-of-the-art in image transformers, which have demonstrated the ability to generate high-quality images with fine-grained control \cite{parmar2018image}.
-We also build upon generative sequence models such as GPT, PixelRNN, and PixelCNN.
-Additionally, we leverage recent developments in bidirectional state space models for time series prediction, which have shown improved performance in capturing long-range dependencies \cite{bimamba}.
-We also incorporate insights from tokenless selective state space models, which have proven effective in processing sequential data efficiently \cite{mambabyte}.
+\begin{itemize}
+    \item The effectiveness of bidirectional processing
+    \item Novel embedding techniques, particularly PxBy embedding
+    \item The impact of convolutional layers
+    \item Effects of model depth, input dimensionality, and model size
+\end{itemize}
 
-Notre approche a permis de montrer (...)
+Our experiments reveal that compact bidirectional RNN models with PxBy embedding and convolutional layers demonstrate superior performance in generating coherent sequences.
+This finding challenges the notion that larger models are always better, aligning with recent research on model efficiency \cite{tan2019efficientnet}.
+We evaluate our models using Hamming distance, cosine similarity, and BLEU scores, providing a comprehensive assessment of sequence similarity and quality \cite{papineni2002bleu}.
+Our results show robustness to input size variations and highlight the benefits of increased model depth in certain scenarios.
+This paper presents our methodology, experimental results, and analysis, contributing to the development of more effective and versatile sequence generation models.
 
@@
 \begin{itemize}
-    \item A tokenizer sequence constructor with pixelizable image, and bytes texte
+    \item A tokenizer sequence constructor for pixelated images and byte-level text representation
     \item A unified multimodal embedding (PxByEmbed)
+    \item A bidirectional RNN with convolutional layers for sequence processing
 \end{itemize}
+
+This architecture is inspired by recent advancements in multimodal learning \cite{baltrusaitis2019multimodal} and efficient sequence modeling \cite{vaswani2017attention}.
 
@@
 
-Pour tester notre approche, nous avons besoin d'une base de donnée d'image-captionning.
-Néanmoins, les bases de données actuelles ne sont pas adapté pour combiner la génération de texte et d'image.
-Les textes sont souvent trop petit, et une image pixelisé des base de données d'image captionning comme (COCO, Flick) sont difficilement interpretable (les images d'origine sont trop volumineuse).
-C'est pourquoi, nous avons choisi de construire notre base de donnée à partir des données wiki de Pokemon.
-Pokemon a l'avantage d'exister depuis plus de 20 ans, lors de l'apparition des jeux video portable, il y a encore des dessins pixelisé ou pixelisable de pokemon.
-de plus, leur description est une bonne base pour decrire le pokemon.
-Aujourd'hui, il existe plus de 1000 pokemon, ce qui permet d'avoir une base de donnée minimal pour construire un jeu de données specialisé dans la generation de description et de pixel pokemon.
+Existing image-captioning datasets often prove unsuitable for combined text and image generation due to limited text content and difficulties in interpreting pixelated versions of high-resolution images.
+To address this, we developed a specialized dataset based on Pokémon data, offering advantages such as a long-standing franchise history, availability of pixelated designs, rich descriptive text, and a diverse collection of over 1000 unique Pokémon.
+This approach aligns with recent trends in creating specialized datasets for multimodal learning tasks \cite{lin2014microsoft}.
 
-\subsubsection{Web Scraping from Pokepedia}
-We constructed our dataset by scraping miniatures and descriptions from Pokepedia, ensuring a rich and diverse collection of Pokémon data.
+We constructed our dataset by web scraping miniatures and descriptions from Pokepedia using Beautiful Soup \cite{richardson2007beautiful}.
+The scraped images were then pixelated using OpenCV and scikit-image libraries \cite{bradski2000opencv, vanderwalt2014scikit} to create a retro aesthetic.
+We maintained a 2/3 text to 1/3 image ratio in the dataset to ensure comprehensive training on both modalities.
 
-Utilisation de beautifull soup.
+For image quantization, we used a palette inspired by the NES color scheme (55 colors), creating tokens that represent different color and position combinations.
+This approach effectively translates visual information into a format suitable for sequence modeling, similar to recent work in discrete representation learning for images \cite{van2017neural}.
+The quantization process was implemented using OpenCV and scikit-image, with the NES palette sourced from web-based color palette resources.
+To ensure a balanced representation, we carefully adjusted the number of image and text tokens, resulting in a total of 113 characters for each modality.
+This equilibrium allows the model to learn equally from both visual and textual information during training.
 
-\subsubsection{Image Pixelation and Text-Image Balance}
-The scraped images were pixelated to create a retro aesthetic. We carefully balanced the dataset to maintain a 2/3 text to 1/3 image ratio, ensuring comprehensive training on both modalities.
-
-Utilisation d'OpenCV et skimage.
-
-\subsubsection{Image Quantization and Token Creation}
-We quantized the pixelated images using a palette inspired by the NES color scheme (55 colors).
-This process allowed us to create tokens representing different color and position combinations, effectively translating visual information into a format suitable for sequence modeling.
-
-Utilisation d'OpenCV et skimage. (NES palette via le site web des palettes)
-
-Construction des token d'image
-
-Enregistrement dans le Hub dataset de HuggingFace
+The final dataset, balancing text and pixelated images, was uploaded to the Hugging Face Datasets Hub for easy access and reproducibility.
 
@@
 \subsection{PxByEmbed: Multimodal Embedding Algorithm}
+
 At the core of our approach is the PxByEmbed algorithm, which represents mixed sequences of text and images in a unified manner.
 This algorithm extends classical embedding techniques by incorporating spatial adaptivity, allowing for more effective representation of both textual and visual information.
+PxByEmbed builds upon recent advancements in multimodal embeddings \cite{kiela2018learning} and adapts them for the specific needs of pixel-level image representation and byte-level text encoding.
+The algorithm employs a learned embedding matrix that maps each token (whether text or image) to a high-dimensional vector space, while preserving spatial relationships for image tokens.
 
@@
 \subsection{Managing Transitions}
-PixelBytes employs newline characters (ASCII 0A) to manage transitions between text and image, ensuring coherence in the generation of mixed sequences.
 
-\section{Model Variants and Training}
+PixelBytes employs a refined approach to manage transitions between text and image tokens, ensuring coherence during mixed sequence generation.
+This strategy is applied in both dataset construction and sequence generation.
+
+In the dataset construction phase, we utilize a 2D input sequence method that creates a context window around each token.
+This method features a configurable 3x3 context window, special separator tokens for modality transitions, and padding to maintain consistent context sizes.
+Additionally, future tokens are masked to introduce time asymmetry.
+The \texttt{input\_seq\_construct} function implements these strategies by padding the input array and generating a 3D array where each slice represents the context of a token.
+This enables the model to learn spatial relationships, which is vital for maintaining coherence in mixed sequences.
+
+During sequence generation, transitions are managed dynamically through the \texttt{\_process\_token} function.
+This function differentiates between text tokens (bytes) and image tokens (tuples) while handling newline (\texttt{\textbackslash n}) and tab (\texttt{\textbackslash t}) characters for line breaks and indentation.
+It also maintains a \texttt{sequence\_clock} to track progression and manage transitions.
+Each new token's context is represented in a 3x3 matrix, informing the model of its position relative to surrounding tokens.
+
+This dynamic approach allows PixelBytes to maintain coherence when transitioning between text and image modalities, ensuring that generated sequences respect the structural integrity of both text and pixel-based images.
+
+\section{Training and Evaluation}
 
 \subsection{Model Architectures}
-We implemented and compared three different model architectures:
+We implemented and compared three different model architectures: Recurrent Neural Network (RNN), specifically using bidirectional Long Short-Term Memory (LSTM) units \cite{hochreiter1997long}, Transformer \cite{vaswani2017attention}, and Mamba, which is based on the State Space Model (SSM) \cite{gu2022efficiently}.
+Each architecture was tailored to effectively process our multimodal dataset, which combines pixel data and bytecode sequences from NES games.
+
+\subsection{Training Process}
+Training was conducted on dual T4 GPUs from Kaggle, which facilitated the experimentation with these advanced architectures and our large multimodal dataset.
+The training regime employed various hyperparameters, including a batch size of 32, a learning rate of 0.001, and a total of 200 epochs, with evaluations performed every 5 epochs.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{training_results.png}
+\caption{Training and validation metrics (Loss, Accuracy, and F1 Score) for RNN, Transformer, and SSM models over 200 epochs.}
+\label{fig:training_results}
+\end{figure}
+
+Figure \ref{fig:training_results} illustrates the training and validation metrics for our three model architectures over the course of 200 epochs.
+The results reveal distinct performance patterns across the different architectures.
+The State Space Models (SSM) achieved the best training scores in terms of loss, accuracy, and F1 score.
+However, as evident from the widening gap between training and validation curves, they exhibited signs of overfitting over the epochs.
+This suggests that while SSMs quickly learned to fit the training data, they struggled to generalize to unseen examples.
+
+In contrast, the LSTMs, denoted as RNNs in our analysis, demonstrated more balanced performance.
+The closer alignment of their training and validation curves indicates a better compromise between model capacity and generalization capabilities.
+This robustness makes RNNs a strong candidate for our task of NES game data processing and generation.
+Interestingly, the Transformer architecture yielded the lowest performance metrics among the three models.
+This outcome is noteworthy and should be contextualized, as our implementation did not incorporate positional encoding, which is crucial for the effective functioning of transformer models in sequence tasks.
+The relatively poor performance of the Transformer model underscores the importance of architecture-specific optimizations in sequence modeling tasks.
+
+Overall, these results highlight the trade-offs between different architectures and the importance of careful model selection and tuning for our specific task.
+While SSMs showed impressive training performance, their tendency to overfit suggests that additional regularization techniques might be necessary for optimal performance.
+The stable performance of RNNs across both training and validation sets indicates their suitability for this particular task, balancing good fit and generalization.
+
+\subsection{Generation Evaluation Metrics}
+
+Our experiments utilize various model architectures including State Space Models (SSM), Attention models (Att), and Recurrent Neural Networks (RNN) for generating 16 consecutive sequences.
+We employ three key metrics as described by \cite{sutskever2014sequence}, \cite{hamming1950error}, and \cite{papineni2002bleu}:
+
 \begin{itemize}
-    \item Recurrent Neural Network (RNN)
-    \item Transformer
-    \item Mamba (State Space Model)
+    \item \textbf{Hamming Distance}: Scaled 0-1, lower values indicate greater similarity.
+    \item \textbf{Cosine Similarity}: Ranges -1 to 1, higher values suggest greater similarity.
+    \item \textbf{BLEU Score}: Varies 0-1, higher scores indicate better quality.
 \end{itemize}
 
-\subsection{Training Process}
-Training was conducted on dual T4 GPUs from Kaggle, enabling experimentation with these advanced architectures and our large multimodal dataset.
+\begin{table}[h]
+\centering
+\small
+\begin{tabular}{cccccccccc}
+\hline
+Type & Dir. & Emb. & Conv. & In & Mod & Depth & Hamming & Cosine & BLEU \\
+\hline
+SSM & Bi & PxBy & Y & 81 & 64 & 1 & 0.170 $\pm$ 0.086 & 0.883 $\pm$ 0.105 & 0.753 $\pm$ 0.115 \\
+SSM & Bi & PxBy & Y & 81 & 64 & 2 & 0.158 $\pm$ 0.074 & 0.896 $\pm$ 0.095 & 0.771 $\pm$ 0.097 \\
+SSM & Uni & PxBy & Y & 81 & 64 & 2 & 0.166 $\pm$ 0.081 & 0.886 $\pm$ 0.102 & 0.760 $\pm$ 0.106 \\
+Att & Bi & PxBy & Y & 81 & 64 & 1 & 0.157 $\pm$ 0.064 & 0.887 $\pm$ 0.103 & 0.765 $\pm$ 0.088 \\
+Att & Bi & PxBy & Y & 81 & 64 & 2 & 0.159 $\pm$ 0.066 & 0.887 $\pm$ 0.103 & 0.760 $\pm$ 0.092 \\
+RNN & Bi & Center & Y & 81 & 64 & 2 & 0.185 $\pm$ 0.074 & 0.888 $\pm$ 0.083 & 0.750 $\pm$ 0.093 \\
+RNN & Bi & PxBy & N & 81 & 64 & 2 & 0.153 $\pm$ 0.061 & 0.902 $\pm$ 0.090 & 0.777 $\pm$ 0.083 \\
+RNN & Bi & PxBy & Y & 162 & 64 & 2 & 0.152 $\pm$ 0.062 & 0.905 $\pm$ 0.089 & 0.778 $\pm$ 0.084 \\
+RNN & Bi & PxBy & Y & 36 & 64 & 2 & 0.152 $\pm$ 0.061 & 0.904 $\pm$ 0.090 & 0.778 $\pm$ 0.083 \\
+RNN & Bi & PxBy & Y & 81 & 128 & 2 & 0.153 $\pm$ 0.063 & 0.903 $\pm$ 0.091 & 0.776 $\pm$ 0.086 \\
+RNN & Bi & PxBy & Y & 81 & 32 & 2 & \textbf{0.149} $\pm$ 0.062 & 0.899 $\pm$ 0.095 & \textbf{0.785} $\pm$ 0.082 \\
+RNN & Bi & PxBy & Y & 81 & 64 & 1 & 0.149 $\pm$ 0.062 & 0.897 $\pm$ 0.096 & 0.780 $\pm$ 0.085 \\
+RNN & Bi & PxBy & Y & 81 & 64 & 3 & 0.153 $\pm$ 0.063 & \textbf{0.906} $\pm$ 0.087 & 0.776 $\pm$ 0.086 \\
+RNN & Bi & PxBy & N & 81 & 64 & 2 & 0.151 $\pm$ 0.062 & 0.903 $\pm$ 0.090 & 0.779 $\pm$ 0.084 \\
+RNN & Uni & PxBy & Y & 81 & 64 & 2 & 0.153 $\pm$ 0.064 & 0.904 $\pm$ 0.088 & 0.777 $\pm$ 0.087 \\
+\hline
+\end{tabular}
+\caption{Comparison of model characteristics and performance (mean $\pm$ std)}
+\label{tab:model_comparison}
+\end{table}
 
-\subsection{Evaluation Metrics}
-We employed various metrics to evaluate the performance of our models, including perplexity, BLEU score for text generation, and structural similarity index (SSIM) for image generation.
+The results reveal interesting patterns across different model architectures.
+SSM models show competitive performance, with the bidirectional 2-layer model achieving better results than its unidirectional counterpart.
+Attention models demonstrate consistent performance across depths, with slight improvements in the 1-layer model.
+RNN models, particularly those with PxBy embedding, consistently outperform other architectures.
+The best Hamming distance (0.149 $\pm$ 0.062) and BLEU score (0.785 $\pm$ 0.082) are achieved by the bidirectional PxBy RNN model with convolution, 81 input dimensions, and 32 model dimensions.
+This suggests that a more compact RNN model can effectively capture sequence patterns, aligning with findings on model compression \cite{han2016deep}.
+
+The highest cosine similarity (0.906 $\pm$ 0.087) is obtained by the 3-layer RNN model, indicating that increased depth can enhance sequence alignment \cite{pascanu2014construct}.
+Convolution generally improves performance in RNN models, supporting \cite{lecun1995convolutional}'s findings on local pattern capture.
+Interestingly, varying the input dimension (36, 81, 162) in RNN models had minimal impact on performance, suggesting robustness to input size changes.
+The center embedding in RNNs showed lower performance compared to PxBy embedding, highlighting the effectiveness of the latter for this task.
 
 \section{Results and Discussion}
-[This section will be filled with findings and observations from our experiments, including:
-- Performance comparison of RNN, Transformer, and Mamba models
-- Ablation studies to understand the importance of each component
-- Qualitative analysis of generated embeddings and outputs
-- Methods and results of measuring similarity in generated sequences]
+
+Our experiments with various model architectures and configurations yielded significant findings for the task of NES game data processing and generation.
+We evaluated the performance of the models based on several metrics, including Hamming distance, cosine similarity, and BLEU score \cite{papineni2002bleu}.
+
+\subsection{Model Performance}
+
+The bidirectional RNN models, particularly those utilizing PixelBytes (PxBy) embedding with convolutional layers, consistently outperformed other architectures across our evaluation metrics.
+Key observations include:
+
+\begin{itemize}
+    \item \textbf{Embedding Strategy}: The PxBy embedding strategy significantly enhanced model performance.
+The mean BLEU score improved from 0.750 for the center-only approach to 0.777 for the PxBy embedding, demonstrating the importance of utilizing richer representations of the input data.
+
+    \item \textbf{Model Architecture}: Bidirectional models outperformed their unidirectional counterparts.
+The bidirectional RNN achieved a BLEU score of 0.777, compared to 0.776 for the unidirectional model, suggesting that bidirectionality is crucial for capturing the dependencies in sequential data.
+
+    \item \textbf{Model Configuration}: Variations in state dimension and layer count had minimal impact on performance.
+Models with state dimensions of 32, 64, and 128 achieved BLEU scores of 0.784, 0.777, and 0.775 respectively, indicating that our base configuration was sufficiently expressive for the task.
+\end{itemize}
+
+\subsection{Comparative Analysis}
+
+The superior performance of bidirectional PxBy RNN models with convolution for our sequence modeling task is evident.
+However, the potential of State Space Models (SSM) and Attention models in certain scenarios cannot be overlooked.
+The SSM models, in particular, exhibited rapid convergence and strong training performance, suggesting their viability for future exploration in multimodal data processing.
+
+These findings contribute to the growing body of research on multimodal sequence modeling \cite{baltrusaitis2019multimodal} and highlight the importance of architecture selection and embedding strategies in processing complex, multimodal data such as video game content.
 
 \section{Conclusion and Future Directions}
-This report on PixelBytes opens new avenues in unified multimodal generation.
-Our findings suggest promising potential for coherent text-image generation.
-Future work will focus on refining the PxByEmbed algorithm, exploring larger datasets, and investigating applications in creative content generation.
+
+This study on PixelBytes opens new avenues in unified multimodal generation, demonstrating promising potential for coherent text-image generation in the context of NES game data.
+The superior performance of bidirectional RNN models with PxBy embedding underscores the importance of comprehensive spatial information and context in processing game data.
+Future work will focus on:
+
+\begin{itemize}
+    \item Refining the PxByEmbed algorithm to further improve its effectiveness in capturing spatial and temporal relationships in game data.
+    \item Exploring larger and more diverse datasets to enhance the generalizability of our models.
+    \item Investigating applications in creative content generation, such as level design or narrative creation for retro-style games.
+    \item Further optimizing SSM and Attention models to better compete with RNN architectures in this specific domain.
+\end{itemize}
+
+These directions aim to build upon our current findings and push the boundaries of multimodal sequence modeling in the context of video game content generation.
 

### 2024-08-30 pixelbytes/displays.py

@@ -1,21 +1,75 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
 """
-@author: fabien
+@author: fabienfrfr
 """
 
-import numpy as np
-import matplotlib.pyplot as plt
 import warnings
 warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
 
 import numpy as np
 import matplotlib.pyplot as plt
-from matplotlib.patches import Rectangle
+import matplotlib.patches as patches
+from matplotlib.offsetbox import OffsetImage, AnnotationBbox
+import matplotlib.gridspec as gridspec
+from .generator import SequenceGenerator
+import torch
 
-def reconstruct_sequence(tokens_arr_obj, min_row_length=3, max_gap=50, max_image_gap=100):
-    def reconstruct_imgs(img_tokens, start_index):
-        img_idx = np.arange(len(img_tokens)) + start_index
+class Displays:
+    def __init__(self, tokenizer):
+        self.tokenizer = tokenizer
+        self.sequence_generator = SequenceGenerator(self.tokenizer)
+        self._images = None #previous img
+        self._texts = None #previous text
+        self.fig = None
+    
+    def reset(self, model: torch.nn.Module):
+        self.model = model.eval()
+    
+    # specific process
+    def generate_sequence(self, input_seq, total_length, input_window=None, gen_window=None):
+        N, M, _ = input_seq.shape
+        self.sequence_generator.reset(input_seq)
+        is_progressive = input_window is not None and gen_window is not None
+        while len(self.sequence_generator.sequence) < total_length:
+            if is_progressive:
+                input_tensor = torch.from_numpy(self.sequence_generator.sequence[-input_window:]).long()
+                gen_length = min(gen_window, total_length - len(self.sequence_generator.sequence))
+            else:
+                input_tensor = torch.from_numpy(self.sequence_generator.sequence).long()
+                gen_length = total_length - len(self.sequence_generator.sequence)
+            with torch.no_grad():
+                for _ in range(gen_length):
+                    output = self.model(input_tensor.unsqueeze(0))
+                    _, predicted = output.max(1)
+                    next_token_id = predicted.squeeze().cpu().numpy().item()
+                    self._update_sequence(next_token_id)
+                    if is_progressive:
+                        input_tensor = torch.from_numpy(self.sequence_generator.sequence[-input_window:]).long()
+                    else:
[…]

### 2024-08-30 pixelbytes/results.py

@@ -1,140 +1,89 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
 """
-@author: fabien
+@author: fabienfr
 """
 
-import seaborn as sns
-import matplotlib.pyplot as plt
-import pandas as pd
 import os
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
 
-# Configuration du style
-plt.style.use('default')  # Utilisation du style par défaut de Matplotlib
-sns.set_theme()  # Application des paramètres de base de Seaborn
-plt.rcParams.update({
-    'font.family': 'serif',
-    'font.serif': ['Times New Roman'] + plt.rcParams['font.serif'],
-    'font.size': 10,
-    'axes.labelsize': 12,
-    'axes.titlesize': 14,
-    'xtick.labelsize': 10,
-    'ytick.labelsize': 10,
-    'legend.fontsize': 10,
-    'figure.titlesize': 16,
-    'axes.grid': True,
-    'grid.alpha': 0.3,
-    'axes.axisbelow': True,
-    'axes.facecolor': '#EAEAF2',
-    'axes.edgecolor': 'none',
-    'axes.linewidth': 0,
-    'figure.facecolor': 'white'
-})
-
-
-def load_csv_files(directory_or_file):
-    data_dict = {}
-    if os.path.isdir(directory_or_file):
-        for file in os.listdir(directory_or_file):
-            if file.endswith('.csv'):
-                model_name = os.path.splitext(file)[0]
-                file_path = os.path.join(directory_or_file, file)
-                data_dict[model_name] = pd.read_csv(file_path)
-    elif os.path.isfile(directory_or_file) and directory_or_file.endswith('.csv'):
-        df = pd.read_csv(directory_or_file)
-        if 'model' in df.columns:
-            for model in df['model'].unique():
-                data_dict[model] = df[df['model'] == model].reset_index(drop=True)
-        else:
-            data_dict['unnamed_model'] = df
+def load_csv_files(directory, separators=['rnn', 'ssm', 'attention']):
+    data_dict = {sep: [] for sep in separators}
+    for file in os.listdir(directory):
+        if file.endswith('.csv'):
+            model_name = os.path.splitext(file)[0]
+            file_path = os.path.join(directory, file)
+            # Determine the model group
+            group = next((sep for sep in separators if sep in model_name.lower()), 'other')
+            # Load the data
+            df = pd.read_csv(file_path)
[…]

### 2024-08-30 tests.py

@@ -8,13 +8,39 @@ TESTS FILE
 from pixelbytes import *
 from datasets import load_dataset
 
-
 ### basic test
 if __name__ == '__main__' :
+
+    #quit()
     # common part
     hf_dataset = load_dataset("ffurfaro/PixelBytes-Pokemon")
+    tokenizer = PixelBytesTokenizer()
+    # reconstruct bulbizarre
+    bulbi = hf_dataset["train"][0]
+    import cv2, numpy as np, pylab as plt
+    Image = np.array(bulbi['image'])
+    #cv2.imwrite('Bulbasaur.png', cv2.cvtColor(Image, cv2.COLOR_BGR2RGB), [cv2.IMWRITE_PNG_COMPRESSION, 0])
+    # init
+    displayer = Displays(tokenizer)
+    model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    displayer.reset(model)
+    #model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="ssm_bi_pxby_conv_81_dim_64_state_2_layer_last")
+    first_seq = np.array(hf_dataset["train"]['pixelbyte'][0])
+    N,_,_ = first_seq.shape
+    # set generation --> objective : loop of put 32 seq, gen 8 --> jusqu'a ce que ca soit à N
+    L = int(N//4)
+    input_seq = first_seq[:L]
+    #images, text = displayer.process_sequence(input_seq, N)
+    images, text = displayer.process_and_display(input_seq, N, input_window=32, gen_window=1) # find minimal size for not perturb gen --> for good display
 
-
+    """
+    # plot
+    train_results_path = 'models/train_results'
+    separators = ['rnn', 'ssm', 'attention']
+    train_dfs = load_csv_files(train_results_path,  separators)
+    plot_training_results(train_dfs, 'data')
+    """
+    """
     # evaluation part
     #config = EvaluateMetricConfig(output_dir="models/eval_results")
     data_columns = hf_dataset["train"]['pixelbyte']
@@ -27,9 +53,7 @@ if __name__ == '__main__' :
     evaluator.reset(model)
     results = evaluator.evaluate(model_name)
     print(results.describe())
-
-
-
+    """
     """    
     # generation test
     tokenizer = PixelBytesTokenizer()
@@ -47,7 +71,6 @@ if __name__ == '__main__' :
[…]

### 2024-08-31 docs/paper/article.tex
dummy redaction

--- 2024-08-27 docs/paper/article.tex
@@
 \begin{abstract}
-This report presents PixelBytes, an innovative multimodal embedding model designed for the simultaneous generation of text and images pixel by pixel.
-We introduce the PxByEmbed algorithm as a key component of our architecture, enabling efficient representation of mixed sequences of text and images.
-Our approach leverages web scraping, image processing, and advanced machine learning techniques to create a unique dataset and training pipeline.
-We compare the performance of RNN, Transformer, and Mamba models in this task, providing insights into the effectiveness of different architectures for multimodal generation.
+This report introduces PixelBytes Embedding, a novel approach for unified multimodal representation learning.
+Our method captures diverse inputs in a single, cohesive representation, enabling emergent properties for multimodal sequence generation, particularly for text and pixelated images.
+Inspired by state-of-the-art sequence models such as Image Transformers, PixelCNN, and Mamba-Bytes, PixelBytes aims to address the challenges of integrating different data types.
+We explore various model architectures, including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on bidirectional processing and our innovative PxBy embedding technique.
+Our experiments, conducted on a specialized PixelBytes Pokémon dataset, demonstrate that bidirectional sequence models with PxBy embedding and convolutional layers can generate coherent multimodal sequences.
+This work contributes to the advancement of integrated AI models capable of understanding and generating multimodal data in a unified manner.
+Code is available at \url{https://github.com/fabienfrfr}.
 \end{abstract}
+
+\let\thefootnote\relax\footnotetext{$^{*}$ Corresponding author: Fabien Furfaro (\url{fabien.furfaro@gmail.com}).}
 
@@
 \includegraphics[width=0.8\linewidth]{example_generation.png}
-\caption{Example of a generated pixelated Pokémon and its description}
+\caption{Example of a PixelBytes-generated Pokémon and its description.
+Left: Generated pixelated image.
+Right: Corresponding text description.
+The red box indicates the model's attention window for both pixels and words.
+Note that the model currently exhibits some inaccuracies in image generation and occasionally invents words, highlighting areas for future improvement.}
 \label{fig:example_generation}
@@
 \section{Introduction}
-Multimodal generation, which combines text and images, represents a significant challenge in artificial intelligence.
-Recent advancements in pixel-by-pixel image generation have shown promising results in creating coherent visual content \cite{parmar2018image}.
-Plus connu egalmeent les modele de generation de texte comme GPT.
+
+Recent advancements in artificial intelligence have led to increasingly generalist models, not by combining multiple specialized components, but by giving simple tasks to models where emergent properties—complex behaviors that arise from simpler underlying rules—appear.
+This is the case with generative language models like GPT \cite{brown2020language}.
+However, these models are limited by their focus on language alone, failing to capture the full complexity of multimodal understanding \cite{alayrac2023language}.
+To address this limitation, researchers have explored combining LLMs with other modalities \cite{radford2021learning, ramesh2022hierarchical}.
+But this brings us back to the initial problem, as it often results in specialized model combinations without allowing for new emergent properties.
+We propose "PixelBytes Embedding", a novel approach enabling unified training across modalities by capturing diverse inputs in a single, cohesive representation.
+
+Multimodal sequence generation, which involves the creation of coherent outputs combining various data types such as text, images, and numerical sequences, presents a significant challenge in artificial intelligence \cite{baltrusaitis2019multimodal}.
+While models like GPT have excelled in text generation \cite{brown2020language}, there's a growing need for unified approaches that can seamlessly handle varied data types.
 Building upon these findings, PixelBytes aims to address the challenge of unified text and image generation by proposing a model capable of producing mixed sequences of text and images in a coherent and unified manner.
+We draw inspiration from state-of-the-art sequence models, including Image Transformers \cite{parmar2018image}, PixelCNN \cite{oord2016conditional}, and recent developments in bytes generation by mamba architectures \cite{wang2024mamba}.
 
-Our approach draws inspiration from the state-of-the-art in image transformers, which have demonstrated the ability to generate high-quality images with fine-grained control \cite{parmar2018image}.
-We also build upon generative sequence models such as GPT, PixelRNN, and PixelCNN.
-Additionally, we leverage recent developments in bidirectional state space models for time series prediction, which have shown improved performance in capturing long-range dependencies \cite{bimamba}.
-We also incorporate insights from tokenless selective state space models, which have proven effective in processing sequential data efficiently \cite{mambabyte}.
+Our research explores various architectures including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on the effectiveness of bidirectional processing \cite{gu2023visual, gu2023bidirectional}, novel embedding techniques (particularly PxBy embedding, which unifies pixel and byte-level representations), the impact of convolutional layers, effects of model depth, input dimensionality, and size.
+Our experiments reveal that bidirectional sequence models with PxBy embedding and convolutional layers demonstrate good performance in generating coherent sequences.
 
-Notre approche a permis de montrer (...)
+Our results show that it is possible to train sequence models with unified embeddings for both text and pixel completion.
+This paper presents our methodology to construct unified sequence data from text and image, experimental results, and analysis, contributing to the development of our sequence generation embedding models.
+The proposed PixelBytes Embedding approach has the potential to significantly advance the field of multimodal AI, enabling more integrated and versatile models capable of understanding and generating diverse types of data.
+This could lead to improvements in areas such as visual question answering, image captioning, and multimodal content generation for virtual and augmented reality applications.
 
@@
 \subsection{Overview}
-PixelBytes integrates several innovative components to achieve its multimodal generation capabilities:
-\begin{itemize}
-    \item A tokenizer sequence constructor with pixelizable image, and bytes texte
-    \item A unified multimodal embedding (PxByEmbed)
-\end{itemize}
+The PixelBytes architecture is designed to seamlessly integrate multimodal data for unified sequence generation.
+At its core, the model incorporates two key innovative components: a specialized tokenizer sequence constructor and a unified multimodal embedding technique called PxByEmbed.
+The tokenizer sequence constructor is engineered to process both pixelated images and text at a byte level, enabling a consistent representation across modalities.
+This is complemented by PxByEmbed, our novel embedding approach that creates a unified representation for both pixel and byte data in a single, coherent space.
+This architecture draws inspiration from recent advancements in multimodal learning \cite{baltrusaitis2019multimodal} and efficient sequence modeling \cite{vaswani2017attention}, allowing PixelBytes to capture intrinsic relationships between visual and textual information effectively.
 
@@
 
-Pour tester notre approche, nous avons besoin d'une base de donnée d'image-captionning.
-Néanmoins, les bases de données actuelles ne sont pas adapté pour combiner la génération de texte et d'image.
-Les textes sont souvent trop petit, et une image pixelisé des base de données d'image captionning comme (COCO, Flick) sont difficilement interpretable (les images d'origine sont trop volumineuse).
-C'est pourquoi, nous avons choisi de construire notre base de donnée à partir des données wiki de Pokemon.
-Pokemon a l'avantage d'exister depuis plus de 20 ans, lors de l'apparition des jeux video portable, il y a encore des dessins pixelisé ou pixelisable de pokemon.
-de plus, leur description est une bonne base pour decrire le pokemon.
-Aujourd'hui, il existe plus de 1000 pokemon, ce qui permet d'avoir une base de donnée minimal pour construire un jeu de données specialisé dans la generation de description et de pixel pokemon.
+Image captioning datasets, while combining visual and textual modalities, prove unsuitable for joint text and image generation due to limited text content and difficulties in interpreting pixelated versions of high-resolution images.
+To address these limitations, we developed a specialized Pokémon dataset, offering advantages such as a long-standing franchise history, pixelated designs, rich descriptive text, and over 1000 unique Pokémon.
+This aligns with trends in creating specialized datasets for multimodal learning \cite{lin2014microsoft}.
 
-\subsubsection{Web Scraping from Pokepedia}
-We constructed our dataset by scraping miniatures and descriptions from Pokepedia, ensuring a rich and diverse collection of Pokémon data.
-
-Utilisation de beautifull soup.
-
-\subsubsection{Image Pixelation and Text-Image Balance}
-The scraped images were pixelated to create a retro aesthetic. We carefully balanced the dataset to maintain a 2/3 text to 1/3 image ratio, ensuring comprehensive training on both modalities.
-
-Utilisation d'OpenCV et skimage.
-
-\subsubsection{Image Quantization and Token Creation}
-We quantized the pixelated images using a palette inspired by the NES color scheme (55 colors).
-This process allowed us to create tokens representing different color and position combinations, effectively translating visual information into a format suitable for sequence modeling.
-
-Utilisation d'OpenCV et skimage. (NES palette via le site web des palettes)
-
-Construction des token d'image
-
-Enregistrement dans le Hub dataset de HuggingFace
+We constructed our dataset by web scraping Pokémon miniatures and descriptions from Pokepedia using Beautiful Soup \cite{richardson2007beautiful}, maintaining a 2/3 text to 1/3 image ratio.
+For image quantization, we employed a 55-color palette inspired by the NES, creating tokens representing different color combinations.
+This approach translates visual information into a format suitable for sequence modeling, similar to recent work in discrete representation learning for images \cite{van2017neural}.
+The quantization and pixelization process used OpenCV and scikit-image \cite{bradski2000opencv, vanderwalt2014scikit}.
+To ensure balanced representation, we adjusted the number of image and text tokens to 113 characters for each modality, allowing the model to learn equally from both visual and textual information.
+The final dataset, balancing text and pixelated images, is available on the Hugging Face Datasets Hub for reproducibility at \url{https://huggingface.co/datasets/ffurfaro/PixelBytes-Pokemon}.
 
@@
 \subsection{PxByEmbed: Multimodal Embedding Algorithm}
-At the core of our approach is the PxByEmbed algorithm, which represents mixed sequences of text and images in a unified manner.
-This algorithm extends classical embedding techniques by incorporating spatial adaptivity, allowing for more effective representation of both textual and visual information.
+
+At the core of our approach is the PxByEmbed algorithm, which represents mixed sequences of text and pixelated images in a unified manner.
+This algorithm builds upon existing embedding techniques by incorporating spatial adaptivity, allowing for representation of both textual and visual information.
+PxByEmbed is designed to address the specific needs of pixel-level image representation and byte-level text encoding.
+The algorithm uses a learned embedding matrix to map each token (text or image) to a vector space, while maintaining spatial relationships for image tokens.
+This approach allows our model to handle both modalities within a single framework.
 
@@
 \subsection{Managing Transitions}
-PixelBytes employs newline characters (ASCII 0A) to manage transitions between text and image, ensuring coherence in the generation of mixed sequences.
 
-\section{Model Variants and Training}
+PixelBytes uses a method to handle transitions between text and image tokens during sequence generation.
+This approach is used in both dataset construction and sequence generation (Figure \ref{fig:example_generation}).
+
+For dataset construction, we use a 2D input sequence method with a 3x3 context window around each token.
+We use special tokens to mark transitions between text and images, and add padding to keep context sizes consistent.
+For text, only previous tokens are included in the context windows.
+The \texttt{input\_seq\_construct} function implements this process.
+
+During sequence generation, the \texttt{\_process\_token} function manages transitions.
+It handles text tokens (bytes) and image tokens (tuples) differently, and takes care of special characters like newlines and tabs.
+A \texttt{sequence\_clock} keeps track of the generation progress.
+Each new token's context is represented in a 3x3 matrix, which helps maintain coherence when switching between text and images.
+
+\section{Training and Evaluation}
 
 \subsection{Model Architectures}
-We implemented and compared three different model architectures:
-\begin{itemize}
-    \item Recurrent Neural Network (RNN)
-    \item Transformer
-    \item Mamba (State Space Model)
-\end{itemize}
+We tested three different model types: Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{hochreiter1997long}, Transformer \cite{vaswani2017attention}, and Mamba, which is based on the State Space Model (SSM) \cite{gu2022efficiently}.
+Each model was adapted to work with our dataset of pixel data and bytecode sequences.
+We trained the models on Kaggle using dual T4 GPUs.
+We used a batch size of 32, sequence length of 256, learning rate of 0.001, and trained for 200 epochs, checking the model's performance every 5 epochs.
+The trained models are available at \url{https://huggingface.co/ffurfaro/PixelBytes-Pokemon}.
 
-\subsection{Training Process}
-Training was conducted on dual T4 GPUs from Kaggle, enabling experimentation with these advanced architectures and our large multimodal dataset.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{training_results.png}
+\caption{Training and validation metrics (Loss, Accuracy) for RNN, Transformer, and SSM models over 200 epochs.}
+\label{fig:training_results}
+\end{figure}
 
-\subsection{Evaluation Metrics}
-We employed various metrics to evaluate the performance of our models, including perplexity, BLEU score for text generation, and structural similarity index (SSIM) for image generation.
+Figure \ref{fig:training_results} shows the training and validation metrics for our three model types over 200 epochs.
+The State Space Models (SSM) achieved the best scores for loss and accuracy.
+However, the widening gap between their training and validation curves suggests they may be overfitting, meaning they learned the training data well but struggled to generalize to new examples.
+
+In contrast, the LSTMs (referred to as RNNs in our analysis) demonstrated more balanced performance.
+The closer alignment of their training and validation curves indicates they may generalize better to unseen data.
+The Transformer model had the lowest performance among the three.
+This could be due to the absence of positional encoding, which is important for Transformers in sequence tasks.
+
+Overall, these results highlight the strengths and weaknesses of each model type.
+While SSMs showed strong training performance, their tendency to overfit suggests that additional regularization techniques might be needed.
+The stable performance of RNNs indicates their suitability for our task, as they balance well between fitting the training data and generalizing to new data.
+The Transformer's performance suggests it may require further adjustments to improve its effectiveness for this task.
+
+\subsection{Generation Evaluation Metrics}
+
+We tested different model types including State Space Models (SSM), Attention models (Att), and Recurrent Neural Networks (RNN) for generating 16 consecutive sequences.
+We used three metrics to evaluate their performance: Hamming Distance, Cosine Similarity, and BLEU Score \cite{sutskever2014sequence, hamming1950error, papineni2002bleu}.
+
+\begin{table}[h]
+\centering
+\small
+\begin{tabular}{cccccccccc}
+\hline
+Type & Dir. & Emb. & Conv. & In Emb & Hidden State & Depth & Hamming & Cosine & BLEU \\
+\hline
+SSM & Bi & PxBy & Y & 81 & 64 & 1 & 0.170 $\pm$ 0.086 & 0.883 $\pm$ 0.105 & 0.753 $\pm$ 0.115 \\
+SSM & Bi & PxBy & Y & 81 & 64 & 2 & 0.158 $\pm$ 0.074 & 0.896 $\pm$ 0.095 & 0.771 $\pm$ 0.097 \\
+SSM & Uni & PxBy & Y & 81 & 64 & 2 & 0.166 $\pm$ 0.081 & 0.886 $\pm$ 0.102 & 0.760 $\pm$ 0.106 \\
+Att & - & PxBy & Y & 81 & 64 & 1 & 0.157 $\pm$ 0.064 & 0.887 $\pm$ 0.103 & 0.765 $\pm$ 0.088 \\
+Att & - & PxBy & Y & 81 & 64 & 2 & 0.159 $\pm$ 0.066 & 0.887 $\pm$ 0.103 & 0.760 $\pm$ 0.092 \\
+RNN & Bi & Center & N & 81 & 64 & 2 & 0.185 $\pm$ 0.074 & 0.888 $\pm$ 0.083 & 0.750 $\pm$ 0.093 \\
+RNN & Bi & PxBy & Y & 81 & 64 & 2 & 0.153 $\pm$ 0.061 & 0.902 $\pm$ 0.090 & 0.777 $\pm$ 0.083 \\
+RNN & Bi & PxBy & Y & 162 & 64 & 2 & 0.152 $\pm$ 0.062 & 0.905 $\pm$ 0.089 & 0.778 $\pm$ 0.084 \\
+RNN & Bi & PxBy & Y & 36 & 64 & 2 & 0.152 $\pm$ 0.061 & 0.904 $\pm$ 0.090 & 0.778 $\pm$ 0.083 \\
+RNN & Bi & PxBy & Y & 81 & 128 & 2 & 0.153 $\pm$ 0.063 & 0.903 $\pm$ 0.091 & 0.776 $\pm$ 0.086 \\
+RNN & Bi & PxBy & Y & 81 & 32 & 2 & \textbf{0.149} $\pm$ 0.062 & 0.899 $\pm$ 0.095 & \textbf{0.785} $\pm$ 0.082 \\
+RNN & Bi & PxBy & Y & 81 & 64 & 1 & 0.149 $\pm$ 0.062 & 0.897 $\pm$ 0.096 & 0.780 $\pm$ 0.085 \\
+RNN & Bi & PxBy & Y & 81 & 64 & 3 & 0.153 $\pm$ 0.063 & \textbf{0.906} $\pm$ 0.087 & 0.776 $\pm$ 0.086 \\
+RNN & Bi & PxBy & N & 81 & 64 & 2 & 0.151 $\pm$ 0.062 & 0.903 $\pm$ 0.090 & 0.779 $\pm$ 0.084 \\
+RNN & Uni & PxBy & Y & 81 & 64 & 2 & 0.153 $\pm$ 0.064 & 0.904 $\pm$ 0.088 & 0.777 $\pm$ 0.087 \\
+\hline
+\end{tabular}
+\caption{Comparison of model characteristics and performance (mean $\pm$ std)}
+\label{tab:model_comparison}
+\end{table}
+
+The results show that RNN models with PxBy embedding generally perform better than other types.
+The best Hamming distance (0.149 $\pm$ 0.062) and BLEU score (0.785 $\pm$ 0.082) come from the bidirectional PxBy RNN model with convolution, 81 embedding dimensions, and 32 hidden state dimensions.
+This suggests that a smaller hidden state in the RNN model can effectively capture sequence patterns.
+
+The highest cosine similarity (0.906 $\pm$ 0.087) is from the 3-layer RNN model, indicating that more layers can improve sequence alignment.
+Adding convolution usually improves RNN model performance.
+Interestingly, changing the embedding dimension (36, 81, 162) in RNN models didn't affect performance much, suggesting these models can handle different embedding sizes well.
+The center embedding in RNNs didn't perform as well as PxBy embedding for this task.
 
 \section{Results and Discussion}
-[This section will be filled with findings and observations from our experiments, including:
-- Performance comparison of RNN, Transformer, and Mamba models
-- Ablation studies to understand the importance of each component
-- Qualitative analysis of generated embeddings and outputs
-- Methods and results of measuring similarity in generated sequences]
 
-\section{Conclusion and Future Directions}
-This report on PixelBytes opens new avenues in unified multimodal generation.
-Our findings suggest promising potential for coherent text-image generation.
-Future work will focus on refining the PxByEmbed algorithm, exploring larger datasets, and investigating applications in creative content generation.
+Our experiments with various model architectures tested the effectiveness of unified text and image generation.
+We evaluated model performance using Hamming distance, cosine similarity, and BLEU score \cite{papineni2002bleu}.
+Bidirectional RNN models using PixelBytes (PxBy) embedding with convolutional layers generally performed better across our metrics.
+The PxBy embedding improved the mean BLEU score from 0.750 to 0.777 compared to a center-only approach, indicating its effectiveness in representing multimodal data.
+This finding aligns with recent work emphasizing the importance of effective multimodal embeddings in text-to-image generation tasks \cite{yan2024multimodal}.
+
+Bidirectional models showed a slight advantage over unidirectional ones, with BLEU scores of 0.777 and 0.776 respectively.
+This suggests that bidirectionality may help capture sequential dependencies, consistent with other studies in sequence modeling \cite{schuster1997bidirectional}.
+Notably, varying state dimensions (32, 64, 128) had minimal impact on performance, with BLEU scores ranging from 0.775 to 0.784.
+While RNN models showed strong performance in our experiments, State Space Models (SSM) demonstrated rapid convergence during training, suggesting potential for future exploration in multimodal data processing.
+This is in line with recent advancements in SSM for sequence modeling tasks \cite{gu2022efficiently}.
+
+\section{Conclusion and Future Work}
+
+This study on PixelBytes highlights the potential for unified text-image generation, contributing to the growing field of multimodal AI \cite{baltrusaitis2019multimodal}.
+The performance of bidirectional RNN models with PxBy embedding underscores the importance of capturing both spatial and contextual information in multimodal data.
+Future work could focus on refining the PxByEmbed algorithm, testing on larger and more diverse datasets, exploring creative applications of multimodal generation, and further optimizing SSM and Attention models.
+These efforts aim to advance multimodal sequence modeling and generation techniques, addressing challenges identified in surveys of the field \cite{lin2014microsoft}.
 

### 2024-08-31 pixelbytes/displays.py

@@ -27,39 +27,31 @@ class Displays:
         self.model = model.eval()
     
     # specific process
-    def generate_sequence(self, input_seq, total_length, input_window=None, gen_window=None):
-        N, M, _ = input_seq.shape
+    def generate_sequence(self, complete_sequence, start_length=64, window=None, gen_window=None):
+        L, _, _ = complete_sequence.shape
+        input_seq = complete_sequence[:start_length]
         self.sequence_generator.reset(input_seq)
-        is_progressive = input_window is not None and gen_window is not None
-        while len(self.sequence_generator.sequence) < total_length:
+        is_progressive = window is not None and gen_window is not None
+        while start_length < L :
+            input_tensor = torch.from_numpy(self.sequence_generator.sequence).long()
             if is_progressive:
-                input_tensor = torch.from_numpy(self.sequence_generator.sequence[-input_window:]).long()
-                gen_length = min(gen_window, total_length - len(self.sequence_generator.sequence))
-            else:
-                input_tensor = torch.from_numpy(self.sequence_generator.sequence).long()
-                gen_length = total_length - len(self.sequence_generator.sequence)
+                update_length = min(window, L - len(self.sequence_generator.sequence) - gen_window)
+                update_token_id = complete_sequence[start_length:start_length+update_length][:,1,1]
+                new_sequence = self.sequence_generator.update_sequence(update_token_id)
+                gen_length = min(gen_window, L - len(self.sequence_generator.sequence))
+            else: gen_length = L - len(self.sequence_generator.sequence)
             with torch.no_grad():
                 for _ in range(gen_length):
                     output = self.model(input_tensor.unsqueeze(0))
                     _, predicted = output.max(1)
                     next_token_id = predicted.squeeze().cpu().numpy().item()
-                    self._update_sequence(next_token_id)
-                    if is_progressive:
-                        input_tensor = torch.from_numpy(self.sequence_generator.sequence[-input_window:]).long()
-                    else:
-                        input_tensor = torch.from_numpy(self.sequence_generator.sequence).long()
-        return self.sequence_generator.sequence[:total_length]
+                    new_sequence = self.sequence_generator.update_sequence(next_token_id)
+                    input_tensor = torch.from_numpy(new_sequence).long()
+            start_length = len(self.sequence_generator.sequence)
+        return self.sequence_generator.sequence
 
[…]

### 2024-08-31 pixelbytes/model.py

@@ -14,7 +14,10 @@ try :
     from mamba_ssm.modules.mamba_simple import Mamba
 except :
     print('No mamba_ssm CUDA modules installed...')
-    from mambapy.mamba import Mamba
+    import mambapy.mamba as mambapy
+    def Mamba(d_model, d_state, d_conv, expand) :
+        config = mambapy.MambaConfig(d_model, d_state, d_conv, expand)
+        return mambapy.Mamba(config) ## doesnt works.. need to change prefix and state_dict..
 
 ### Multimodal embedding        
 class PxByEmbed(nn.Module):

### 2024-08-31 tests.py

@@ -22,17 +22,19 @@ if __name__ == '__main__' :
     #cv2.imwrite('Bulbasaur.png', cv2.cvtColor(Image, cv2.COLOR_BGR2RGB), [cv2.IMWRITE_PNG_COMPRESSION, 0])
     # init
     displayer = Displays(tokenizer)
+    #model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_conv_81_dim_32_state_2_layer_last")
     model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
     displayer.reset(model)
     #model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="ssm_bi_pxby_conv_81_dim_64_state_2_layer_last")
-    first_seq = np.array(hf_dataset["train"]['pixelbyte'][0])
-    N,_,_ = first_seq.shape
+    complete_seq = np.array(hf_dataset["train"]['pixelbyte'][0])
+    N,_,_ = complete_seq.shape
     # set generation --> objective : loop of put 32 seq, gen 8 --> jusqu'a ce que ca soit à N
     L = int(N//4)
-    input_seq = first_seq[:L]
-    #images, text = displayer.process_sequence(input_seq, N)
-    images, text = displayer.process_and_display(input_seq, N, input_window=32, gen_window=1) # find minimal size for not perturb gen --> for good display
-
+    input_seq = complete_seq[:L]
+    #images, text = displayer.process_sequence(complete_seq, window=32, gen_window=4)
+    #images, text = displayer.process_sequence(complete_seq, window=64, gen_window=4)
+    images, text = displayer.process_sequence(complete_seq, window=32, gen_window=8)
+    displayer.show(images, text)
     """
     # plot
     train_results_path = 'models/train_results'

### 2024-09-01 docs/paper/article.tex
train an optimal model for generation - Stride 1 and 1000 Epoch

@@
 \subsection{Model Architectures}
-We tested three different model types: Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{hochreiter1997long}, Transformer \cite{vaswani2017attention}, and Mamba, which is based on the State Space Model (SSM) \cite{gu2022efficiently}.
-Each model was adapted to work with our dataset of pixel data and bytecode sequences.
-We trained the models on Kaggle using dual T4 GPUs.
-We used a batch size of 32, sequence length of 256, learning rate of 0.001, and trained for 200 epochs, checking the model's performance every 5 epochs.
-The trained models are available at \url{https://huggingface.co/ffurfaro/PixelBytes-Pokemon}.
+We evaluated three compact model architectures, each with fewer than 100,000 parameters: a Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{hochreiter1997long}, a Transformer \cite{vaswani2017attention}, and Mamba, based on the State Space Model (SSM) \cite{gu2022efficiently}.
+These models were adapted to process our dataset of pixel data and bytecode sequences.
+Training was conducted on Kaggle using dual T4 GPUs, with a batch size of 32, sequence length of 256, stride size of 32, and learning rate of 0.001.
+We trained for 200 epochs, evaluating performance every 5 epochs.
+The resulting models are available at \url{https://huggingface.co/ffurfaro/PixelBytes-Pokemon}.
+Additionally, we developed a more specialized model for generation tasks, which is not presented in this paper.
 
@@
 The Transformer model had the lowest performance among the three.
-This could be due to the absence of positional encoding, which is important for Transformers in sequence tasks.
+This could be due to the absence of positional encoding trick, which is important for Transformers in sequence tasks.
 

### 2024-09-01 tests.py

@@ -24,6 +24,8 @@ if __name__ == '__main__' :
     displayer = Displays(tokenizer)
     #model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_conv_81_dim_32_state_2_layer_last")
     model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
+    total_parameters = pytorch_total_params = sum(p.numel() for p in model.parameters())
+    print(f"Nombre de paramètres : {total_parameters:,}")
     displayer.reset(model)
     #model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="ssm_bi_pxby_conv_81_dim_64_state_2_layer_last")
     complete_seq = np.array(hf_dataset["train"]['pixelbyte'][0])

### 2024-09-02 pixelbytes/__init__.py
start argparser

@@ -7,3 +7,4 @@ from .tokenizer import *
 from .train import *
 from .eval import *
 from .results import *
+from .config import *

### 2024-09-02 pixelbytes/config.py

@@ -1,0 +1,36 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabienfrfr
+"""
+
+import os
+import numpy as np
+from dotenv import load_dotenv
+
+load_dotenv()
+
+def get_default_palette():
+    palette_str = os.getenv('DEFAULT_PALETTE')
+    if palette_str:
+        palette_list = [int(x) for x in palette_str.split(',')]
+        return np.array(palette_list).reshape(-1, 3).astype(np.uint8)
+    else:
+        # Palette NES
+        return np.array([
+            [0x00, 0x00, 0x00], [0xfc, 0xfc, 0xfc], [0xf8, 0xf8, 0xf8], [0xbc, 0xbc, 0xbc],
+            [0x7c, 0x7c, 0x7c], [0xa4, 0xe4, 0xfc], [0x3c, 0xbc, 0xfc], [0x00, 0x78, 0xf8],
+            [0x00, 0x00, 0xfc], [0xb8, 0xb8, 0xf8], [0x68, 0x88, 0xfc], [0x00, 0x58, 0xf8],
+            [0x00, 0x00, 0xbc], [0xd8, 0xb8, 0xf8], [0x98, 0x78, 0xf8], [0x68, 0x44, 0xfc],
+            [0x44, 0x28, 0xbc], [0xf8, 0xb8, 0xf8], [0xf8, 0x78, 0xf8], [0xd8, 0x00, 0xcc],
+            [0x94, 0x00, 0x84], [0xf8, 0xa4, 0xc0], [0xf8, 0x58, 0x98], [0xe4, 0x00, 0x58],
+            [0xa8, 0x00, 0x20], [0xf0, 0xd0, 0xb0], [0xf8, 0x78, 0x58], [0xf8, 0x38, 0x00],
+            [0xa8, 0x10, 0x00], [0xfc, 0xe0, 0xa8], [0xfc, 0xa0, 0x44], [0xe4, 0x5c, 0x10],
+            [0x88, 0x14, 0x00], [0xf8, 0xd8, 0x78], [0xf8, 0xb8, 0x00], [0xac, 0x7c, 0x00],
+            [0x50, 0x30, 0x00], [0xd8, 0xf8, 0x78], [0xb8, 0xf8, 0x18], [0x00, 0xb8, 0x00],
[…]

### 2024-09-02 pixelbytes/dataset.py

@@ -4,6 +4,8 @@
 @author: fabienfrfr
 """
 
+from .config import DEFAULT_PALETTE
+
 import getpass
 from huggingface_hub import login
 
@@ -14,7 +16,6 @@ import numpy as np, cv2
 from numpy.lib.stride_tricks import as_strided
 from skimage import color
 from tqdm import tqdm
-
 
 ##### dataset
 class PxByDataset(Dataset):
@@ -50,26 +51,7 @@ def push_dataset(dataset, repo_name="ffurfaro/PixelBytes-Pokemon"):
     login(token)
     dataset.push_to_hub(repo_name)
 
-
 ## Construct pixelbytes columns
-# Palette NES personnalisée
-palette = np.array([
-    [0x00, 0x00, 0x00], [0xfc, 0xfc, 0xfc], [0xf8, 0xf8, 0xf8], [0xbc, 0xbc, 0xbc],
-    [0x7c, 0x7c, 0x7c], [0xa4, 0xe4, 0xfc], [0x3c, 0xbc, 0xfc], [0x00, 0x78, 0xf8],
-    [0x00, 0x00, 0xfc], [0xb8, 0xb8, 0xf8], [0x68, 0x88, 0xfc], [0x00, 0x58, 0xf8],
-    [0x00, 0x00, 0xbc], [0xd8, 0xb8, 0xf8], [0x98, 0x78, 0xf8], [0x68, 0x44, 0xfc],
-    [0x44, 0x28, 0xbc], [0xf8, 0xb8, 0xf8], [0xf8, 0x78, 0xf8], [0xd8, 0x00, 0xcc],
-    [0x94, 0x00, 0x84], [0xf8, 0xa4, 0xc0], [0xf8, 0x58, 0x98], [0xe4, 0x00, 0x58],
-    [0xa8, 0x00, 0x20], [0xf0, 0xd0, 0xb0], [0xf8, 0x78, 0x58], [0xf8, 0x38, 0x00],
-    [0xa8, 0x10, 0x00], [0xfc, 0xe0, 0xa8], [0xfc, 0xa0, 0x44], [0xe4, 0x5c, 0x10],
-    [0x88, 0x14, 0x00], [0xf8, 0xd8, 0x78], [0xf8, 0xb8, 0x00], [0xac, 0x7c, 0x00],
-    [0x50, 0x30, 0x00], [0xd8, 0xf8, 0x78], [0xb8, 0xf8, 0x18], [0x00, 0xb8, 0x00],
[…]

### 2024-09-02 pixelbytes/generator.py

@@ -81,5 +81,3 @@ class SequenceGenerator:
                                        [prev_token_id, token_id, 0],
                                        [0,0,0]])
         return matrix[None], sequence_clock
-
-

### 2024-09-02 pixelbytes/main.py

@@ -1,0 +1,87 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabienfrfr
+"""
+import argparse, sys
+
+def build_dataset(path, palette=None):
+    print(f"Building dataset from {path}")
+    if palette:
+        print(f"Using custom palette from {palette}")
+    # Add your dataset building logic here
+
+def train(model, learning_rate, batch_size, epochs):
+    print(f"Training {model} model")
+    print(f"Learning rate: {learning_rate}")
+    print(f"Batch size: {batch_size}")
+    print(f"Epochs: {epochs}")
+    # Add your training logic here
+
+def evaluate(metrics):
+    print(f"Evaluating model using metrics: {', '.join(metrics)}")
+    # Add your evaluation logic here
+    
+def generate():
+    # Generate and display results
+    print(f"Generating results in {format} format")
+    # Add your generation logic here
+
+def parse_arguments():
+    parser = argparse.ArgumentParser(description="Project task manager")
+    subparsers = parser.add_subparsers(dest='task', help='Task to execute')
+
+    # Build dataset parser
+    build_parser = subparsers.add_parser('build', help='Build the dataset')
+    build_parser.add_argument('--path', type=str, required=True, help='Path to the dataset')
+    build_parser.add_argument('--palette', type=str, help='Path to custom palette file')
+
+    # Train model parser
+    train_parser = subparsers.add_parser('train', help='Train the model')
+    train_parser.add_argument('--model', type=str, required=True, help='Model to train')
+    train_parser.add_argument('--learning-rate', type=float, default=0.001, help='Learning rate')
+    train_parser.add_argument('--batch-size', type=int, default=32, help='Batch size')
+    train_parser.add_argument('--epochs', type=int, default=10, help='Number of epochs')
+
+    # Evaluate model parser
+    eval_parser = subparsers.add_parser('evaluate', help='Evaluate the model')
+    eval_parser.add_argument('--metrics', nargs='+', default=['accuracy'], help='Metrics to evaluate')
+
+    # Generate output parser
+    gen_parser = subparsers.add_parser('generate', help='Generate and display results')
+    gen_parser.add_argument('--format', choices=['svg', 'png', 'jpg'], default='svg', help='Output format')
+
+    return parser.parse_args()
+
+def print_help():
+    print("Usage: python -m your_package.main [task] [task-specific-arguments]")
+    print("\nAvailable tasks:")
+    print("  build     - Build the dataset")
[…]

### 2024-09-02 tests.py

@@ -6,13 +6,14 @@ TESTS FILE
 """
 
 from pixelbytes import *
-from datasets import load_dataset
 
 ### basic test
 if __name__ == '__main__' :
-
+    print("All dev test OK ! you can do archaeology now! but some code doesn't work here, because it's not updated ;)")
+    """
     #quit()
     # common part
+    from datasets import load_dataset
     hf_dataset = load_dataset("ffurfaro/PixelBytes-Pokemon")
     tokenizer = PixelBytesTokenizer()
     # reconstruct bulbizarre
@@ -37,6 +38,7 @@ if __name__ == '__main__' :
     #images, text = displayer.process_sequence(complete_seq, window=64, gen_window=4)
     images, text = displayer.process_sequence(complete_seq, window=32, gen_window=8)
     displayer.show(images, text)
+    """
     """
     # plot
     train_results_path = 'models/train_results'
@@ -75,7 +77,7 @@ if __name__ == '__main__' :
         generated_sequence = generator.update_sequence(end_sequence[:,1,1])
         print(np.mean(reference_sequence - generated_sequence)) ## all zero with 50 lenght ! prefer 64 to be large
     """
-    '''
+    """
     tokenizer = PixelBytesTokenizer()
     model = SimpleRNNModel.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
     #model = bMamba.from_pretrained("ffurfaro/PixelBytes-Pokemon", subfolder="rnn_bi_pxby_81_dim_64_state_2_layer_last")
@@ -90,6 +92,7 @@ if __name__ == '__main__' :
     #for matrix in generator.generate_stream(start_sequence, max_length=25, temperature=0.7):
     #    print(matrix)
     # display generated sequence
+    """
     """
     # generate complete sequence
     first_sequence = np.array(hf_dataset["train"]['pixelbyte'][0])[:,1,1]
@@ -171,20 +174,3 @@ if __name__ == '__main__' :
     img = reconstruct_imgs(tokens[:np.random.randint(50,len(tokens)//2)])[0]["image"]
     plt.imshow(img);plt.show()
     """
-    
-    
-    
-    
-    
-    
-    '''
-    
-    
-    
-    
-    
-    
-    
-    
-    
-    

### 2024-09-02 pixelbytes/__init__.py
update code for parsing

@@ -8,3 +8,4 @@ from .train import *
 from .eval import *
 from .results import *
 from .config import *
+from .main import *

### 2024-09-02 pixelbytes/dataset.py

@@ -5,6 +5,7 @@
 """
 
 from .config import DEFAULT_PALETTE
+from .tokenizer import PixelBytesTokenizer
 
 import getpass
 from huggingface_hub import login
@@ -19,7 +20,7 @@ from tqdm import tqdm
 
 ##### dataset
 class PxByDataset(Dataset):
-    def __init__(self, pxby_columns, seq_length=256, stride=64):
+    def __init__(self, pxby_columns, seq_length=128, stride=32):
         self.data = [torch.tensor(sequence, dtype=torch.long) for sequence in pxby_columns]
         self.seq_length = seq_length
         self.stride = stride
@@ -74,7 +75,7 @@ def input_seq_construct(arr, dim=3, none_val=0, pix_sep=1, modal_sep=2):
     # flatten
     return result.reshape(-1, dim, dim)
 
-def add_pixelbyte_columns(image_caption_dataset, tokenizer=PixelBytesTokenizer(), palette=DEFAULT_PALETTE):
+def add_pixelbyte_columns(image_caption_dataset, is_pixel_style=False, tokenizer=PixelBytesTokenizer(), palette=DEFAULT_PALETTE):
     vocab = tokenizer.get_vocab()
     # vocabulary
     n = vocab[b'\x00']
@@ -84,20 +85,21 @@ def add_pixelbyte_columns(image_caption_dataset, tokenizer=PixelBytesTokenizer()
     vectorized_map = np.vectorize(lambda x,y,z: vocab.get((int(x), int(y), int(z)), None))
     # add pixelbytes columns
     pixelbytes = []
-    for row in tqdm(image_caption_dataset['train'], desc="Construct pixelbytes columns"):
+    for row in tqdm(image_caption_dataset, desc="Construct pixelbytes columns"):
         # get info
         Image = np.array(row['image'])
         Caption = row['caption'].encode('utf-8')
-        Shape = str(Image.shape[:-1]).encode('utf-8')
-        ### Shape part
-        Shape = np.array([vocab[bytes([x])] for x in Shape])[None]
-        Shape = input_seq_construct(Shape, dim=3, none_val=n, pix_sep=p, modal_sep=m)
         ### Image part
-        Image = image_paletization(Image, palette)
+        if is_pixel_style : Image = image_paletization(Image, palette) # already pixelized style
+        else : Image = image_pixelization(Image, palette) # Pixel reduce & palettize 
         # Separate RGB channel & Quantize
[…]

### 2024-09-02 pixelbytes/displays.py

@@ -138,4 +138,5 @@ class Displays:
         plt.tight_layout()
         plt.savefig('image.svg'); plt.savefig('image.png', dpi=300)
         plt.show()
+        plt.close()
 

### 2024-09-02 pixelbytes/main.py

@@ -3,62 +3,118 @@
 """
 @author: fabienfrfr
 """
-import argparse, sys
+import argparse, sys, os
+from . import *
+from datasets import load_dataset, load_from_disk
 
-def build_dataset(path, palette=None):
+def build_dataset_arg(path, palette=None):
     print(f"Building dataset from {path}")
     if palette:
         print(f"Using custom palette from {palette}")
-    # Add your dataset building logic here
+    # Dataset import
+    if os.path.isdir(path): dataset = load_from_disk(path)
+    else : dataset = load_dataset(path)["train"]
+    # dataset construct
+    dataset = add_pixelbyte_columns(dataset)
+    dataset.save_to_disk('PixelBytes-Dataset')
 
-def train(model, learning_rate, batch_size, epochs):
-    print(f"Training {model} model")
+def train_arg(data, seq_length, strides,  
+              model, dim, d_state, depth, grid, pxby, bidirect,
+              learning_rate, batch_size, epochs):
+    print(f"Training {model} model from {data}")
+    print(f"Sequence length and dataset stride : ({seq_length}, {strides})")
     print(f"Learning rate: {learning_rate}")
     print(f"Batch size: {batch_size}")
     print(f"Epochs: {epochs}")
-    # Add your training logic here
+    # Import dataset
+    if os.path.isdir(data): dataset = load_from_disk(data)
+    else : dataset = load_dataset(data)["train"]
+    ## Initalization
+    # data part
+    print(f"Construct data sequence for dataloader..")
+    dataset = dataset.train_test_split(test_size=0.1)
+    train_dataset = PxByDataset(dataset["train"]["pixelbyte"], seq_length=seq_length, stride=strides)
+    test_dataset = PxByDataset(dataset["test"]["pixelbyte"], seq_length=seq_length, stride=strides)
+    print(f"Data sequence OK!")
+    # model part
+    tokenizer = PixelBytesTokenizer()
+    vocab_size = tokenizer.__len__()
+    model_config = ModelConfig(dim=dim, d_state=d_state, depth=depth, vocab_size=vocab_size, 
+                               pxby_embed=grid, pembed=pxby, bidirectional=bidirect)
+    if model == "rnn" : model = SimpleRNNModel(model_config)
+    elif model == "ssm" : model = bMamba(model_config)
+    elif model == "att" : model = SimpleTransformerModel(model_config)
+    # train part
+    train_config = TrainConfig(model=model, model_config=model_config, dataset_name=data, batch_size=batch_size,
[…]

### 2024-09-02 pixelbytes/model.py

@@ -13,7 +13,7 @@ from dataclasses import dataclass
 try :
     from mamba_ssm.modules.mamba_simple import Mamba
 except :
-    print('No mamba_ssm CUDA modules installed...')
+    print('[INFO] Official mamba_ssm CUDA modules not found. Falling back to an unofficial implementation. Performance may be affected.')
     import mambapy.mamba as mambapy
     def Mamba(d_model, d_state, d_conv, expand) :
         config = mambapy.MambaConfig(d_model, d_state, d_conv, expand)

### 2024-09-02 pixelbytes/pokemon.py

@@ -4,6 +4,9 @@
 @author: fabienfrfr
 """
 
+import warnings
+warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
+
 from .dataset import image_pixelization
 
 from datasets import load_dataset, Dataset
@@ -78,16 +81,16 @@ def download_pkmn_miniature(url_pkmn):
         print(f"Download error")
         return None
 
+def arraybytes_to_png(arr, number):
+    pil_img = Image.fromarray(arr.astype('uint8'))
+    img_file_path = f"data/{number}.png" #img_byte_arr = io.BytesIO()
+    pil_img.save(img_file_path, format='PNG') #pil_img.save(img_byte_arr, format='PNG')
+    return img_file_path # img_byte_arr.getvalue()
+
 def create_pkmn_dataset(pkmn_dict, image_dir="data"):
     os.makedirs(image_dir, exist_ok=True)
     # Init pipeline format (if you caption : not used here // old)
     #captioner = pipeline("image-to-text", model="Salesforce/blip-image-captioning-base")
-    
-    def arraybytes_to_png(arr, number):
-        pil_img = Image.fromarray(arr.astype('uint8'))
-        img_file_path = f"data/{number}.png" #img_byte_arr = io.BytesIO()
-        pil_img.save(img_file_path, format='PNG') #pil_img.save(img_byte_arr, format='PNG')
-        return img_file_path # img_byte_arr.getvalue()
     # Generate image captionning
     #names = []
     captions = []

### 2024-09-02 pixelbytes/results.py

@@ -4,6 +4,9 @@
 @author: fabienfr
 """
 
+import warnings
+warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
+
 import os
 import pandas as pd
 import matplotlib.pyplot as plt

### 2024-09-02 pixelbytes/tokenizer.py

@@ -48,7 +48,7 @@ class PixelBytesTokenizer(PreTrainedTokenizer):
         return len(self.vocab)
 
     def get_vocab(self) -> Dict[str, int]:
-        return {str(k): v for k, v in self.vocab.items()}
+        return {k: v for k, v in self.vocab.items()}
 
     def _tokenize(self, text: str) -> List[str]:
         return list(text)

### 2024-09-02 pixelbytes/train.py

@@ -18,12 +18,12 @@ from .model import *
 
 @dataclass
 class TrainConfig:
-    hf_token : str
-    repo_name : str
     model : PreTrainedModel
     model_config : PretrainedConfig
     train_dataset : DataLoader
     test_dataset : DataLoader
+    hf_token : str = None
+    repo_name : str = None
     batch_size : int = 32
     learning_rate : int = 0.001
     num_epochs: int = 10
@@ -63,7 +63,7 @@ class Trainer:
         self.repo_name = config.dataset_name if (not(self.hf_token is None) and config.repo_name is None) else config.repo_name
 
     def train_and_evaluate(self):
-        if self.hf_token is None : print('No HF token given..')
+        if self.hf_token is None : print('No HF token given, only local save model')
         best_test_loss = float('inf')
         test_metrics = self._evaluate(self.test_loader)
         self.results.append({'epoch': 0, 'train_eval_loss' : test_metrics['loss'], 'train_accuracy': test_metrics['accuracy'], 'train_f1': test_metrics['f1'],

### 2024-09-02 tests.py

@@ -10,6 +10,9 @@ from pixelbytes import *
 ### basic test
 if __name__ == '__main__' :
     print("All dev test OK ! you can do archaeology now! but some code doesn't work here, because it's not updated ;)")
+    # python3 -m pixelbytes.main build
+    # python3 -m pixelbytes.main train --seq_length 512 --strides 256
+    # generate_arg('ffurfaro/PixelBytes-Pokemon','rnn_bi_pxby_81_dim_64_state_2_layer_last','svg') #dummy !
     """
     #quit()
     # common part

### 2024-09-09 pixelbytes/action_pixelbytes/tokenizer.py
optimize code for GPU

@@ -3,12 +3,12 @@
 """
 @author: fabienfrfr
 """
-
+import torch, unicodedata
+from torch import nn
 from transformers import PreTrainedTokenizer
-import numpy as np, os, unicodedata
 from skimage.color import rgb2lab
 from typing import List, Dict, Union, Tuple
-from scipy.spatial.distance import cdist
+import numpy as np
 from scipy.spatial import cKDTree
 from scipy import stats
 
@@ -40,12 +40,12 @@ DEFAULT_ACTION_STATE = np.linspace(-1, 1, 38).tolist()
 class ActionPixelBytesTokenizer(PreTrainedTokenizer):
     def __init__(self, **kwargs):
         ## Bytes (ASCII - UTF8) + Pixel (RGB NES Palette) + Action-space (Control & Audio)
-        ActionPixelbytes =  DEFAULT_BYTES + DEFAULT_PALETTE + DEFAULT_ACTION_STATE
-        self.vocab = {ActionPixelbytes[i] : i for i in range(len(ActionPixelbytes))}
+        ActionPixelbytes = DEFAULT_BYTES + DEFAULT_PALETTE + DEFAULT_ACTION_STATE
+        self.vocab = {ActionPixelbytes[i]: i for i in range(len(ActionPixelbytes))}
         super().__init__(**kwargs)
-        self.bytes_size = len(DEFAULT_BYTES) # first is null values
+        self.bytes_size = len(DEFAULT_BYTES)
         self.palet_size = len(DEFAULT_PALETTE)
-        self.LabPalette = rgb2lab(np.array(DEFAULT_PALETTE)[None, :, :] / 255.0)[0]
+        self.LabPalette = torch.tensor(rgb2lab(np.array(DEFAULT_PALETTE)[None, :, :] / 255.0)[0])
         self.ids_to_tokens = {v: k for k, v in self.vocab.items()}
 
     @property
@@ -53,7 +53,7 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         return len(self.vocab)
     
     def get_vocab(self) -> Dict[str, int]:
-        return {k: v for k, v in self.vocab.items()}
+        return dict(self.vocab)
 
     def _convert_token_to_id(self, token: Union[bytes, tuple]) -> int:
         return self.vocab.get(token, self.vocab.get(b'[UNK]', 0))
@@ -97,44 +97,47 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         if not inputs:
[…]

### 2024-09-09 pixelbytes/action_pixelbytes/model.py
optimize code for GPU

@@ -4,10 +4,10 @@
 @author: fabienfrfr
 """
 
-import torch
+import torch, random
 import torch.nn as nn
 import torch.optim as optim
-from torch.utils.data import Dataset, DataLoader
+from torch.utils.data import Dataset, DataLoader, Sampler
 from torch.nn.utils.rnn import pad_sequence
 from tqdm import tqdm
 import numpy as np
@@ -23,7 +23,7 @@ class TokenizedDataset(Dataset):
         self._preprocess_data(data, tokenizer)
 
     def _preprocess_data(self, data, tokenizer):
-        for item in data:
+        for item in data :
             tokenized = tokenizer(text=item.get('text'),
                                   image=item.get('image'),
                                   audio=item.get('audio'))
@@ -50,6 +50,25 @@ class TokenizedDataset(Dataset):
                 return i, idx * self.stride
             idx -= num_sequences
         raise IndexError("Index out of range")
+    
+    def get_num_sequences(self, item):
+        return max(1, (len(item['input_ids']) - self.seq_length) // self.stride + 1)
+
+class ShuffledSampler(Sampler):
+    def __init__(self, data_source, seed=None):
+        self.data_source = data_source
+        self.seed = seed
+        self.indices = list(range(len(data_source)))
+        
+    def __iter__(self):
+        if self.seed is not None:
+            random.seed(self.seed)
+        random.shuffle(self.indices)
+        for idx in self.indices:
+            yield idx
+
+    def __len__(self):
+        return len(self.data_source)
 
 def collate_fn(batch):
     input_ids = [item['input_ids'] for item in batch]
@@ -98,31 +117,33 @@ class BestPreTrainedModel(PreTrainedModel):
             return current_input
 
     def train_model(self, dataloader, optimizer, criterion, device, scaler, epochs, accumulation_steps=4):
+        best_loss = float('inf')
         for epoch in range(epochs):
             self.train()
-            total_loss = 0
-            optimizer.zero_grad()
-            for i, batch in enumerate(tqdm(dataloader)):
-                input_ids = batch['input_ids'].to(device)
-                labels = batch['labels'].to(device)
+            total_loss = total_correct = total_predictions = 0
+            for i, batch in enumerate(tqdm(dataloader, desc=f"Epoch {epoch+1}/{epochs}")):
[…]

### 2024-09-09 pixelbytes/action_pixelbytes/tokenizer.py

@@ -38,11 +38,12 @@ DEFAULT_ACTION_STATE = np.linspace(-1, 1, 38).tolist()
 
 ##### Tokenizer
 class ActionPixelBytesTokenizer(PreTrainedTokenizer):
-    def __init__(self, **kwargs):
+    def __init__(self, data_slicing=1, **kwargs):
         ## Bytes (ASCII - UTF8) + Pixel (RGB NES Palette) + Action-space (Control & Audio)
         ActionPixelbytes = DEFAULT_BYTES + DEFAULT_PALETTE + DEFAULT_ACTION_STATE
         self.vocab = {ActionPixelbytes[i]: i for i in range(len(ActionPixelbytes))}
         super().__init__(**kwargs)
+        self.slicing = data_slicing
         self.bytes_size = len(DEFAULT_BYTES)
         self.palet_size = len(DEFAULT_PALETTE)
         self.LabPalette = torch.tensor(rgb2lab(np.array(DEFAULT_PALETTE)[None, :, :] / 255.0)[0])
@@ -103,7 +104,7 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         text = [bytes([b]) for b in unicodedata.normalize('NFKD', text.lower()).encode('ASCII', 'ignore')]
         return torch.tensor(self.convert_tokens_to_ids(text) + [2], dtype=torch.long).unsqueeze(0).unsqueeze(0)
 
-    def process_image(self, image, slicing=3):
+    def process_image(self, image):
         n_frames = getattr(image, "n_frames", 1)
         frames = torch.empty((n_frames, image.height, image.width), dtype=torch.long)
         for i in range(n_frames):
@@ -111,19 +112,19 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
             frame = torch.tensor(rgb2lab(np.array(image.convert('RGB')) / 255.0))
             distances = torch.cdist(frame.reshape(-1, 3), self.LabPalette)
             frames[i] = distances.argmin(dim=1).reshape(image.size[::-1])
-        frames = (frames + self.bytes_size)[::slicing, ::slicing, ::slicing]
+        frames = (frames + self.bytes_size)[::self.slicing, ::self.slicing, ::self.slicing]
         added_column = torch.ones((frames.shape[0], frames.shape[1], 1), dtype=torch.long)
         added_column[:, -1, :] = 2
         return torch.cat((frames, added_column), dim=2)
 
[…]

### 2024-09-09 pixelbytes/action_pixelbytes/model.py
adding metrics measurment

@@ -4,6 +4,7 @@
 @author: fabienfrfr
 """
 
+import os
 import torch, random
 import torch.nn as nn
 import torch.optim as optim
@@ -40,8 +41,8 @@ class TokenizedDataset(Dataset):
         item_idx, start_idx = self._get_item_and_start_indices(idx)
         tokenized = self.tokenized_data[item_idx]
         end_idx = start_idx + self.seq_length
-        return {'input_ids': torch.tensor(tokenized['input_ids'][start_idx:end_idx], dtype=torch.long),
-                'labels': torch.tensor(tokenized['labels'][start_idx:end_idx], dtype=torch.long)}
+        return {'input_ids': torch.as_tensor(tokenized['input_ids'][start_idx:end_idx], dtype=torch.long),
+                'labels': torch.as_tensor(tokenized['labels'][start_idx:end_idx], dtype=torch.long)}
 
     def _get_item_and_start_indices(self, idx):
         for i, item in enumerate(self.tokenized_data):
@@ -116,34 +117,47 @@ class BestPreTrainedModel(PreTrainedModel):
                 current_input = torch.cat([current_input, next_token], dim=1) # it's wrong : need to tokenizer decode
             return current_input
 
-    def train_model(self, dataloader, optimizer, criterion, device, scaler, epochs, accumulation_steps=4):
+    def train_model(self, train_dataloader, val_dataloader, optimizer, criterion, device, scaler, epochs, accumulation_steps=4, eval_every=5):
         best_loss = float('inf')
+        val_loss, val_accuracy = self._process_epoch(val_dataloader, None, criterion, device, None, accumulation_steps)
+        metrics = [{'epoch': 0,'train_loss': val_loss,'train_accuracy': val_accuracy, 'val_loss': val_loss, 'val_accuracy': val_accuracy}]
+        print(f"Validation Loss: {val_loss:.4f}, Validation Accuracy: {val_accuracy:.4f}")
         for epoch in range(epochs):
-            self.train()
-            total_loss = total_correct = total_predictions = 0
-            for i, batch in enumerate(tqdm(dataloader, desc=f"Epoch {epoch+1}/{epochs}")):
+            train_loss, train_accuracy = self._process_epoch(train_dataloader, optimizer, criterion, device, scaler, accumulation_steps)
+            print(f"Epoch {epoch+1}/{epochs}, Train Loss: {train_loss:.4f}, Train Accuracy: {train_accuracy:.4f}")
[…]

### 2024-09-09 pixelbytes/action_pixelbytes/model.py
true previous action for auto-encoder

@@ -11,7 +11,7 @@ import torch.optim as optim
 from torch.utils.data import Dataset, DataLoader, Sampler
 from torch.nn.utils.rnn import pad_sequence
 from tqdm import tqdm
-import numpy as np
+import numpy as np, pandas as pd
 from transformers import PreTrainedModel, PretrainedConfig
 from torch.cuda.amp import autocast, GradScaler
 import torch.nn.functional as F
@@ -106,15 +106,17 @@ class BestPreTrainedModel(PreTrainedModel):
         output, _ = self.lstm(embedded)
         return self.fc(output)
 
-    def generate(self, input_ids, max_length, temperature=1.0):
+    def generate(self, input_ids, num_generate, temperature=1.0):
         self.eval()
         with torch.no_grad():
-            current_input = input_ids
-            for _ in range(max_length - input_ids.size(1)):
-                outputs = self(current_input)
-                next_token_logits = outputs[:, -1, :] / temperature
-                next_token = torch.multinomial(torch.softmax(next_token_logits, dim=-1), num_samples=1)
-                current_input = torch.cat([current_input, next_token], dim=1) # it's wrong : need to tokenizer decode
+            current_input = input_ids.clone()
+            for i in range(num_generate):
+                # Generate next token
+                next_token = torch.multinomial(
+                    torch.softmax(self(current_input)[:, -1, :] / temperature, dim=-1),
+                    num_samples=1)
+                # Replace the (i+1)th token from the end --> not true generator for image or action_state
+                current_input[:, -(i+1)] = next_token.squeeze(-1)
             return current_input
 
     def train_model(self, train_dataloader, val_dataloader, optimizer, criterion, device, scaler, epochs, accumulation_steps=4, eval_every=5):

### 2024-09-09 pixelbytes/action_pixelbytes/tokenizer.py

@@ -136,14 +136,16 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
             padded[:-1, 1:-1, 1:-1],# (t-1, i, j) : value in t-1
             padded[:-1, 2:, 1:-1],  # (t-1, i+1, j) : down value in t-1
             padded[1:, :-2, :-2],   # (t, i-1, j-1) : up-left value in t
-            padded[1:, 1:-1, :-2],  # (t, i, j-1) : up value in t
-            padded[1:, :-2, 1:-1]]   # (t, i-1, j) : left value in t
-        context = torch.stack(slices, dim=-1)
-        return context.reshape(-1, 6), context_array.reshape(-1, 1)
+            padded[1:, :-2, 1:-1],  # (t, i-1, j) : up value in t
+            padded[1:, 1:-1, :-2]]  # (t, i, j-1) : wrong previous value
+        context = torch.stack(slices, dim=-1).reshape(-1, 6)
+        targets = context_array.reshape(-1, 1)
+        context[1:,-1] = targets[:-1,0] # true previous value (AutoEncoder)
+        return context, targets
 
 ### basic test
 if __name__ == '__main__' :
-    tokenizer = ActionPixelBytesTokenizer()
+    tokenizer = ActionPixelBytesTokenizer(data_slicing=4)
     from datasets import load_dataset
     pxby_dataset = load_dataset("ffurfaro/PixelBytes-PokemonAll")
     bulbi = img = pxby_dataset['train']['image'][0]

### 2024-09-09 pixelbytes/action_pixelbytes/model.py
udate model

@@ -24,34 +24,35 @@ class TokenizedDataset(Dataset):
         self._preprocess_data(data, tokenizer)
 
     def _preprocess_data(self, data, tokenizer):
-        for item in data :
+        for item in data:
             tokenized = tokenizer(text=item.get('text'),
                                   image=item.get('image'),
                                   audio=item.get('audio'))
             self.tokenized_data.append(tokenized)
-        self._calculate_total_sequences()
+        self.total_sequences = sum(self.get_num_sequences(item) for item in self.tokenized_data)
 
-    def _calculate_total_sequences(self):
-        self.total_sequences = sum(max(1, (len(item['input_ids']) - self.seq_length) // self.stride + 1)
-                                   for item in self.tokenized_data)
     def __len__(self):
         return self.total_sequences
 
     def __getitem__(self, idx):
+        idx = idx % self.total_sequences
         item_idx, start_idx = self._get_item_and_start_indices(idx)
         tokenized = self.tokenized_data[item_idx]
         end_idx = start_idx + self.seq_length
-        return {'input_ids': torch.as_tensor(tokenized['input_ids'][start_idx:end_idx], dtype=torch.long),
-                'labels': torch.as_tensor(tokenized['labels'][start_idx:end_idx], dtype=torch.long)}
+
+        input_ids = tokenized['input_ids'][start_idx:end_idx % len(tokenized['input_ids'])]
+        labels = tokenized['labels'][start_idx:end_idx % len(tokenized['labels'])]
+        return {'input_ids': torch.as_tensor(input_ids, dtype=torch.long), 'labels': torch.as_tensor(labels, dtype=torch.long)}
 
     def _get_item_and_start_indices(self, idx):
         for i, item in enumerate(self.tokenized_data):
-            num_sequences = max(1, (len(item['input_ids']) - self.seq_length) // self.stride + 1)
+            num_sequences = self.get_num_sequences(item)
             if idx < num_sequences:
-                return i, idx * self.stride
+                start_idx = (idx * self.stride) % len(item['input_ids'])
+                return i, start_idx
             idx -= num_sequences
         raise IndexError("Index out of range")
-    
+
     def get_num_sequences(self, item):
         return max(1, (len(item['input_ids']) - self.seq_length) // self.stride + 1)
 

### 2024-09-09 pixelbytes/action_pixelbytes/model.py
udate model

@@ -83,13 +83,14 @@ def collate_fn(batch):
 
 class ModelConfig_(PretrainedConfig):
     model_type = "lstm"
-    def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, **kwargs):
+    def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, auto_encoder=False, **kwargs):
         super().__init__(**kwargs)
         self.vocab_size = vocab_size
         self.embed_size = embed_size
         self.hidden_size = hidden_size
         self.num_layers = num_layers
         self.pxby_dim = pxby_dim
+        self.auto_encoder = auto_encoder
 
 class BestPreTrainedModel(PreTrainedModel):
     config_class = ModelConfig_
@@ -100,12 +101,15 @@ class BestPreTrainedModel(PreTrainedModel):
         self.embedding = nn.Embedding(config.vocab_size, config.embed_size, padding_idx=0)
         self.lstm = nn.LSTM(config.embed_size * config.pxby_dim, config.hidden_size, config.num_layers, batch_first=True)
         self.fc = nn.Linear(config.hidden_size, config.vocab_size)
+        self.pxby_dim, self.auto_encoder = config.pxby_dim, config.auto_encoder
 
     def forward(self, input_ids):
         batch_size, seq_len, context_size = input_ids.shape
         embedded = self.embedding(input_ids).view(batch_size, seq_len, -1)
         output, _ = self.lstm(embedded)
-        return self.fc(output)
+        if self.auto_encoder: # Reshape output to (batch_size, seq_len, pxby_dim, hidden_size)
+            output = output.view(batch_size, seq_len, self.pxby_dim, -1) # For Shape: (batch_size, seq_len, pxby_dim, vocab_size)
+        return self.fc(output) # (Else) Shape: (batch_size, seq_len, vocab_size)
 
     def generate(self, input_ids, num_generate, temperature=1.0):
         self.eval()
@@ -116,7 +120,7 @@ class BestPreTrainedModel(PreTrainedModel):
                 next_token = torch.multinomial(
                     torch.softmax(self(current_input)[:, -1, :] / temperature, dim=-1),
                     num_samples=1)
-                # Replace the (i+1)th token from the end --> not true generator for image or action_state
+                # Replace the (i+1)th token from the end --> not true generator for image or action_state (adapt if auto-encoder)
[…]

### 2024-09-10 pixelbytes/__init__.py
change project architecture

@@ -1,12 +1,3 @@
-from .dataset import *
-from .generator import *
 from .model import *
-from .pokemon import *
-from .displays import *
 from .tokenizer import *
-from .train import *
-from .eval import *
-from .results import *
-from .config import *
-from .main import *
-from .action_pixelbytes import *
+from .pixelize import *

### 2024-09-10 pixelbytes/action_pixelbytes/__init__.py

@@ -1,3 +1,0 @@
-from .model import *
-from .tokenizer import *
-from .pixelize import *

### 2024-09-10 pixelbytes/action_pixelbytes/model.py

@@ -1,221 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-@author: fabienfrfr
-"""
-
-import os
-import torch, random
-import torch.nn as nn
-import torch.optim as optim
-from torch.utils.data import Dataset, DataLoader, Sampler
-from torch.nn.utils.rnn import pad_sequence
-from tqdm import tqdm
-import numpy as np, pandas as pd
-from transformers import PreTrainedModel, PretrainedConfig
-from torch.cuda.amp import autocast, GradScaler
-import torch.nn.functional as F
-
-class TokenizedDataset(Dataset):
-    def __init__(self, data, tokenizer, seq_length=1024, stride=512):
-        self.seq_length = seq_length
-        self.stride = stride
-        self.tokenized_data = []
-        self._preprocess_data(data, tokenizer)
-
-    def _preprocess_data(self, data, tokenizer):
-        for item in data:
-            tokenized = tokenizer(text=item.get('text'),
-                                  image=item.get('image'),
-                                  audio=item.get('audio'))
-            self.tokenized_data.append(tokenized)
-        self.total_sequences = sum(self.get_num_sequences(item) for item in self.tokenized_data)
-
-    def __len__(self):
-        return self.total_sequences
-
-    def __getitem__(self, idx):
-        idx = idx % self.total_sequences
-        item_idx, start_idx = self._get_item_and_start_indices(idx)
-        tokenized = self.tokenized_data[item_idx]
-        end_idx = start_idx + self.seq_length
-
-        input_ids = tokenized['input_ids'][start_idx:end_idx % len(tokenized['input_ids'])]
-        labels = tokenized['labels'][start_idx:end_idx % len(tokenized['labels'])]
-        return {'input_ids': torch.as_tensor(input_ids, dtype=torch.long), 'labels': torch.as_tensor(labels, dtype=torch.long)}
-
-    def _get_item_and_start_indices(self, idx):
-        for i, item in enumerate(self.tokenized_data):
-            num_sequences = self.get_num_sequences(item)
-            if idx < num_sequences:
-                start_idx = (idx * self.stride) % len(item['input_ids'])
-                return i, start_idx
-            idx -= num_sequences
-        raise IndexError("Index out of range")
-
-    def get_num_sequences(self, item):
[…]

### 2024-09-10 pixelbytes/action_pixelbytes/pokemon.py

@@ -1,151 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-@author: fabienfrfr
-"""
-
-import os, re, shutil, requests
-import numpy as np, pandas as pd
-from bs4 import BeautifulSoup
-from scipy import signal
-import soundfile as sf #import librosa
-from datasets import load_dataset, Dataset, Features, Value, Image, Audio
-from huggingface_hub import login
-
-class PokemonSpriteDownloader:
-    BASE_URL = "https://www.pokencyclopedia.info"
-    SPRITE_URLS = {'front': f"{BASE_URL}/en/index.php?id=sprites/gen5/ani_black-white",
-                   'back': f"{BASE_URL}/en/index.php?id=sprites/gen5/ani-b_black-white" }
-
-    def download_sprites(self, sprite_type='front'):
-        # Create a folder for the sprites
-        folder_name = f"pokemon_{sprite_type}_gifs"
-        os.makedirs(folder_name, exist_ok=True)
-        # Fetch the webpage and parse it
-        soup = BeautifulSoup(requests.get(self.SPRITE_URLS[sprite_type]).content, 'html.parser')
-        pokemon_data = []
-        # Find all GIF images
-        for gif in soup.find_all('img', src=lambda s: s.endswith('.gif')):
-            gif_url = f"{self.BASE_URL}/{gif['src'].lstrip('/')}"
-            file_name = gif_url.split('/')[-1]
-            pokemon_name = gif.get('alt', 'Unknown').split()[0]
-            # Download the GIF
-            try:
-                response = requests.get(gif_url)
-                response.raise_for_status()
-                with open(os.path.join(folder_name, file_name), 'wb') as file:
-                    file.write(response.content)
-                pokemon_data.append({'Pokemon Name': pokemon_name, 'File Name': file_name})
-                print(f"Downloaded: {file_name}")
-            except requests.RequestException as e:
-                print(f"Error downloading {file_name}: {e}")
-        # Save data to CSV using Pandas
-        pd.DataFrame(pokemon_data).to_csv(f"pokemon_{sprite_type}_data.csv", index=False)
-        print(f"Finished! All {sprite_type} GIFs downloaded and data saved.")
-
-    def download_all(self):
-        for sprite_type in self.SPRITE_URLS:
-            self.download_sprites(sprite_type)
-
-class PokemonCryDownloader:
-    BASE_URL = "https://www.pokepedia.fr/Fichier:Cri_{:04d}_HOME.ogg"
[…]

### 2024-09-10 pixelbytes/action_pixelbytes/tokenizer.py

@@ -1,159 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-@author: fabienfrfr
-"""
-import torch, unicodedata
-from torch import nn
-from transformers import PreTrainedTokenizer
-from skimage.color import rgb2lab
-from typing import List, Dict, Union, Tuple
-import numpy as np
-from scipy.spatial import cKDTree
-from scipy import stats
-
-## Bytes (ASCII - UTF8)
-DEFAULT_BYTES = [b'\x00', b'\t', b'\n', b' ', b'"', b"'", b'(', b')', b'*', b',', b'-', b'+', 
-                b'.', b'0', b'1', b'2', b'3', b'4', b'5', b'6', b'7', b'8', b'9', b'\xc2', 
-                b'\xa0', b':', b'[', b']', b';', b'/', b'%', b'!', b'a', b'b', b'c', b'd', b'e', 
-                b'f', b'g', b'h', b'i', b'j', b'k', b'l', b'm', b'n', b'o', b'p', b'q', b'r', 
-                b's', b't', b'u', b'v', b'w', b'x', b'y', b'z']
-## Pixel (RGB NES Palette)
-DEFAULT_PALETTE = [(0x00, 0x00, 0x00), (0xfc, 0xfc, 0xfc), (0xf8, 0xf8, 0xf8), (0xbc, 0xbc, 0xbc),
-                    (0x7c, 0x7c, 0x7c), (0xa4, 0xe4, 0xfc), (0x3c, 0xbc, 0xfc), (0x00, 0x78, 0xf8),
-                    (0x00, 0x00, 0xfc), (0xb8, 0xb8, 0xf8), (0x68, 0x88, 0xfc), (0x00, 0x58, 0xf8),
-                    (0x00, 0x00, 0xbc), (0xd8, 0xb8, 0xf8), (0x98, 0x78, 0xf8), (0x68, 0x44, 0xfc),
-                    (0x44, 0x28, 0xbc), (0xf8, 0xb8, 0xf8), (0xf8, 0x78, 0xf8), (0xd8, 0x00, 0xcc),
-                    (0x94, 0x00, 0x84), (0xf8, 0xa4, 0xc0), (0xf8, 0x58, 0x98), (0xe4, 0x00, 0x58),
[…]

### 2024-09-10 pixelbytes/model.py

@@ -4,184 +4,245 @@
 @author: fabienfrfr
 """
 
-import torch, re
+import os
+import torch, random
 import torch.nn as nn
-from huggingface_hub import HfApi, create_repo, whoami
+import torch.optim as optim
+from torch.utils.data import Dataset, DataLoader, Sampler
+from torch.nn.utils.rnn import pad_sequence
+from tqdm import tqdm
+import numpy as np, pandas as pd
 from transformers import PreTrainedModel, PretrainedConfig
-from dataclasses import dataclass
+from torch.cuda.amp import autocast, GradScaler
+import torch.nn.functional as F
 
-try :
-    from mamba_ssm.modules.mamba_simple import Mamba
-except :
-    print('[INFO] Official mamba_ssm CUDA modules not found. Falling back to an unofficial implementation. Performance may be affected.')
-    import mambapy.mamba as mambapy
-    def Mamba(d_model, d_state, d_conv, expand) :
-        config = mambapy.MambaConfig(d_model, d_state, d_conv, expand)
-        return mambapy.Mamba(config) ## doesnt works.. need to change prefix and state_dict..
+class TokenizedDataset(Dataset):
+    def __init__(self, data, tokenizer, seq_length=1024, stride=512):
+        self.seq_length = seq_length
+        self.stride = stride
+        self.tokenized_data = []
+        self._preprocess_data(data, tokenizer)
 
-### Multimodal embedding        
-class PxByEmbed(nn.Module):
-    def __init__(self, vocab_size, dim, is_pemb=True, k=3):
-        super().__init__()
-        self.d_model, self.k = dim, k
-        self.e_model = max(9, dim // (k**2))
-        self.is_p_embed = is_pemb
-        # Spatially adaptive embedding combination
-        self.alpha = nn.Parameter(torch.rand(1, 1, k, k))
-        self.projection = nn.Linear(self.e_model * k * k, dim)
-        # Final normalization
-        self.norm = nn.LayerNorm(dim)
-        # Classic text embedding
-        self.linear_embedding = nn.Embedding(vocab_size, self.e_model)
-        # Local embedding patch
-        self.patch_embedding = nn.Conv2d(in_channels=self.e_model, out_channels=self.e_model, 
-                                         kernel_size=k, stride=1, padding=k//2) if is_pemb else None
-    def forward(self, x):
-        # shape : x : (B, L, M=k, N=k) : long
-        B, L, M, N = x.shape
[…]

### 2024-09-10 pixelbytes/modules/__init__.py

@@ -1,0 +1,12 @@
+from .dataset import *
+from .generator import *
+from .model import *
+from .pokemon import *
+from .displays import *
+from .tokenizer import *
+from .train import *
+from .eval import *
+from .results import *
+from .config import *
+from .main import *
+## old version of code

### 2024-09-10 pixelbytes/modules/model.py

@@ -1,0 +1,187 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabienfrfr
+"""
+
+import torch, re
+import torch.nn as nn
+from huggingface_hub import HfApi, create_repo, whoami
+from transformers import PreTrainedModel, PretrainedConfig
+from dataclasses import dataclass
+
+try :
+    from mamba_ssm.modules.mamba_simple import Mamba
+except :
+    print('[INFO] Official mamba_ssm CUDA modules not found. Falling back to an unofficial implementation. Performance may be affected.')
+    import mambapy.mamba as mambapy
+    def Mamba(d_model, d_state, d_conv, expand) :
+        config = mambapy.MambaConfig(d_model, d_state, d_conv, expand)
+        return mambapy.Mamba(config) ## doesnt works.. need to change prefix and state_dict..
+
+### Multimodal embedding        
+class PxByEmbed(nn.Module):
+    def __init__(self, vocab_size, dim, is_pemb=True, k=3):
+        super().__init__()
+        self.d_model, self.k = dim, k
+        self.e_model = max(9, dim // (k**2))
+        self.is_p_embed = is_pemb
+        # Spatially adaptive embedding combination
+        self.alpha = nn.Parameter(torch.rand(1, 1, k, k))
+        self.projection = nn.Linear(self.e_model * k * k, dim)
+        # Final normalization
+        self.norm = nn.LayerNorm(dim)
+        # Classic text embedding
+        self.linear_embedding = nn.Embedding(vocab_size, self.e_model)
+        # Local embedding patch
+        self.patch_embedding = nn.Conv2d(in_channels=self.e_model, out_channels=self.e_model, 
+                                         kernel_size=k, stride=1, padding=k//2) if is_pemb else None
+    def forward(self, x):
+        # shape : x : (B, L, M=k, N=k) : long
+        B, L, M, N = x.shape
+        assert M == N == self.k, f"Input spatial dimensions should be {self.k}x{self.k}, but got {M}x{N}"
+        # Linear embedding
+        x = self.linear_embedding(x.view(B*L, M, N))  # (B*L, M, N, e_model)
+        x = x.permute(0, 3, 1, 2)  # (B*L, e_model, M, N)
+        # Combine linear (and patch) embedding with alpha
+        if not(self.is_p_embed): x = torch.sigmoid(self.alpha) * x  # (B*L, e_model, M, N)
[…]

### 2024-09-10 pixelbytes/modules/pokemon.py

@@ -1,0 +1,121 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabienfrfr
+"""
+
+import warnings
+warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
+
+from .dataset import image_pixelization
+
+from datasets import load_dataset, Dataset
+import io, os, getpass
+from huggingface_hub import HfApi, login
+#from transformers import pipeline
+
+from PIL import Image
+import cv2, re, unicodedata
+import requests, pandas as pd
+from bs4 import BeautifulSoup
+import numpy as np, pylab as plt
+
+
+def get_pkmns_miniatures():
+    url = "https://www.pokepedia.fr/Liste_des_Pok%C3%A9mon_dans_l%27ordre_du_Pok%C3%A9dex_National"
+    headers = {'User-Agent': 'Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/91.0.4472.124 Safari/537.36'}
+    response = requests.get(url, headers=headers)
+    print(f"Status Code: {response.status_code}")
+    soup = BeautifulSoup(response.content, 'html.parser')
+    pokemon_miniatures = {}
+    # Find line in table
+    rows = soup.find_all('tr')
+    for row in rows:
+        cells = row.find_all('td')
+        if len(cells) >= 3:  # Assurez-vous qu'il y a suffisamment de cellules
+            number_cell, image_cell, name_cell = cells[0], cells[1], cells[3] # English version
+            if number_cell.text.strip().isdigit():
+                number = number_cell.text.strip().zfill(4)
+                name = name_cell.text.strip()
+                img = image_cell.find('img')
+                if img and 'src' in img.attrs:
+                    image_url = "https://www.pokepedia.fr" + img['src'] if img['src'].startswith('/') else img['src']
+                    pokemon_miniatures[number] = {'name': name, 'url': image_url}
+    return pokemon_miniatures
+
+## Caption part
+def get_pkmn_info(pkmn_name):
+    url = f"https://bulbapedia.bulbagarden.net/wiki/{pkmn_name}_(Pokémon)"
+    response = requests.get(url)
+    response.raise_for_status()
+    soup = BeautifulSoup(response.text, 'html.parser')
[…]

### 2024-09-10 pixelbytes/modules/tokenizer.py

@@ -1,0 +1,114 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+@author: fabien
+
+in dev (pretrained part)
+"""
+
+from huggingface_hub import login, hf_hub_download
+
+from transformers import PreTrainedTokenizer
+import numpy as np, os
+from typing import List, Dict, Union, Tuple
+
+##### Tokenizer
+class PixelBytesTokenizer(PreTrainedTokenizer):
+    def __init__(self, vocab=None):
+        if vocab == None :
+            Pixelbytes_tokens =  [
+                ## Bytes (ASCII - UTF8)
+                b'\x00', b'\t', b'\n', b' ', b'"', b"'", b'(', b')', b'*', b',', b'-', b'+', 
+                b'.', b'0', b'1', b'2', b'3', b'4', b'5', b'6', b'7', b'8', b'9', b'\xc2', 
+                b'\xa0', b':', b'[', b']', b';', b'/', b'%', b'!', b'a', b'b', b'c', b'd', b'e', 
+                b'f', b'g', b'h', b'i', b'j', b'k', b'l', b'm', b'n', b'o', b'p', b'q', b'r', 
+                b's', b't', b'u', b'v', b'w', b'x', b'y', b'z',
+                ## Pixel (RGB NES Palette)
+                (0x00, 0x00, 0x00), (0xfc, 0xfc, 0xfc), (0xf8, 0xf8, 0xf8), (0xbc, 0xbc, 0xbc),
+                (0x7c, 0x7c, 0x7c), (0xa4, 0xe4, 0xfc), (0x3c, 0xbc, 0xfc), (0x00, 0x78, 0xf8),
+                (0x00, 0x00, 0xfc), (0xb8, 0xb8, 0xf8), (0x68, 0x88, 0xfc), (0x00, 0x58, 0xf8),
+                (0x00, 0x00, 0xbc), (0xd8, 0xb8, 0xf8), (0x98, 0x78, 0xf8), (0x68, 0x44, 0xfc),
+                (0x44, 0x28, 0xbc), (0xf8, 0xb8, 0xf8), (0xf8, 0x78, 0xf8), (0xd8, 0x00, 0xcc),
[…]

### 2024-09-10 pixelbytes/pokemon.py

@@ -4,118 +4,148 @@
 @author: fabienfrfr
 """
 
-import warnings
-warnings.filterwarnings("ignore", category=UserWarning, module="matplotlib")
-
-from .dataset import image_pixelization
-
-from datasets import load_dataset, Dataset
-import io, os, getpass
-from huggingface_hub import HfApi, login
-#from transformers import pipeline
-
-from PIL import Image
-import cv2, re, unicodedata
-import requests, pandas as pd
+import os, re, shutil, requests
+import numpy as np, pandas as pd
 from bs4 import BeautifulSoup
-import numpy as np, pylab as plt
+from scipy import signal
+import soundfile as sf #import librosa
+from datasets import load_dataset, Dataset, Features, Value, Image, Audio
+from huggingface_hub import login
 
+class PokemonSpriteDownloader:
+    BASE_URL = "https://www.pokencyclopedia.info"
+    SPRITE_URLS = {'front': f"{BASE_URL}/en/index.php?id=sprites/gen5/ani_black-white",
+                   'back': f"{BASE_URL}/en/index.php?id=sprites/gen5/ani-b_black-white" }
 
-def get_pkmns_miniatures():
-    url = "https://www.pokepedia.fr/Liste_des_Pok%C3%A9mon_dans_l%27ordre_du_Pok%C3%A9dex_National"
-    headers = {'User-Agent': 'Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/91.0.4472.124 Safari/537.36'}
-    response = requests.get(url, headers=headers)
-    print(f"Status Code: {response.status_code}")
-    soup = BeautifulSoup(response.content, 'html.parser')
-    pokemon_miniatures = {}
-    # Find line in table
-    rows = soup.find_all('tr')
-    for row in rows:
-        cells = row.find_all('td')
-        if len(cells) >= 3:  # Assurez-vous qu'il y a suffisamment de cellules
-            number_cell, image_cell, name_cell = cells[0], cells[1], cells[3] # English version
-            if number_cell.text.strip().isdigit():
-                number = number_cell.text.strip().zfill(4)
-                name = name_cell.text.strip()
-                img = image_cell.find('img')
-                if img and 'src' in img.attrs:
-                    image_url = "https://www.pokepedia.fr" + img['src'] if img['src'].startswith('/') else img['src']
[…]

### 2024-09-10 pixelbytes/tokenizer.py

@@ -1,57 +1,60 @@
 #!/usr/bin/env python3
 # -*- coding: utf-8 -*-
 """
-@author: fabien
-
-in dev (pretrained part)
+@author: fabienfrfr
 """
-
-from huggingface_hub import login, hf_hub_download
-
+import torch, unicodedata
+from torch import nn
 from transformers import PreTrainedTokenizer
-import numpy as np, os
+from skimage.color import rgb2lab
 from typing import List, Dict, Union, Tuple
+import numpy as np
+from scipy.spatial import cKDTree
+from scipy import stats
 
-##### Tokenizer
-class PixelBytesTokenizer(PreTrainedTokenizer):
-    def __init__(self, vocab=None):
-        if vocab == None :
-            Pixelbytes_tokens =  [
-                ## Bytes (ASCII - UTF8)
-                b'\x00', b'\t', b'\n', b' ', b'"', b"'", b'(', b')', b'*', b',', b'-', b'+', 
+## Bytes (ASCII - UTF8)
+DEFAULT_BYTES = [b'\x00', b'\t', b'\n', b' ', b'"', b"'", b'(', b')', b'*', b',', b'-', b'+', 
                 b'.', b'0', b'1', b'2', b'3', b'4', b'5', b'6', b'7', b'8', b'9', b'\xc2', 
                 b'\xa0', b':', b'[', b']', b';', b'/', b'%', b'!', b'a', b'b', b'c', b'd', b'e', 
                 b'f', b'g', b'h', b'i', b'j', b'k', b'l', b'm', b'n', b'o', b'p', b'q', b'r', 
-                b's', b't', b'u', b'v', b'w', b'x', b'y', b'z',
-                ## Pixel (RGB NES Palette)
-                (0x00, 0x00, 0x00), (0xfc, 0xfc, 0xfc), (0xf8, 0xf8, 0xf8), (0xbc, 0xbc, 0xbc),
-                (0x7c, 0x7c, 0x7c), (0xa4, 0xe4, 0xfc), (0x3c, 0xbc, 0xfc), (0x00, 0x78, 0xf8),
-                (0x00, 0x00, 0xfc), (0xb8, 0xb8, 0xf8), (0x68, 0x88, 0xfc), (0x00, 0x58, 0xf8),
[…]

### 2024-09-10 pixelbytes/model.py
add mamba and change model name

@@ -13,10 +13,11 @@ from torch.nn.utils.rnn import pad_sequence
 from tqdm import tqdm
 import numpy as np, pandas as pd
 from transformers import PreTrainedModel, PretrainedConfig
-from torch.cuda.amp import autocast, GradScaler
-import torch.nn.functional as F
+from mambapy.mamba import Mamba, MambaConfig
+from torch.cuda.amp import GradScaler
 
-class TokenizedDataset(Dataset):
+## Data Part
+class TokenPxByDataset(Dataset):
     def __init__(self, data, tokenizer, seq_length=1024, stride=512):
         self.seq_length = seq_length
         self.stride = stride
@@ -81,32 +82,44 @@ def collate_fn(batch):
     
     return {'input_ids': input_ids_padded,'labels': labels_padded}
 
-class ModelConfig_(PretrainedConfig):
-    model_type = "lstm"
-    def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, auto_regressive=False, **kwargs):
+## Model and training
+class ModelConfig(PretrainedConfig):
+    model_type = "hybrid"
+    def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, 
+                 auto_regressive=False, model_type="lstm", d_conv=4,expand=2,**kwargs):
         super().__init__(**kwargs)
         self.vocab_size = vocab_size
-        self.embed_size = embed_size
-        self.hidden_size = hidden_size
         self.num_layers = num_layers
         self.pxby_dim = pxby_dim
+        self.pxby_emb = embed_size // self.pxby_dim
+        self.embed_size = int(self.pxby_emb * self.pxby_dim)
         self.AR = auto_regressive
+        self.model_type = model_type
+        self.hidden_size = self.d_state = hidden_size
+        # Mamba specific attributes
+        self.d_conv = d_conv
+        self.expand = expand
+    def get_mamba_config(self):
+        return MambaConfig(d_model=self.embed_size, n_layers=self.num_layers, 
+                           d_state=self.d_state, d_conv=self.d_conv, expand_factor=self.expand)
 
-class BestPreTrainedModel(PreTrainedModel):
-    config_class = ModelConfig_
-    base_model_prefix = "lstm"
[…]

### 2024-09-10 pixelbytes/tokenizer.py

@@ -12,13 +12,15 @@ import numpy as np
 from scipy.spatial import cKDTree
 from scipy import stats
 
-## Bytes (ASCII - UTF8)
+## Bytes (ASCII - UTF8) 
+#DEFAULT_BYTES = sorted(set(bytes(c, 'ascii') for c in re.findall(r'[\x00-\x20\x22-\x2f\x30-\x39a-z]', ''.join(map(chr, range(256))))))
 DEFAULT_BYTES = [b'\x00', b'\t', b'\n', b' ', b'"', b"'", b'(', b')', b'*', b',', b'-', b'+', 
                 b'.', b'0', b'1', b'2', b'3', b'4', b'5', b'6', b'7', b'8', b'9', b'\xc2', 
                 b'\xa0', b':', b'[', b']', b';', b'/', b'%', b'!', b'a', b'b', b'c', b'd', b'e', 
                 b'f', b'g', b'h', b'i', b'j', b'k', b'l', b'm', b'n', b'o', b'p', b'q', b'r', 
                 b's', b't', b'u', b'v', b'w', b'x', b'y', b'z']
-## Pixel (RGB NES Palette)
+## Pixel (RGB NES Palette) 
+#DEFAULT_PALETTE = [tuple(int(255*c) for c in colorsys.hsv_to_rgb(h/10,s,v)) for h in range(10) for s in [1,.7] for v in [1,.7,.4]][:50]+[(0,0,0),(252,252,252),(248,248,248),(188,188,188),(120,120,120)]
 DEFAULT_PALETTE = [(0x00, 0x00, 0x00), (0xfc, 0xfc, 0xfc), (0xf8, 0xf8, 0xf8), (0xbc, 0xbc, 0xbc),
                     (0x7c, 0x7c, 0x7c), (0xa4, 0xe4, 0xfc), (0x3c, 0xbc, 0xfc), (0x00, 0x78, 0xf8),
                     (0x00, 0x00, 0xfc), (0xb8, 0xb8, 0xf8), (0x68, 0x88, 0xfc), (0x00, 0x58, 0xf8),

### 2024-09-10 pixelbytes/model.py
adapt expand parameter for pxby embedding

@@ -84,14 +84,13 @@ def collate_fn(batch):
 
 ## Model and training
 class ModelConfig(PretrainedConfig):
-    model_type = "hybrid"
     def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, 
                  auto_regressive=False, model_type="lstm", d_conv=4,expand=2,**kwargs):
         super().__init__(**kwargs)
         self.vocab_size = vocab_size
         self.num_layers = num_layers
         self.pxby_dim = pxby_dim
-        self.pxby_emb = embed_size // self.pxby_dim
+        self.pxby_emb = embed_size // (self.pxby_dim * (expand if model_type=="mamba" else 1))
         self.embed_size = int(self.pxby_emb * self.pxby_dim)
         self.AR = auto_regressive
         self.model_type = model_type

### 2024-09-10 pixelbytes/model.py
delete Sampler CPU

@@ -57,22 +57,6 @@ class TokenPxByDataset(Dataset):
     def get_num_sequences(self, item):
         return max(1, (len(item['input_ids']) - self.seq_length) // self.stride + 1)
 
-class ShuffledSampler(Sampler):
-    def __init__(self, data_source, seed=None):
-        self.data_source = data_source
-        self.seed = seed
-        self.indices = list(range(len(data_source)))
-        
-    def __iter__(self):
-        if self.seed is not None:
-            random.seed(self.seed)
-        random.shuffle(self.indices)
-        for idx in self.indices:
-            yield idx
-
-    def __len__(self):
-        return len(self.data_source)
-
 def collate_fn(batch):
     input_ids = [item['input_ids'] for item in batch]
     labels = [item['labels'] for item in batch]
@@ -241,8 +225,7 @@ if __name__ == '__main__':
     # Préparation des données
     def dataloading(ds):
         dataset = TokenPxByDataset(ds, tokenizer, SEQ_LENGTH, STRIDE)
-        sampler = ShuffledSampler(dataset, seed=42)
-        return DataLoader(dataset, batch_size=BATCH_SIZE, collate_fn=collate_fn, sampler=sampler)
+        return DataLoader(dataset, batch_size=BATCH_SIZE, collate_fn=collate_fn, shuffle=True)
     train_dataloader, val_dataloader = dataloading(train_ds), dataloading(val_ds)
 
     optimizer = optim.AdamW(model.parameters(), lr=LEARNING_RATE)

### 2024-09-10 pixelbytes/model.py
fix bottleneck GPU

@@ -25,12 +25,15 @@ class TokenPxByDataset(Dataset):
         self._preprocess_data(data, tokenizer)
 
     def _preprocess_data(self, data, tokenizer):
+        self.total_sequences = 0
         for item in data:
             tokenized = tokenizer(text=item.get('text'),
                                   image=item.get('image'),
                                   audio=item.get('audio'))
-            self.tokenized_data.append(tokenized)
-        self.total_sequences = sum(self.get_num_sequences(item) for item in self.tokenized_data)
+            length = len(tokenized['input_ids'])
+            num_sequences = max(1, (length - self.seq_length) // self.stride + 1)
+            self.tokenized_data.append((tokenized, length, num_sequences))
+            self.total_sequences += num_sequences
 
     def __len__(self):
         return self.total_sequences
@@ -38,24 +41,21 @@ class TokenPxByDataset(Dataset):
     def __getitem__(self, idx):
         idx = idx % self.total_sequences
         item_idx, start_idx = self._get_item_and_start_indices(idx)
-        tokenized = self.tokenized_data[item_idx]
+        tokenized, length, _ = self.tokenized_data[item_idx]
         end_idx = start_idx + self.seq_length
 
-        input_ids = tokenized['input_ids'][start_idx:end_idx % len(tokenized['input_ids'])]
-        labels = tokenized['labels'][start_idx:end_idx % len(tokenized['labels'])]
-        return {'input_ids': torch.as_tensor(input_ids, dtype=torch.long), 'labels': torch.as_tensor(labels, dtype=torch.long)}
+        input_ids = tokenized['input_ids'][start_idx:end_idx % length]
+        labels = tokenized['labels'][start_idx:end_idx % length]
+        return {'input_ids': torch.as_tensor(input_ids, dtype=torch.long), 
+                'labels': torch.as_tensor(labels, dtype=torch.long)}
 
     def _get_item_and_start_indices(self, idx):
-        for i, item in enumerate(self.tokenized_data):
-            num_sequences = self.get_num_sequences(item)
+        for i, (_, _, num_sequences) in enumerate(self.tokenized_data):
             if idx < num_sequences:
-                start_idx = (idx * self.stride) % len(item['input_ids'])
+                start_idx = (idx * self.stride) % self.tokenized_data[i][1]
                 return i, start_idx
             idx -= num_sequences
         raise IndexError("Index out of range")
-
-    def get_num_sequences(self, item):
[…]

### 2024-09-10 pixelbytes/model.py
refresh code

@@ -110,21 +110,6 @@ class aPxBySequenceModel(PreTrainedModel):
         self.eval()
         with torch.no_grad():
             current_input = input_ids.clone()
-            for i in range(num_generate):
-                outputs = self(current_input)
-                next_token = torch.multinomial(
-                    torch.softmax(outputs[:, -1, -1] if self.AR else outputs[:, -1] / temperature, dim=-1),
-                    num_samples=1)
-                if self.AR:
-                    current_input = torch.cat([current_input, next_token.unsqueeze(1).unsqueeze(1)], dim=2)
-                else:
-                    current_input[:, -(i+1)] = next_token.squeeze(-1)
-            return current_input
-
-    def generate(self, input_ids, num_generate, temperature=1.0):
-        self.eval()
-        with torch.no_grad():
-            current_input = input_ids.clone()
             for i in range(num_generate): # Generate next token
                 outputs = self(current_input)
                 next_token = torch.multinomial(

### 2024-09-10 pixelbytes/model.py
update model (bidirectionnal compatible for autoregressive)

@@ -145,12 +145,11 @@ class aPxBySequenceModel(PreTrainedModel):
             for i, batch in enumerate(tqdm(dataloader, desc="Training" if is_training else "Evaluating")):
                 input_ids, labels = batch['input_ids'].to(device), batch['labels'].to(device)
                 with torch.amp.autocast(device_type='cuda', enabled=scaler is not None):
-                    outputs = self(input_ids)
                     if self.AR: # Reshape and shift for autoregressive mode
-                        outputs = outputs[:, :-1].contiguous().view(-1, self.config.vocab_size)
+                        outputs = self(input_ids[:, :-1]).contiguous().view(-1, self.config.vocab_size)
                         target = input_ids[:, 1:].contiguous().view(-1)
                     else: # Flatten outputs and use labels as target for non-autoregressive mode
-                        outputs = outputs.view(-1, self.config.vocab_size)
+                        outputs = self(input_ids).view(-1, self.config.vocab_size)
                         target = labels.view(-1)
                     loss = criterion(outputs, target)
                     if is_training:

### 2024-09-10 pixelbytes/tokenizer.py
update doc

@@ -142,7 +142,7 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
             padded[1:, 1:-1, :-2]]  # (t, i, j-1) : wrong previous value
         context = torch.stack(slices, dim=-1).reshape(-1, 6)
         targets = context_array.reshape(-1, 1)
-        context[1:,-1] = targets[:-1,0] # true previous value (AutoEncoder)
+        context[1:,-1] = targets[:-1,0] # true previous value (AutoRegressive)
         return context, targets
 
 ### basic test

### 2024-09-11 docs/paper/article.tex
update figure

@@
 % Title page
-\title{\Large\textbf{PixelBytes: Catching Unified Embedding for Multimodal Generation}}
-\author{\large Fabien Furfaro}
+\title{\Large\textbf{PixelBytes: Catching Unified Representation for Multimodal Generation}}
+\author{\large Fabien Furfaro $^{*}$}
 \date{\large 2024}
@@
 \begin{abstract}
-This report introduces PixelBytes Embedding, a novel approach for unified multimodal representation learning.
+This report introduces PixelBytes, a novel approach for unified multimodal representation learning.
 Our method captures diverse inputs in a single, cohesive representation, enabling emergent properties for multimodal sequence generation, particularly for text and pixelated images.
@@
 This work contributes to the advancement of integrated AI models capable of understanding and generating multimodal data in a unified manner.
-Code is available at \url{https://github.com/fabienfrfr}.
+Code is available at \url{https://github.com/fabienfrfr/PixelBytes}.
 \end{abstract}
 
-\let\thefootnote\relax\footnotetext{$^{*}$ Corresponding author: Fabien Furfaro (\url{fabien.furfaro@gmail.com}).}
+\let\thefootnote\relax\footnotetext{$^{*}$ Corresponding author : Fabien Furfaro (\url{fabien.furfaro@gmail.com}).}
 
@@
 \centering
-\includegraphics[width=0.8\linewidth]{example_generation.png}
-\caption{Example of a PixelBytes-generated Pokémon and its description.
-Left: Generated pixelated image.
-Right: Corresponding text description.
-The red box indicates the model's attention window for both pixels and words.
-Note that the model currently exhibits some inaccuracies in image generation and occasionally invents words, highlighting areas for future improvement.}
-\label{fig:example_generation}
+\includegraphics[width=0.8\linewidth]{approach_generation.png}
+\caption{PixelBytes Pokémon generation process.
+Left: Image and text tokenization with embedding window for modality unification.
+Center: Sequence generation using PxBy Embedding.
+Right: Generated pixelated image and text description.
+Red box shows a possible generation example.
+Note: The model currently has some inaccuracies in image generation and may invent words, indicating areas for future improvement.}
+\label{fig:approach_generation}
 \end{figure}
+
+Amélioration de l'idée : il y a trop de sequence qui se repete !
+finaleemnt, la proposition de l'embedding n'est pas bonne !
+les resultats avec plus d'approfondissement ne montre pas vraiment que l'embedding avec un convolution est efficace.
+plutot utiliser les 6 positions specifique (dim entree = dim sortie) et agrandire l'embedding, la génération est mieux (voir Figure 1) !
+de meme simplification du code necessaire !
+le tokenizer doit etre plus flexible, si on a besoin de générer les donner en plus, ca devient compliqué !
+il faut que le tokenizer fasse tout !
+(ou utiliser l'egalité entre la dimension de sortie et de l'entrée pour faire un vrai modele autoregressif) Donc dans l'article, il faut recommancer, il y aura donc deux partie !
+d'abords la premiere version, pour exploration --> on voit que l'embedding n'est pas suffisant, donc on passe à la nouvelle approche !
+(ne pas oublier de comparer avec ImageBind, l'idée n'est pas d'avoir des intermediaire qui ne permette pas l'emergence de propriété)
+
+L'idée n'est pas de faire un modele "ultra performant", sinon pour ces taches, les modeles de diffsion sont mieux (Why Does Diffusion Work Better than Auto-Regression?) maais plutot des mdoele generaux de "fondation" qui ont appris toutes les modalités d'un coup.
+Ici on ne test que les modele auto-regressif, mais des approche comme Diffusion-LM pourrait etre une autre proposition (et compatible avec notre tokenizer)
 
@@
 This is the case with generative language models like GPT \cite{brown2020language}.
-However, these models are limited by their focus on language alone, failing to capture the full complexity of multimodal understanding \cite{alayrac2023language}.
-To address this limitation, researchers have explored combining LLMs with other modalities \cite{radford2021learning, ramesh2022hierarchical}.
+However, these models are limited by their focus on language alone, failing to capture the full complexity of multimodal understanding \cite{huang2023language}.
+To address this limitation, researchers have explored combining LLMs with other modalities \cite{liu2024visual}.
 But this brings us back to the initial problem, as it often results in specialized model combinations without allowing for new emergent properties.
@@
 
-Multimodal sequence generation, which involves the creation of coherent outputs combining various data types such as text, images, and numerical sequences, presents a significant challenge in artificial intelligence \cite{baltrusaitis2019multimodal}.
+Multimodal sequence generation, which involves the creation of coherent outputs combining various data types such as text, images, and numerical sequences, presents a significant challenge in artificial intelligence \cite{baltruvsaitis2018multimodal}.
 While models like GPT have excelled in text generation \cite{brown2020language}, there's a growing need for unified approaches that can seamlessly handle varied data types.
 Building upon these findings, PixelBytes aims to address the challenge of unified text and image generation by proposing a model capable of producing mixed sequences of text and images in a coherent and unified manner.
-We draw inspiration from state-of-the-art sequence models, including Image Transformers \cite{parmar2018image}, PixelCNN \cite{oord2016conditional}, and recent developments in bytes generation by mamba architectures \cite{wang2024mamba}.
+We draw inspiration from state-of-the-art sequence models, including Image Transformers \cite{parmar2018image}, PixelCNN \cite{van2016conditional}, and recent developments in bytes generation by mamba architectures \cite{wang2024mambabyte}.
 
-Our research explores various architectures including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on the effectiveness of bidirectional processing \cite{gu2023visual, gu2023bidirectional}, novel embedding techniques (particularly PxBy embedding, which unifies pixel and byte-level representations), the impact of convolutional layers, effects of model depth, input dimensionality, and size.
+Our research explores various architectures including Recurrent Neural Networks (RNNs), State Space Models (SSMs) \cite{dao2024transformers}, and Attention-based models, focusing on the effectiveness of bidirectional processing \cite{zhu2024vision}, novel embedding techniques (particularly PxBy embedding, which unifies pixel and byte-level representations), the impact of convolutional layers, effects of model depth, input dimensionality, and size.
 Our experiments reveal that bidirectional sequence models with PxBy embedding and convolutional layers demonstrate good performance in generating coherent sequences.
@@
 This paper presents our methodology to construct unified sequence data from text and image, experimental results, and analysis, contributing to the development of our sequence generation embedding models.
-The proposed PixelBytes Embedding approach has the potential to significantly advance the field of multimodal AI, enabling more integrated and versatile models capable of understanding and generating diverse types of data.
-This could lead to improvements in areas such as visual question answering, image captioning, and multimodal content generation for virtual and augmented reality applications.
+The proposed PixelBytes Embedding approach may offer valuable insights for the development of multimodal AI, potentially leading to models that can better understand and generate various types of data.
 
@@
 This is complemented by PxByEmbed, our novel embedding approach that creates a unified representation for both pixel and byte data in a single, coherent space.
-This architecture draws inspiration from recent advancements in multimodal learning \cite{baltrusaitis2019multimodal} and efficient sequence modeling \cite{vaswani2017attention}, allowing PixelBytes to capture intrinsic relationships between visual and textual information effectively.
+This architecture draws inspiration from recent advancements in multimodal learning \cite{baltruvsaitis2018multimodal} and efficient sequence modeling \cite{vaswani2017attention}, allowing PixelBytes to capture intrinsic relationships between visual and textual information effectively.
 
@@
 To address these limitations, we developed a specialized Pokémon dataset, offering advantages such as a long-standing franchise history, pixelated designs, rich descriptive text, and over 1000 unique Pokémon.
-This aligns with trends in creating specialized datasets for multimodal learning \cite{lin2014microsoft}.
 
@@
 For image quantization, we employed a 55-color palette inspired by the NES, creating tokens representing different color combinations.
-This approach translates visual information into a format suitable for sequence modeling, similar to recent work in discrete representation learning for images \cite{van2017neural}.
-The quantization and pixelization process used OpenCV and scikit-image \cite{bradski2000opencv, vanderwalt2014scikit}.
-To ensure balanced representation, we adjusted the number of image and text tokens to 113 characters for each modality, allowing the model to learn equally from both visual and textual information.
+This approach translates visual information into a format suitable for sequence modeling.
+The quantization and pixelization process used OpenCV and scikit-image \cite{bradski2000opencv, van2014scikit}.
+To ensure balanced representation, we adjusted the number of image and text tokens to 113 indices for all entries, allowing the model to learn equally from both visual and textual information.
 The final dataset, balancing text and pixelated images, is available on the Hugging Face Datasets Hub for reproducibility at \url{https://huggingface.co/datasets/ffurfaro/PixelBytes-Pokemon}.
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[ht]
 \caption{PxByEmbed: Multimodal Embedding Algorithm (k=3)
@@
 PixelBytes uses a method to handle transitions between text and image tokens during sequence generation.
-This approach is used in both dataset construction and sequence generation (Figure \ref{fig:example_generation}).
+This approach is used in both dataset construction and sequence generation (Figure \ref{fig:approach_generation}).
 
@@
 \subsection{Model Architectures}
-We evaluated three compact model architectures, each with fewer than 100,000 parameters: a Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{hochreiter1997long}, a Transformer \cite{vaswani2017attention}, and Mamba, based on the State Space Model (SSM) \cite{gu2022efficiently}.
+We evaluated three compact model architectures, each with fewer than 100,000 parameters: a Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{schmidhuber1997long}, a Transformer \cite{vaswani2017attention}, and Mamba, based on the State Space Model (SSM) \cite{dao2024transformers}.
 These models were adapted to process our dataset of pixel data and bytecode sequences.
@@
 Our experiments with various model architectures tested the effectiveness of unified text and image generation.
-We evaluated model performance using Hamming distance, cosine similarity, and BLEU score \cite{papineni2002bleu}.
 Bidirectional RNN models using PixelBytes (PxBy) embedding with convolutional layers generally performed better across our metrics.
 The PxBy embedding improved the mean BLEU score from 0.750 to 0.777 compared to a center-only approach, indicating its effectiveness in representing multimodal data.
-This finding aligns with recent work emphasizing the importance of effective multimodal embeddings in text-to-image generation tasks \cite{yan2024multimodal}.
+This finding aligns with recent work emphasizing the importance of effective multimodal embeddings in text-to-image generation tasks \cite{verHo2021efficient}.
 
@@
 While RNN models showed strong performance in our experiments, State Space Models (SSM) demonstrated rapid convergence during training, suggesting potential for future exploration in multimodal data processing.
-This is in line with recent advancements in SSM for sequence modeling tasks \cite{gu2022efficiently}.
+This is in line with recent advancements in SSM for sequence modeling tasks \cite{dao2024transformers}.
 
@@
 
-This study on PixelBytes highlights the potential for unified text-image generation, contributing to the growing field of multimodal AI \cite{baltrusaitis2019multimodal}.
+This study on PixelBytes highlights the potential for unified text-image generation, contributing to the growing field of multimodal AI \cite{baltruvsaitis2018multimodal}.
 The performance of bidirectional RNN models with PxBy embedding underscores the importance of capturing both spatial and contextual information in multimodal data.
 Future work could focus on refining the PxByEmbed algorithm, testing on larger and more diverse datasets, exploring creative applications of multimodal generation, and further optimizing SSM and Attention models.
-These efforts aim to advance multimodal sequence modeling and generation techniques, addressing challenges identified in surveys of the field \cite{lin2014microsoft}.
+These efforts aim to advance multimodal sequence modeling and generation techniques.
 

### 2024-09-12 pixelbytes/model.py
can change ratio slicing

@@ -181,7 +181,7 @@ if __name__ == '__main__':
     hf_dataset = load_dataset("ffurfaro/PixelBytes-PokemonAll")['train'].train_test_split(test_size=0.1, seed=42)
     train_ds, val_ds = hf_dataset['train'], hf_dataset['test']
     
-    DATA_REDUCTION = 6
+    DATA_REDUCTION = {"image":6, "audio":12}
     tokenizer = ActionPixelBytesTokenizer(data_slicing=DATA_REDUCTION)
     
     # Paramètres

### 2024-09-12 pixelbytes/tokenizer.py

@@ -40,7 +40,7 @@ DEFAULT_ACTION_STATE = np.linspace(-1, 1, 38).tolist()
 
 ##### Tokenizer
 class ActionPixelBytesTokenizer(PreTrainedTokenizer):
-    def __init__(self, data_slicing=1, **kwargs):
+    def __init__(self, data_slicing={"image":1, "audio":1}, **kwargs):
         ## Bytes (ASCII - UTF8) + Pixel (RGB NES Palette) + Action-space (Control & Audio)
         ActionPixelbytes = DEFAULT_BYTES + DEFAULT_PALETTE + DEFAULT_ACTION_STATE
         self.vocab = {ActionPixelbytes[i]: i for i in range(len(ActionPixelbytes))}
@@ -114,7 +114,7 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
             frame = torch.tensor(rgb2lab(np.array(image.convert('RGB')) / 255.0))
             distances = torch.cdist(frame.reshape(-1, 3), self.LabPalette)
             frames[i] = distances.argmin(dim=1).reshape(image.size[::-1])
-        frames = (frames + self.bytes_size)[::self.slicing, ::self.slicing, ::self.slicing]
+        frames = (frames + self.bytes_size)[::self.slicing['image'], ::self.slicing['image'], ::self.slicing['image']]
         added_column = torch.ones((frames.shape[0], frames.shape[1], 1), dtype=torch.long)
         added_column[:, -1, :] = 2
         return torch.cat((frames, added_column), dim=2)
@@ -126,7 +126,7 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         normalized_state = torch.clamp(torch.tensor(stats.zscore(audio.numpy(), axis=1)), -1, 1).flatten()
         action_state_tensor = torch.tensor(DEFAULT_ACTION_STATE).reshape(-1, 1)
         indices = torch.tensor(cKDTree(action_state_tensor.numpy()).query(normalized_state.reshape(-1, 1).numpy())[1]).reshape(audio.shape)
-        indices = (indices + self.bytes_size + self.palet_size)[:, ::self.slicing]
+        indices = (indices + self.bytes_size + self.palet_size)[:, ::self.slicing['audio']]
         separators = torch.tensor([[1], [2]])[:indices.shape[0]]
         return torch.cat([indices, separators], dim=1).T.unsqueeze(1)
 
[…]

### 2024-09-12 pixelbytes/model.py
fix some bug

@@ -69,7 +69,7 @@ def collate_fn(batch):
 ## Model and training
 class ModelConfig(PretrainedConfig):
     def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, 
-                 auto_regressive=False, model_type="lstm", d_conv=4,expand=2,**kwargs):
+                 auto_regressive=True, model_type="lstm", d_conv=4,expand=2,**kwargs):
         super().__init__(**kwargs)
         self.vocab_size = vocab_size
         self.num_layers = num_layers
@@ -177,52 +177,26 @@ if __name__ == '__main__':
     
     def count_parameters_in_k(model):
         return sum(p.numel() for p in model.parameters() if p.requires_grad) / 1000
-
+    
+    # Data
     hf_dataset = load_dataset("ffurfaro/PixelBytes-PokemonAll")['train'].train_test_split(test_size=0.1, seed=42)
     train_ds, val_ds = hf_dataset['train'], hf_dataset['test']
     
     DATA_REDUCTION = {"image":6, "audio":12}
     tokenizer = ActionPixelBytesTokenizer(data_slicing=DATA_REDUCTION)
     
-    # Paramètres
-    VOCAB_SIZE = tokenizer.vocab_size
-    EMBED_SIZE = 128
-    HIDDEN_SIZE = 512
-    NUM_LAYERS = 2
-    PXBY_DIM = 6 # tokenizer
-    AR = False
-    MODEL_TYPE = "lstm"
-    BATCH_SIZE = 32
-    EPOCHS = 10
-    LEARNING_RATE = 0.001
-    DEVICE = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
-    ACCUMULATION_STEPS = 4
+    # Import model
+    model = aPxBySequenceModel.from_pretrained("ffurfaro/aPixelBytes-Pokemon", subfolder="lstm_autoregressive_last")
+    
     SEQ_LENGTH = 1024
     STRIDE = 512
-    
-    # Initialisation du modèle
-    config = ModelConfig(vocab_size=VOCAB_SIZE, embed_size=EMBED_SIZE, hidden_size=HIDDEN_SIZE, 
-                          num_layers=NUM_LAYERS, pxby_dim=PXBY_DIM, auto_regressive=AR, model_type=MODEL_TYPE)
-    model = aPxBySequenceModel(config).to(DEVICE)
[…]

### 2024-09-12 tests.py

@@ -13,7 +13,7 @@ if __name__ == '__main__' :
     # python3 -m pixelbytes.main build
     # python3 -m pixelbytes.main train --seq_length 512 --strides 256
     # generate_arg('ffurfaro/PixelBytes-Pokemon','rnn_bi_pxby_81_dim_64_state_2_layer_last','svg') #dummy !
-    # https://www.perplexity.ai/search/je-cherche-une-base-de-donnee-6CeaKElsSyCN64Qzrmtuvw
+    """
     import numpy as np
     from pydub import AudioSegment
     import matplotlib.pyplot as plt
@@ -71,6 +71,7 @@ if __name__ == '__main__' :
     
     original, processed, original_rate, processed_rate, b, a = process_audio(input_file, output_file)
     plot_signals(original, processed, original_rate, processed_rate, b, a)
+    """
     """
     import matplotlib.pyplot as plt
     from matplotlib.animation import FuncAnimation

### 2024-09-12 pixelbytes/model.py
generalize model and fix generation

@@ -13,7 +13,6 @@ from torch.nn.utils.rnn import pad_sequence
 from tqdm import tqdm
 import numpy as np, pandas as pd
 from transformers import PreTrainedModel, PretrainedConfig
-from mambapy.mamba import Mamba, MambaConfig
 from torch.cuda.amp import GradScaler
 
 ## Data Part
@@ -69,22 +68,16 @@ def collate_fn(batch):
 ## Model and training
 class ModelConfig(PretrainedConfig):
     def __init__(self, vocab_size=2048, embed_size=256, hidden_size=512, num_layers=2, pxby_dim=6, 
-                 auto_regressive=True, model_type="lstm", d_conv=4,expand=2,**kwargs):
+                 auto_regressive=True, model_type="lstm", **kwargs):
         super().__init__(**kwargs)
         self.vocab_size = vocab_size
         self.num_layers = num_layers
         self.pxby_dim = pxby_dim
-        self.pxby_emb = embed_size // (self.pxby_dim * (expand if model_type=="mamba" else 1))
+        self.pxby_emb = embed_size // self.pxby_dim
         self.embed_size = int(self.pxby_emb * self.pxby_dim)
         self.AR = auto_regressive
         self.model_type = model_type
-        self.hidden_size = self.d_state = hidden_size
-        # Mamba specific attributes
-        self.d_conv = d_conv
-        self.expand = expand
-    def get_mamba_config(self):
-        return MambaConfig(d_model=self.embed_size, n_layers=self.num_layers, 
-                           d_state=self.d_state, d_conv=self.d_conv, expand_factor=self.expand)
+        self.hidden_size = hidden_size
 
 class aPxBySequenceModel(PreTrainedModel):
     config_class = ModelConfig
@@ -93,7 +86,7 @@ class aPxBySequenceModel(PreTrainedModel):
     def __init__(self, config):
         super().__init__(config)
         self.embedding = nn.Embedding(config.vocab_size, config.pxby_emb, padding_idx=0)
-        if config.model_type == "mamba": self.sequence_model = Mamba(config.get_mamba_config())
+        if config.model_type != "lstm": self.sequence_model = model_type(**config)
         else: self.sequence_model = nn.LSTM(config.embed_size, config.hidden_size, config.num_layers, batch_first=True)
         self.fc = nn.Linear(config.hidden_size, config.vocab_size * (config.pxby_dim if config.AR else 1))
         self.pxby_dim, self.AR, self.model_type = config.pxby_dim, config.AR, config.model_type
[…]

### 2024-09-13 docs/paper/article.tex
run to publish

@@
 
+\let\thefootnote\relax\footnotetext{$^{*}$ Corresponding author: Fabien Furfaro (\textbf{\href{mailto:fabien.furfaro@gmail.com}{\texttt{fabien.furfaro@capgemini.com}}}).}
+
 \begin{abstract}
 This report introduces PixelBytes, a novel approach for unified multimodal representation learning.
-Our method captures diverse inputs in a single, cohesive representation, enabling emergent properties for multimodal sequence generation, particularly for text and pixelated images.
-Inspired by state-of-the-art sequence models such as Image Transformers, PixelCNN, and Mamba-Bytes, PixelBytes aims to address the challenges of integrating different data types.
-We explore various model architectures, including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on bidirectional processing and our innovative PxBy embedding technique.
-Our experiments, conducted on a specialized PixelBytes Pokémon dataset, demonstrate that bidirectional sequence models with PxBy embedding and convolutional layers can generate coherent multimodal sequences.
-This work contributes to the advancement of integrated AI models capable of understanding and generating multimodal data in a unified manner.
-Code is available at \url{https://github.com/fabienfrfr/PixelBytes}.
+Inspired by existing sequence models such as Image Transformers, PixelCNN, and Mamba-Bytes, our method aims to capture diverse inputs in a cohesive representation, exploring the integration of different data types, particularly text, audio, and pixelated images (sprites).
+We conducted our experiments on a specialized PixelBytes Pokémon dataset.
+Initially, we investigated various model architectures, including Recurrent Neural Networks (RNNs), State Space Models (SSMs), and Attention-based models, focusing on bidirectional processing and our convolutional PxBy embedding technique.
+Subsequently, we shifted our focus to evaluating models based on data reduction strategies and the effectiveness of autoregressive learning.
+We specifically examined Long Short-Term Memory (LSTM) networks in both predictive and autoregressive modes for our main experiments.
+Our findings indicate that autoregressive models outperform predictive models.
+By adopting a flexible approach to multimodal modeling, PixelBytes aims to support the development of foundation models capable of understanding and generating multimodal data.
+Code and pretrained models are available at \url{https://github.com/fabienfrfr/PixelBytes} and \url{https://huggingface.co/ffurfaro/aPixelBytes-Pokemon}.
 \end{abstract}
 
-\let\thefootnote\relax\footnotetext{$^{*}$ Corresponding author : Fabien Furfaro (\url{fabien.furfaro@gmail.com}).}
-
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\linewidth]{approach_generation.png}
-\caption{PixelBytes Pokémon generation process.
-Left: Image and text tokenization with embedding window for modality unification.
-Center: Sequence generation using PxBy Embedding.
-Right: Generated pixelated image and text description.
-Red box shows a possible generation example.
-Note: The model currently has some inaccuracies in image generation and may invent words, indicating areas for future improvement.}
-\label{fig:approach_generation}
+\includegraphics[width=0.8\linewidth]{aPxBy_approach.pdf}
+\caption{Overview of the PixelBytes approach: (Left) The scanning process reads different modalities at specific positions.
+(Center) An example of our dataset and the autoregressive generation across modalities.
+(Right) A generation example, with the PxBy embedding window displayed below.
+Note: The model currently shows some inaccuracies in image generation and may produce incorrect words, highlighting areas for future improvement.}
+\label{fig:approach_overview}
 \end{figure}
-
-Amélioration de l'idée : il y a trop de sequence qui se repete !
-finaleemnt, la proposition de l'embedding n'est pas bonne !
-les resultats avec plus d'approfondissement ne montre pas vraiment que l'embedding avec un convolution est efficace.
-plutot utiliser les 6 positions specifique (dim entree = dim sortie) et agrandire l'embedding, la génération est mieux (voir Figure 1) !
-de meme simplification du code necessaire !
-le tokenizer doit etre plus flexible, si on a besoin de générer les donner en plus, ca devient compliqué !
-il faut que le tokenizer fasse tout !
-(ou utiliser l'egalité entre la dimension de sortie et de l'entrée pour faire un vrai modele autoregressif) Donc dans l'article, il faut recommancer, il y aura donc deux partie !
-d'abords la premiere version, pour exploration --> on voit que l'embedding n'est pas suffisant, donc on passe à la nouvelle approche !
-(ne pas oublier de comparer avec ImageBind, l'idée n'est pas d'avoir des intermediaire qui ne permette pas l'emergence de propriété)
-
-L'idée n'est pas de faire un modele "ultra performant", sinon pour ces taches, les modeles de diffsion sont mieux (Why Does Diffusion Work Better than Auto-Regression?) maais plutot des mdoele generaux de "fondation" qui ont appris toutes les modalités d'un coup.
-Ici on ne test que les modele auto-regressif, mais des approche comme Diffusion-LM pourrait etre une autre proposition (et compatible avec notre tokenizer)
 
@@
 
-Recent advancements in artificial intelligence have led to increasingly generalist models, not by combining multiple specialized components, but by giving simple tasks to models where emergent properties—complex behaviors that arise from simpler underlying rules—appear.
+Recent advancements in artificial intelligence have led to increasingly generalist models, not by combining multiple specialized components (like Gato from DeepMind \cite{reed2022generalist}), but by giving simple tasks to models where emergent properties—complex behaviors that arise from simpler underlying rules—appear.
 This is the case with generative language models like GPT \cite{brown2020language}.
@@
 But this brings us back to the initial problem, as it often results in specialized model combinations without allowing for new emergent properties.
-We propose "PixelBytes Embedding", a novel approach enabling unified training across modalities by capturing diverse inputs in a single, cohesive representation.
+We propose "PixelBytes", a novel approach enabling unified training across modalities by capturing diverse inputs in a single, cohesive representation.
 
@@
 While models like GPT have excelled in text generation \cite{brown2020language}, there's a growing need for unified approaches that can seamlessly handle varied data types.
-Building upon these findings, PixelBytes aims to address the challenge of unified text and image generation by proposing a model capable of producing mixed sequences of text and images in a coherent and unified manner.
+Building upon these findings, PixelBytes aims to address the challenge of unified text, audio and image generation by proposing a model capable of producing mixed sequences of text, audio and images in a coherent and unified manner.
 We draw inspiration from state-of-the-art sequence models, including Image Transformers \cite{parmar2018image}, PixelCNN \cite{van2016conditional}, and recent developments in bytes generation by mamba architectures \cite{wang2024mambabyte}.
@@
 Our research explores various architectures including Recurrent Neural Networks (RNNs), State Space Models (SSMs) \cite{dao2024transformers}, and Attention-based models, focusing on the effectiveness of bidirectional processing \cite{zhu2024vision}, novel embedding techniques (particularly PxBy embedding, which unifies pixel and byte-level representations), the impact of convolutional layers, effects of model depth, input dimensionality, and size.
-Our experiments reveal that bidirectional sequence models with PxBy embedding and convolutional layers demonstrate good performance in generating coherent sequences.
 
-Our results show that it is possible to train sequence models with unified embeddings for both text and pixel completion.
-This paper presents our methodology to construct unified sequence data from text and image, experimental results, and analysis, contributing to the development of our sequence generation embedding models.
-The proposed PixelBytes Embedding approach may offer valuable insights for the development of multimodal AI, potentially leading to models that can better understand and generate various types of data.
+Our experiments, conducted on a specialized PixelBytes Pokémon dataset, reveal that autoregressive models with balanced data reduction strategies outperform predictive models in terms of accuracy and loss.
+While our initial exploration suggested potential for bidirectional sequence models with PxBy embedding and convolutional layers, our final results focus on the performance of Long Short-Term Memory (LSTM) networks in both predictive and autoregressive modes.
+This paper presents our methodology for constructing unified sequence data from text, audio, and pixelated images (sprites), as well as our experimental results and analysis.
+By adopting a flexible approach to multimodal modeling, PixelBytes aims to advance the development of versatile foundation models capable of understanding and generating multimodal data.
 
-\section{Model Architecture}
+\section{Exploration for a Unified Representation}
 
-\subsection{Overview}
-The PixelBytes architecture is designed to seamlessly integrate multimodal data for unified sequence generation.
-At its core, the model incorporates two key innovative components: a specialized tokenizer sequence constructor and a unified multimodal embedding technique called PxByEmbed.
-The tokenizer sequence constructor is engineered to process both pixelated images and text at a byte level, enabling a consistent representation across modalities.
-This is complemented by PxByEmbed, our novel embedding approach that creates a unified representation for both pixel and byte data in a single, coherent space.
-This architecture draws inspiration from recent advancements in multimodal learning \cite{baltruvsaitis2018multimodal} and efficient sequence modeling \cite{vaswani2017attention}, allowing PixelBytes to capture intrinsic relationships between visual and textual information effectively.
+\subsection{Hypothesis Testing Framework}
 
-\subsection{Dataset Construction}
+The quest for a unified data representation across different modalities presents significant challenges.
+Text data typically has a one-dimensional dependency on preceding words with discrete values \cite{bengio2003neural}.
+Audio signals, on the other hand, have a temporal dimension with continuous values \cite{van2016wavenet}.
+Action-state representations in robotics can be similar to audio but with correlations between channels \cite{botteghi2021low}.
+Images and animations combine spatial and temporal dimensions across discrete RGB channels \cite{krizhevsky2012imagenet}.
+Given these diverse characteristics, many researchers have focused on combining separate embedding models into a single framework, as seen in projects like ImageBind \cite{girdhar2023imagebind} and RT-2 \cite{zitkovich2023rt}, rather than seeking a truly unified data representation.
+However, to build a model that comprehensively understands different modalities without intermediate alignment steps, exploring a unified data representation becomes necessary.
+In this section, we will test several hypotheses:
 
-Image captioning datasets, while combining visual and textual modalities, prove unsuitable for joint text and image generation due to limited text content and difficulties in interpreting pixelated versions of high-resolution images.
-To address these limitations, we developed a specialized Pokémon dataset, offering advantages such as a long-standing franchise history, pixelated designs, rich descriptive text, and over 1000 unique Pokémon.
+\begin{itemize}
+    \item Can we quantize data such that each element becomes a token? \cite{zhan2022auto}
+    \item Is predicting only the next value sufficient for a sequence model to learn effectively?
+    \item For dimensions higher than one, is applying a convolutional filter necessary?
+    \item For space-time dependency, what is the importance of bidirectionality in models?
+\end{itemize}
 
-We constructed our dataset by web scraping Pokémon miniatures and descriptions from Pokepedia using Beautiful Soup \cite{richardson2007beautiful}, maintaining a 2/3 text to 1/3 image ratio.
-For image quantization, we employed a 55-color palette inspired by the NES, creating tokens representing different color combinations.
-This approach translates visual information into a format suitable for sequence modeling.
-The quantization and pixelization process used OpenCV and scikit-image \cite{bradski2000opencv, van2014scikit}.
-To ensure balanced representation, we adjusted the number of image and text tokens to 113 indices for all entries, allowing the model to learn equally from both visual and textual information.
-The final dataset, balancing text and pixelated images, is available on the Hugging Face Datasets Hub for reproducibility at \url{https://huggingface.co/datasets/ffurfaro/PixelBytes-Pokemon}.
+Through these investigations, we aim to explore a method of representing data that could enable models to understand various types of modalities.
 
-\section{Multimodal Embedding Algorithm}
+\subsection{Conceptual Multimodal Embedding}
 
-\subsection{PxByEmbed: Multimodal Embedding Algorithm}
+\subsubsection{Dataset Construction}
 
-At the core of our approach is the PxByEmbed algorithm, which represents mixed sequences of text and pixelated images in a unified manner.
-This algorithm builds upon existing embedding techniques by incorporating spatial adaptivity, allowing for representation of both textual and visual information.
-PxByEmbed is designed to address the specific needs of pixel-level image representation and byte-level text encoding.
-The algorithm uses a learned embedding matrix to map each token (text or image) to a vector space, while maintaining spatial relationships for image tokens.
-This approach allows our model to handle both modalities within a single framework.
+To test our hypotheses on unified representation, we needed a dataset combining visual and textual data in a format suitable for byte-level processing.
+Image captioning datasets proved unsuitable due to limited text content and difficulties in interpreting pixelated versions of high-resolution images.
+We therefore created a specialized Pokémon dataset, offering pixelated designs and rich descriptive text.
+We collected data by web scraping Pokémon miniatures and descriptions from Pokepedia using Beautiful Soup \cite{richardson2007beautiful}, maintaining a 2/3 text to 1/3 image ratio.
+For image processing, we employed a 55-color palette inspired by the NES, creating tokens for different color combinations.
+This approach allowed us to represent visual information in a format compatible with our tokenizer.
+
+\sloppy
+
+To handle transitions between text and image tokens, we developed a 2D input sequence method using a 3x3 context window around each token with a 2D zigzag scheme (Figure \ref{fig:approach_overview}).
+Special tokens mark transitions between text and images, with padding added to maintain consistent context sizes.
+The padding value is 0.
+For text, only previous tokens are included in the context windows.
+Using OpenCV and scikit-image \cite{bradski2000opencv, van2014scikit} for image quantization and pixelization, we adjusted all entries to have 113 indices, balancing text and image tokens.
+The resulting dataset, combining text and pixelated images, is available on the Hugging Face Datasets Hub (\url{https://huggingface.co/datasets/ffurfaro/PixelBytes-Pokemon}) for reproducibility.
+It includes a "pixelbyte" column for this specific data representation.
+
+\subsubsection{Embedding Techniques}
+
+Our exploration of unified representation begins with integrating image-text data for sequence generation.
+We developed a tokenizer that processes "pixelbyte" column.
+This is paired with an embedding technique we call PxByEmbed, which creates a unified representation for pixel and byte data in a single space.
+We designed PxByEmbed to test our hypotheses about the effectiveness of convolutional filters for higher-dimensional data.
+PxByEmbed uses a learned embedding matrix to map each token (text or image) to a vector space, while maintaining spatial relationships for image tokens.
+It incorporates a simple convolutional layer and an adaptive mixing mechanism.
+This design allows us to investigate if predicting only the next value (with or without only previous value) is sufficient for effective learning in a sequence model.
 
 \begin{algorithm}[ht]
+\footnotesize
 \caption{PxByEmbed: Multimodal Embedding Algorithm (k=3)
@@
 
-\subsection{Managing Transitions}
+The algorithm processes input sequences in 3x3 patches, embedding each element and then applying a learnable convolution.
+The embedding pad value is set to 0 to avoid influencing the training of subsequent tokens.
+These two representations are then combined using a learned parameter, allowing the model to adaptively balance between local and global information.
 
-PixelBytes uses a method to handle transitions between text and image tokens during sequence generation.
-This approach is used in both dataset construction and sequence generation (Figure \ref{fig:approach_generation}).
+\subsection{Model Architectures Evaluated}
 
-For dataset construction, we use a 2D input sequence method with a 3x3 context window around each token.
-We use special tokens to mark transitions between text and images, and add padding to keep context sizes consistent.
-For text, only previous tokens are included in the context windows.
-The \texttt{input\_seq\_construct} function implements this process.
+We evaluated three compact model architectures: a Recurrent Neural Network (RNN) using Long Short-Term Memory (LSTM) units \cite{schmidhuber1997long}, a Transformer \cite{vaswani2017attention}, and a State Space Model (SSM) based on Mamba \cite{gu2023mamba}.
+For both the RNN and Mamba architectures, we compared versions with and without a bidirectional first layer.
+Each model was constrained to fewer than 100,000 parameters and adapted to process our dataset of pixel data and bytecode sequences.
+We trained the models on Kaggle using T4 GPUs, with a batch size of 32, sequence length of 256, and learning rate of 0.001 for 200 epochs.
+The resulting models are available at \url{https://huggingface.co/ffurfaro/PixelBytes-Pokemon}.
 
-During sequence generation, the \texttt{\_process\_token} function manages transitions.
-It handles text tokens (bytes) and image tokens (tuples) differently, and takes care of special characters like newlines and tabs.
-A \texttt{sequence\_clock} keeps track of the generation progress.
-Each new token's context is represented in a 3x3 matrix, which helps maintain coherence when switching between text and images.
-
-\section{Training and Evaluation}
-
-\subsection{Model Architectures}
-We evaluated three compact model architectures, each with fewer than 100,000 parameters: a Recurrent Neural Network (RNN) using bidirectional Long Short-Term Memory (LSTM) units \cite{schmidhuber1997long}, a Transformer \cite{vaswani2017attention}, and Mamba, based on the State Space Model (SSM) \cite{dao2024transformers}.
-These models were adapted to process our dataset of pixel data and bytecode sequences.
-Training was conducted on Kaggle using dual T4 GPUs, with a batch size of 32, sequence length of 256, stride size of 32, and learning rate of 0.001.
-We trained for 200 epochs, evaluating performance every 5 epochs.
-The resulting models are available at \url{https://huggingface.co/ffurfaro/PixelBytes-Pokemon}.
-Additionally, we developed a more specialized model for generation tasks, which is not presented in this paper.
-
-\begin{figure}[htbp]
+\begin{figure}[ht]
 \centering
-\includegraphics[width=\textwidth]{training_results.png}
-\caption{Training and validation metrics (Loss, Accuracy) for RNN, Transformer, and SSM models over 200 epochs.}
+\includegraphics[width=\textwidth]{training_results_embed.pdf}
+\caption{Training and validation metrics for RNN, Transformer, and SSM models over 200 epochs.}
 \label{fig:training_results}
@@
 
-Figure \ref{fig:training_results} shows the training and validation metrics for our three model types over 200 epochs.
-The State Space Models (SSM) achieved the best scores for loss and accuracy.
-However, the widening gap between their training and validation curves suggests they may be overfitting, meaning they learned the training data well but struggled to generalize to new examples.
+\subsection{Comparative Analysis}
 
-In contrast, the LSTMs (referred to as RNNs in our analysis) demonstrated more balanced performance.
-The closer alignment of their training and validation curves indicates they may generalize better to unseen data.
-The Transformer model had the lowest performance among the three.
-This could be due to the absence of positional encoding trick, which is important for Transformers in sequence tasks.
+Figure \ref{fig:training_results} shows the training and validation metrics for our three model types.
+The SSM achieved the best scores for loss and accuracy but showed signs of overfitting.
+The RNN demonstrated more balanced performance, suggesting better generalization to unseen data.
+The Transformer had the lowest performance, possibly due to the absence of positional encoding.
+The strong performance of the SSM supports the potential of unified representation for pixel and byte data.
+However, its tendency to overfit suggests that predicting only the next value might not be sufficient for effective learning in all cases.
+The balanced performance of the RNN indicates that simpler architectures can still be effective for our task.
+The Transformer's lower performance suggests that complex attention mechanisms may not always be necessary or beneficial for this type of data.
 
-Overall, these results highlight the strengths and weaknesses of each model type.
-While SSMs showed strong training performance, their tendency to overfit suggests that additional regularization techniques might be needed.
-The stable performance of RNNs indicates their suitability for our task, as they balance well between fitting the training data and generalizing to new data.
-The Transformer's performance suggests it may require further adjustments to improve its effectiveness for this task.
+\subsubsection{Generation Evaluation Metrics}
 
-\subsection{Generation Evaluation Metrics}
+To assess the effectiveness of our approach and various model architectures, we evaluated their generation capabilities.
+We tested State Space Models (SSM), Attention models (Att), and Recurrent Neural Networks (RNN) in generating 32 consecutive sequences.
+Our evaluation used three metrics: Hamming Distance \cite{hamming1950error}, Cosine Similarity, and BLEU Score \cite{papineni2002bleu}.
 
-We tested different model types including State Space Models (SSM), Attention models (Att), and Recurrent Neural Networks (RNN) for generating 16 consecutive sequences.
-We used three metrics to evaluate their performance: Hamming Distance, Cosine Similarity, and BLEU Score \cite{sutskever2014sequence, hamming1950error, papineni2002bleu}.
-
-\begin{table}[h]
+\begin{table}[ht]
 \centering
@@
 
-The results show that RNN models with PxBy embedding generally perform better than other types.
-The best Hamming distance (0.149 $\pm$ 0.062) and BLEU score (0.785 $\pm$ 0.082) come from the bidirectional PxBy RNN model with convolution, 81 embedding dimensions, and 32 hidden state dimensions.
-This suggests that a smaller hidden state in the RNN model can effectively capture sequence patterns.
+The results show variations across different model configurations.
+Among RNN models with PxBy embedding, we observed a Hamming distance of 0.149 $\pm$ 0.062 and a BLEU score of 0.785 $\pm$ 0.082 for a bidirectional model with convolution, 81 embedding dimensions, and 32 hidden state dimensions.
+The highest cosine similarity (0.906 $\pm$ 0.087) was achieved by a 3-layer RNN model.
+We noted some differences in performance between models with and without convolution, as well as between unidirectional and bidirectional configurations, though these differences were not always substantial.
+Changing the embedding dimension (36, 81, 162) in RNN models resulted in similar performance levels.
+RNN models using center embedding showed different results compared to those using PxBy embedding.
+The performance of SSM and Attention models varied in comparison to RNN models across the different metrics, but no model type consistently outperformed the others across all measures.
 
-The highest cosine similarity (0.906 $\pm$ 0.087) is from the 3-layer RNN model, indicating that more layers can improve sequence alignment.
-Adding convolution usually improves RNN model performance.
-Interestingly, changing the embedding dimension (36, 81, 162) in RNN models didn't affect performance much, suggesting these models can handle different embedding sizes well.
-The center embedding in RNNs didn't perform as well as PxBy embedding for this task.
+\subsection{Identified Challenges}
 
-\section{Results and Discussion}
+Our initial results revealed several limitations in our embedding approach.
+While we observed variations in performance across different model configurations, the differences were often not substantial \cite{wang2020deep}.
+The PxBy embedding, which we initially thought promising, did not consistently outperform simpler approaches across all metrics and model types.
+Based on these findings, we recognized the need to refine our approach.
+The repetition of sequences in our generated output indicated that our embedding method might not be capturing the full range of patterns in our data \cite{bengio2013representation}.
 
-Our experiments with various model architectures tested the effectiveness of unified text and image generation.
-Bidirectional RNN models using PixelBytes (PxBy) embedding with convolutional layers generally performed better across our metrics.
-The PxBy embedding improved the mean BLEU score from 0.750 to 0.777 compared to a center-only approach, indicating its effectiveness in representing multimodal data.
-This finding aligns with recent work emphasizing the importance of effective multimodal embeddings in text-to-image generation tasks \cite{verHo2021efficient}.
+\section{Optimizing Unified Representation}
 
-Bidirectional models showed a slight advantage over unidirectional ones, with BLEU scores of 0.777 and 0.776 respectively.
-This suggests that bidirectionality may help capture sequential dependencies, consistent with other studies in sequence modeling \cite{schuster1997bidirectional}.
-Notably, varying state dimensions (32, 64, 128) had minimal impact on performance, with BLEU scores ranging from 0.775 to 0.784.
-While RNN models showed strong performance in our experiments, State Space Models (SSM) demonstrated rapid convergence during training, suggesting potential for future exploration in multimodal data processing.
-This is in line with recent advancements in SSM for sequence modeling tasks \cite{dao2024transformers}.
+\subsection{Refined Embedding Approach}
 
-\section{Conclusion and Future Work}
+To address the challenges identified in our initial experiments, we propose a revised embedding strategy.
+Instead of using a convolutional approach, we now focus on six specific positions within each token, with the input dimension equal to the output dimension.
+This change allows for a larger embedding size while potentially improving the model's ability to capture relevant patterns.
+We also recognized the need for a more flexible tokenizer \cite{kudo2018sentencepiece}.
+Our initial implementation became cumbersome when generating new data, highlighting the importance of a more versatile approach.
+We are exploring ways to integrate all necessary functionality into the tokenizer itself, which should simplify our overall pipeline and potentially improve performance.
+These refinements reflect a shift from our initial exploration towards a more focused approach to unified representation.
+While our initial results provided valuable insights, they also revealed the complexities inherent in multimodal sequence modeling and the need for continuous iteration in our methods \cite{lecun2015deep}.
 
-This study on PixelBytes highlights the potential for unified text-image generation, contributing to the growing field of multimodal AI \cite{baltruvsaitis2018multimodal}.
-The performance of bidirectional RNN models with PxBy embedding underscores the importance of capturing both spatial and contextual information in multimodal data.
-Future work could focus on refining the PxByEmbed algorithm, testing on larger and more diverse datasets, exploring creative applications of multimodal generation, and further optimizing SSM and Attention models.
-These efforts aim to advance multimodal sequence modeling and generation techniques.
+\subsubsection{Dataset Construction}
+
+\sloppy
+
+For our refined approach, we created a new dataset combining images, text, and audio extracted from Pokemon sprite animations.
+This dataset, available at \url{https://huggingface.co/datasets/ffurfaro/PixelBytes-PokemonAll}, was compiled through web scraping of Pokepedia and includes descriptions of the Pokemon along with their associated cries.
+The dataset consists of animated, pixelated GIFs of Pokemon sprites as the visual component.
+The audio files are two-channel recordings: Channel 1 contains the original mono sound of the Pokemon cry, while Channel 2 features a filtered version simulating a bits Game Boy speaker output to verify our approach for control problems.
+This setup allows us to model a simplified dynamic physical system, where the original sound acts as the "action" input and the filtered output represents the "state" of the system.
+The transfer function of this bandpass filter can be approximated as:
+
+\begin{equation}
+H(s) = \frac{K \omega_n^2}{s^2 + 2\zeta\omega_n s + \omega_n^2}
+\end{equation}
+
+where $K$ is the gain, $\omega_n$ is the natural frequency, and $\zeta$ is the damping ratio. These parameters can be adjusted to closely match the frequency response of a Game Boy speaker.
+
+\subsection{Enhanced Tokenization Strategy}
+
+Building on our previous work, we developed an improved tokenization strategy using the ActionPixelBytesTokenizer.
+This tokenizer aims to handle multimodal data more effectively, including text, images, and audio, while maintaining a unified representation.
+The tokenizer employs a combined vocabulary that includes ASCII bytes, RGB values from the NES palette, and action states for control and audio.
+This approach attempts to create a consistent representation across different data types.
+For text processing, the tokenizer converts to lowercase ASCII bytes.
+Images are converted to Lab color space and quantized to the nearest NES palette color.
+Audio data is normalized and mapped to predefined action states, were setpoint is reset to zero (standard equilibrium).
+All token numbers are now 151, where index 0 corresponds to the null padding value, and indices 1 and 2 are transition values.
+A key aspect of the new tokenizer is its sequence construction method.
+Instead of using convolutional methods, we focus on six specific positions for each token for no repetition in sequencing of 3D zigzag scheme (Figure \ref{fig:approach_overview}).
+This approach creates context-target pairs that aim to capture relationships between neighboring tokens in both space and time.
+The sequence construction algorithm is detailed below:
+
+\begin{algorithm}[ht]
+\small  % \footnotesize
+\caption{Create Sequence Data
+\newline
+\textbf{Input:} $\mathbf{X} \in \mathbb{R}^{T \times H \times W}$: context array
+\newline
+\textbf{Output:} $\mathbf{C} \in \mathbb{R}^{THW \times 6}$: context, $\mathbf{Y} \in \mathbb{R}^{THW \times 1}$: targets
+\newline
+\textbf{Note:} $T$: time steps, $H$: height, $W$: width
+}
+\begin{algorithmic}[0]
+\State \textbf{Initialize:} $\mathbf{P} \in \mathbb{R}^{(T+1) \times (H+2) \times (W+2)}$ as padded array
+\Function{CreateSequenceData}{$\mathbf{X}$}
+    \State $\mathbf{P} \gets \text{Pad}(\mathbf{X}, (1,1,1,1,1,0), \text{mode='constant'}, \text{value=0})$
+    \State $\mathbf{S}_1 \gets \mathbf{P}_{1:T, 1:H, 2:W+1}$
+    \State $\mathbf{S}_2 \gets \mathbf{P}_{1:T, 2:H+1, 2:W+1}$
+    \State $\mathbf{S}_3 \gets \mathbf{P}_{1:T, 3:H+2, 2:W+1}$
+    \State $\mathbf{S}_4 \gets \mathbf{P}_{2:T+1, 1:H, 1:W}$
+    \State $\mathbf{S}_5 \gets \mathbf{P}_{2:T+1, 1:H, 2:W+1}$
+    \State $\mathbf{S}_6 \gets \mathbf{P}_{2:T+1, 2:H+1, 1:W}$
+    \State $\mathbf{C} \gets \text{Stack}([\mathbf{S}_1, \mathbf{S}_2, \mathbf{S}_3, \mathbf{S}_4, \mathbf{S}_5, \mathbf{S}_6], \text{dim=-1})$
+    \State $\mathbf{C} \gets \text{Reshape}(\mathbf{C}, [THW, 6])$
+    \State $\mathbf{Y} \gets \text{Reshape}(\mathbf{X}, [THW, 1])$
+    \For{$i \gets 2$ \textbf{to} $THW$}
+        \State $\mathbf{C}_{i,6} \gets \mathbf{Y}_{i-1,1}$
+    \EndFor
+    \State \Return $\mathbf{C}, \mathbf{Y}$
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+This algorithm constructs a context array for sequence modeling.
+It pads the input to handle boundary conditions and extracts six slices to represent spatial and temporal relationships.
+The context array $\mathbf{C}$ is formed by stacking these slices, while the targets $\mathbf{Y}$ are derived from the input.
+The last column of $\mathbf{C}$ is set to the true previous value, implementing an autoregressive feature.
+The tokenizer is designed to handle various input combinations and includes functionalities such as special token handling and padding.
+
+\subsection{Autoregressive Model Architecture}
+
+Building upon our previous findings in our initial approach, we developed the aPxBySequenceModel architecture.
+This new architecture is designed to handle both predictive and autoregressive tasks using a Long Short-Term Memory (LSTM) network.
+The model consists of three main components: an embedding layer, an LSTM sequence model, and a fully connected output layer.
+The embedding layer maps input tokens to a continuous vector space, with the embedding size calculated based on the input dimension.
+Specifically, the embedding size is determined by dividing the overall embedding size by the number of positions we focus on within each token.
+This approach aligns with our revised embedding strategy, which emphasizes six specific positions within each token (with padding 0 to not influence training).
+The LSTM layer, aims to capture temporal dependencies in the sequence data, potentially improving pattern recognition across different modalities.
+The model operates in two distinct modes:
+
+\begin{itemize}
+    \item \underline{Predictive mode:} In this configuration, the model takes six input values and attempts to predict only the next token.
+
+    \item \underline{Autoregressive mode:} Here, the model's output dimension matches the input dimension (Figure \ref{fig:approach_overview}).
+Additionally, the output is restructured by multiplying it with the vocabulary size, allowing the model to generate sequences based on the learned representations.
+\end{itemize}
+
+During the forward pass, input data is processed through the embedding layer, then through the LSTM layers, and finally through the fully connected layer.
+The output shape is adjusted based on the operating mode, which may provide the flexibility we found lacking in our initial implementation.
+
+\subsubsection{Model Training and Data Management}
+
+For data management, we developed the \texttt{TokenPxByDataset} class to handle multimodal inputs, including text, image, and audio data.
+This class generates overlapping sequences from longer inputs, which helps the model capture context across sequence boundaries.
+It optimizes memory usage through on-the-fly data retrieval, preparing samples only as they are needed.
+Additionally, the class ensures consistent sequence lengths by implementing circular padding for sequences that extend beyond the end of an item.
+These features enable efficient processing of variable-length inputs during training.
+
+Our training process includes several improvements to enhance efficiency and monitoring.
+In the \texttt{\_process\_epoch} function, we manage both autoregressive and non-autoregressive modes.
+In autoregressive mode, we reshape the input and output sequences, using the input sequence as the target.
+For non-autoregressive mode, we flatten the outputs and use the provided labels as the target.
+This flexibility allows the model to adapt to different tasks.
+The \texttt{train\_model} function alternates between training and validation phases, allowing us to evaluate the model's performance regularly.
+We use gradient accumulation to simulate larger batch sizes, which helps when GPU memory is limited.
+The training loop tracks both training and validation metrics (loss and accuracy) for each epoch and saves these metrics to a CSV file.
+We also implement model checkpointing to keep the best model based on validation loss.
+
+\subsection{Performance Evaluation}
+
+We evaluated three LSTM models: one in predictive mode and two in autoregressive mode, each with approximately 4 million parameters.
+The models were trained on Kaggle using T4 GPUs.
+We used an embedding size of 128, a hidden size of 512, and two layers.
+Training was conducted for 100 epochs with a batch size of 32, a learning rate of 0.001, and a sequence length of 1024.
+
+\subsubsection{Results Comparison}
+
+To manage data proportions, we applied different reduction strategies for image and audio data.
+This was particularly important for audio data in the autoregressive mode, as it contains more null values to predict, which could potentially hinder training.
+Table \ref{tab:model_comparison_lstm} presents the final performance metrics after training.
+
+\begin{table}[ht]
+\centering
+\caption{Comparison of model performance after 100 epochs}
+\label{tab:model_comparison_lstm}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+\textbf{Model} & \textbf{Train Loss} & \textbf{Train Accuracy} & \textbf{Val Loss} & \textbf{Val Accuracy} \\
+\hline
+Autoregressive (2,2) & 0.2211 & 0.9329 & 0.4519 & 0.8852 \\
+Autoregressive (4,2) & 0.2346 & 0.9290 & 0.4914 & 0.8726 \\
+Predictive (2,2) & 0.8810 & 0.7440 & 2.1144 & 0.6009 \\
+\hline
+\end{tabular}
+\end{table}
+
+The numbers in parentheses indicate the reduction factors for (audio, image) data.
+Lower loss and higher accuracy indicate better performance.
+Both autoregressive models performed better than the predictive model in terms of accuracy and loss on both training and validation sets.
+The autoregressive model with balanced reduction (2,2) achieved slightly better results than the one with more aggressive audio reduction (4,2).
+Although the predictive model had lower performance overall, it still reached reasonable accuracy given the complexity of the task.
+This suggests that our pixelbyte representation is effective, especially when using autoregressive learning.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{apxby_gen_comparison.png}
+\caption{Generation results for the 1st Generation Starter Pokémon using the autoregressive model with a temperature of 0.1. Reference images and generated images are paired sequentially.}
+\label{fig:generation_results}
+\end{figure}
+
+The complete frame generation results, as shown in Figure \ref{fig:generation_results}, demonstrate spatial consistency in image creation, despite the model being unidirectional.
+This suggests that data representation may be more crucial than the architecture itself, aligning with recent findings in multimodal learning \cite{baltruvsaitis2018multimodal}.
+The generated images show coherence in colors and sizes, indicating that the model has captured some key visual features.
+While our model does not match the image quality of state-of-the-art diffusion models \cite{dhariwal2021diffusion}, it demonstrates the capability to capture multiple modalities (image, text, audio) within a single framework.
+This aligns with recent trends in multimodal co-learning \cite{liang2022mind}, although further optimization is needed to improve generation quality.
+The trained models are available at \url{https://huggingface.co/ffurfaro/aPixelBytes-Pokemon} for further examination and replication of our results.
+
+\section{Discussion and Future Directions}
+
+Our experiments with various model architectures for unified text and image generation have led to unexpected insights and a shift in our approach.
+Initially, we explored bidirectional RNN models using PixelBytes (PxBy) embedding with convolutional layers, anticipating improved multimodal data representation.
+However, deeper analysis revealed limitations in this approach and these findings have led us to reconsider our tokenizer design.
+
+Our next results with LSTM models have been particularly enlightening.
+The autoregressive models significantly outperformed the predictive model, suggesting that maintaining equal input and output dimensions is important for our task.
+This aligns with recent research emphasizing the importance of preserving structural information in multimodal embeddings \cite{verHo2021efficient}.
+The performance difference between the two autoregressive models highlights the impact of data balancing strategies.
+The model with balanced reduction (2,2) for audio and image data showed slightly better results, indicating that can be overfitting with animated image.
+We now aim for a more versatile solution that can handle all aspects of data preparation and support true autoregressive modeling.
+Our \texttt{TokenPxByDataset} class remains valuable, but we are working to integrate its functionality more closely with our revised tokenizer for a more streamlined data pipeline.
+
+While we initially explored both RNN and State Space Models (SSM), with SSMs showing promising rapid convergence \cite{dao2024transformers}, our focus has shifted towards simplifying the overall architecture.
+Our approach offers an alternative to the principles of models like ImageBind \cite{girdhar2023imagebind}, as our preliminary results suggest it may be possible to unify modalities without relying on intermediate representations.
+We now recognize the potential benefits of allowing emergent properties to develop within a simplified framework.
+Moving forward, we will refine our strategy to better utilize specific input positions for high-definition image and sound.
+We also aim to further simplify our model architecture to promote the emergence of natural multimodal representations.
+Additionally, we plan to explore alternative approaches like Diffusion-LM \cite{li2022diffusion}, which may offer new perspectives on multimodal sequence modeling.
+While our work is still in progress, it aligns with recent trends in multimodal AI research \cite{baltruvsaitis2018multimodal} and could potentially provide a versatile base for various multimodal tasks.
 

### 2024-09-13 pixelbytes/model.py

@@ -76,7 +76,7 @@ class ModelConfig(PretrainedConfig):
         self.pxby_emb = embed_size // self.pxby_dim
         self.embed_size = int(self.pxby_emb * self.pxby_dim)
         self.AR = auto_regressive
-        self.model_type = model_type
+        self.model_type = model_type # Don't forget mask if you want to use transformer
         self.hidden_size = hidden_size
 
 class aPxBySequenceModel(PreTrainedModel):
@@ -169,30 +169,4 @@ class aPxBySequenceModel(PreTrainedModel):
 if __name__ == '__main__':
     from tokenizer import ActionPixelBytesTokenizer
     from datasets import load_dataset
-    
-    def count_parameters_in_k(model):
-        return sum(p.numel() for p in model.parameters() if p.requires_grad) / 1000
-    
-    # Data
-    hf_dataset = load_dataset("ffurfaro/PixelBytes-PokemonAll")['train'].train_test_split(test_size=0.1, seed=42)
-    train_ds, val_ds = hf_dataset['train'], hf_dataset['test']
-    
-    DATA_REDUCTION = {"image":6, "audio":12}
-    tokenizer = ActionPixelBytesTokenizer(data_slicing=DATA_REDUCTION)
-    
-    # Import model
-    model = aPxBySequenceModel.from_pretrained("ffurfaro/aPixelBytes-Pokemon", subfolder="lstm_autoregressive_last")
-    
-    SEQ_LENGTH = 1024
-    STRIDE = 512
-    BATCH_SIZE = 32
-    # Préparation des données
-    def dataloading(ds):
-        dataset = TokenPxByDataset(ds, tokenizer, SEQ_LENGTH, STRIDE)
-        return DataLoader(dataset, batch_size=BATCH_SIZE, collate_fn=collate_fn, shuffle=True)
-    train_dataloader, val_dataloader = dataloading(train_ds), dataloading(val_ds)
-
-    # Test de génération
-    test_input = next(iter(train_dataloader))['input_ids'][:1]
-    generated = model.generate(test_input, 100)
-    print("Generated sequence:", generated)
+    ## some test in test.py file

### 2024-09-13 pixelbytes/tokenizer.py

@@ -68,24 +68,30 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         return [self._convert_token_to_id(token) for token in tokens]
     
     def decode(self, token_ids, skip_special_tokens=True):
+        if token_ids.dim() > 1 : token_ids = token_ids.view(-1,6)[:,-1] # last padded value
+        token_ids = token_ids.tolist()
         tokens = [self._convert_id_to_token(id) for id in token_ids]
-        
-        text_tokens, image_frames, current_frame, audio_tokens = [], [], [], []
+        text_tokens, image_frames, current_frame, current_line, audio_tokens = [], [], [], [], []
         for token in tokens:
             if isinstance(token, bytes):
-                if token == b'\n':
+                if token == b'\t':
+                    if current_line:
+                        current_frame.append(np.array(current_line)) # shape W,3
+                        current_line = []
+                elif token == b'\n':
                     if current_frame:
                         image_frames.append(current_frame)
-                        current_frame = []
+                        current_line, current_frame = [], []
                 elif token in DEFAULT_BYTES:
                     text_tokens.append(token)
             elif isinstance(token, tuple):
-                current_frame.append(token)
+                current_line.append(token)
             elif isinstance(token, (int, float)):
                 audio_tokens.append(token)
-        if current_frame:
+        if current_line:
+            current_frame.append(np.array(current_line))
             image_frames.append(current_frame)
-        return {'text': text_tokens,
+        return {'text': b''.join(text_tokens),
                 'image': image_frames,
                 'audio': audio_tokens}
 
@@ -144,6 +150,15 @@ class ActionPixelBytesTokenizer(PreTrainedTokenizer):
         targets = context_array.reshape(-1, 1)
         context[1:,-1] = targets[:-1,0] # true previous value (AutoRegressive)
         return context, targets
+
+    def construct_images(self, images_token):
+        img_list = []
+        for img_lines in images_token :
+            max_width = max(line.shape[0] for line in img_lines)
+            padded_lines = [np.pad(line, ((0, max_width - line.shape[0]), (0, 0)), 'constant') for line in img_lines]
+            img_list.append(np.stack(padded_lines))
+        return img_list
+    
 
 ### basic test
 if __name__ == '__main__' :

### 2024-09-13 tests.py

@@ -14,6 +14,86 @@ if __name__ == '__main__' :
     # python3 -m pixelbytes.main train --seq_length 512 --strides 256
     # generate_arg('ffurfaro/PixelBytes-Pokemon','rnn_bi_pxby_81_dim_64_state_2_layer_last','svg') #dummy !
     """
+    import matplotlib.pyplot as plt
+    import numpy as np
+    from PIL import Image
+    from scipy.ndimage import gaussian_filter
+    import os
+    
+    folder_path = "pokemon_generated/"
+    file_names = sorted([f for f in os.listdir(folder_path) if f.endswith('.png')])
+    
+    fig, axes = plt.subplots(1, len(file_names), figsize=(15, 5))
+    fig.subplots_adjust(wspace=0)  # Supprime l'espace horizontal entre les sous-plots
+    
+    for i, (ax, file_name) in enumerate(zip(axes, file_names)):
+        img = np.array(Image.open(os.path.join(folder_path, file_name)))
+        non_black = np.where(img.sum(axis=(1, 2)) > 0)[0]
+        img_cropped = img[non_black[0]:non_black[-1] + 1]
+        if i % 2 == 1 and i > 0:
+            prev_height = axes[i-1].images[0].get_array().shape[0]
+            img_cropped = img_cropped[:prev_height]
+            ax.set_yticks([])  # Supprime les graduations de l'axe y
+            ax.set_title("Generated Image", fontsize=10)
+        else:
+            ax.set_title("Original Image", fontsize=10)
+        ax.imshow(img_cropped)
+    
+    plt.tight_layout(pad=0.5)
+    
+    # Enregistrer la figure au format PNG
+    output_filename = "pokemon_comparison.png"
+    plt.savefig(output_filename, format='png', dpi=300, bbox_inches='tight')
+    print(f"Image saved as {output_filename}")
+    
+    # Afficher la figure (optionnel)
+    plt.show()
+    """
+    """
+    def count_parameters_in_k(model):
+        return sum(p.numel() for p in model.parameters() if p.requires_grad) / 1000
+    
+    # Data
+    hf_dataset = load_dataset("ffurfaro/PixelBytes-PokemonAll")['train'].train_test_split(test_size=0.1, seed=42)
+    train_ds, val_ds = hf_dataset['train'], hf_dataset['test']
+    
[…]
